### 2023-07-26 fits.qmd
Start of paper using renv and targets

---
title: "Forecasting COVID-affected time series"
author:
- familyname: Rostami-Tabar
  othernames: Bahman
  address:
    - Cardiff Business School
    - Wales CF10 3EU
    - United Kingdom
  email: rostami-tabarb@cardiff.ac.uk
  correspondingauthor: true
- familyname: Hyndman
  othernames: Rob J
  address:
    - Department of Econometrics & Business Statistics
    - Clayton VIC 3800
    - Australia
  email: Rob.Hyndman@monash.edu
abstract: "A brief summary of our ideas"
keywords: "blah, blah"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: false
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
format:
  wp-pdf:
    knitr:
      opts_chunk:
        dev: "CairoPDF"
fig-height: 5
fig-width: 8
keep-tex: true
number-sections: true
cite-method: biblatex
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r}
#| label: load
#| cache: false
#| eval: true
# Load all required packages
library(targets)
library(fpp3)

# Set ggplot colours and fonts
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)
theme_set(theme_get() + theme(text = element_text(family = "Fira Sans")))

# Load main data objects from targets folder
tar_config_set(store = here::here("_targets"))
```

# Introduction

Not just about covid. How to handle forecasting when there has been a period of major structural change.

Not interested in identifying structural breaks -- we know they have occured. How to forecast in the presence of such change.

Covid has provided one example that is affecting many forecasters around the world. But similar ideas needed in many other contexts.

Time series models usually assume that the future is similar to the past, at least in how the data evolve. But a big event like covid makes the future different from the past, at least in the short term.

# Handling covid effects when forecasting

Some options are:

1. Use a highly adaptive model -- e.g., ETS with large smoothing parameters

Advantages: simple, easy to implement, fast to compute.

Disadvantages: large prediction intervals, because the model is downweighting the past data. Fails to remember things before the change. Also, the model may not adapt quickly enough to the change.

2. Use a dynamic regression model with intervention covariates

Advantages: retains full memory of the past, and allows the change period to be effectively modelled provided you choose the covariates well.

Disadvantages: requires a lot of thought to choose the covariates well. Assumes that the post-pandemic period will be similar to the pre-pandemic period.

3. Treat the covid period as missing and use a model that handles missing values

Advantages: effectively ignores the change period

Disadvantages: throws away the covid-period data, and so starts the post-covid period with no recent history. Therefore prediction intervals will be relatively large.

4. Estimate what might have been and adjust the data

Advantages: relatively simple and doesn't throw away the covid-period data.

Disadvantages: need to have good estimates of covid-period data, and that may be difficult to obtain. Forecasts are conditional on the estimates being correct.

## Example: Ambulance attendances

```{r}
#| label: plot-data
tar_read(attended_plot)
tar_read(stl_plot)
```

```{r}
tar_load(fit)
fit |> filter(Area == "BC") |> select(ets) |> components() |> autoplot()
```

```{r}
#| label: forecast
tar_load(fc)
fc <- forecast(fit, h=365)
fc |> filter(Area == "BC", .model=="ets") |> autoplot(df)
fc |> filter(Area == "BC", .model=="stl_ets") |> autoplot(df)
fc |> filter(Area == "BC", .model=="arima") |> autoplot(df)
```

### 2023-07-26 R/read.R

@@ -1,0 +1,16 @@
+create_tsibble <- function() {
+  # Read data
+  attendant <- readr::read_csv(here::here("data/attendent_incident.csv"))
+  verified <- readr::read_csv(here::here("data/verified_incident.csv"))
+
+  # Create tsibble
+  left_join(
+    attendant |>
+      pivot_longer(-date, names_to = "Area", values_to = "Attended"),
+    verified |>
+      pivot_longer(-date, names_to = "Area", values_to = "Verified"),
+    by = c("date", "Area")
+  ) |>
+    rename(Date = date) |>
+    as_tsibble(index = Date, key = Area)
+}

### 2023-07-26 _dependencies.R

@@ -1,0 +1,8 @@
+# Some library commands to ensure packages end up in renv
+library(tidyverse)
+library(codetools)
+library(languageserver)
+library(styler)
+library(Cairo)
+library(markdown)
+library(httpuv)

### 2023-07-26 _targets.R

@@ -1,0 +1,48 @@
+# Load packages required to define the pipeline:
+library(targets)
+library(tarchetypes)
+
+# Set target options:
+tar_option_set(
+  packages = c("fpp3")
+  # Set other options as needed.
+)
+
+# Run the R scripts in the R/ folder with your custom functions:
+tar_source()
+# source("other_functions.R") # Source other scripts as needed.
+
+# Replace the target list below with your own:
+list(
+  tar_target(df, create_tsibble()),
+  tar_target(attended_plot, autoplot(df, Attended)),
+  tar_target(stl, # STL Decomposition
+    df |>
+      filter(Area == "BC") |>
+      model(
+        stl = STL(Attended ~ trend(window = 81) + season(period = 7, window = Inf)),
+      )
+  ),
+  tar_target(stl_plot,
+    stl |>
+    components() |>
+    autoplot()
+  ),
+  tar_target(stl_model,
+    decomposition_model(
+      STL(Attended ~ trend(window = 81) + season(period = 7, window = Inf)),
+      ETS(season_adjust ~ season("N"))
+    )
+  ),
+  tar_target(fit,
+    df |>
+      model(
+        ets = ETS(Attended),
+        arima = ARIMA(Attended),
+        stl_ets = stl_model
+      )
+    ),
+  tar_target(fc, forecast(fit, h=365)),
+  tar_quarto(paper,
+    "fits.qmd", extra_files = "references.bib")
+)

### 2023-07-26 fits.qmd
Now compiling

@@
 ---
-title: "Forecasting COVID-affected time series"
+title: "Forecasting interrupted time series"
 author:
@@
 tar_load(fc)
-fc <- forecast(fit, h=365)
+tar_load(df)
 fc |> filter(Area == "BC", .model=="ets") |> autoplot(df)

### 2023-07-26 fits.tex
Now compiling

\documentclass[11pt,a4paper,]{article}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={Forecasting interrupted time series},
            pdfkeywords={blah, blah},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{references.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Forecasting interrupted time series}

%% MONASH STUFF

%% CAPTIONS
\RequirePackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\RequirePackage{bera}
\RequirePackage[charter,expert]{mathdesign}
\RequirePackage[scale=0.9]{sourcecodepro}
\RequirePackage[lf,t]{FiraSans}

%% HEADERS AND FOOTERS
\RequirePackage{fancyhdr}
\pagestyle{fancy}
\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
\makeatletter
\lhead{\textsf{\expandafter{\@title}}}
\makeatother
\rhead{}
\cfoot{}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\RequirePackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\RequirePackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%\RequirePackage[section]{placeins}

%% SECTION TITLES
\RequirePackage[compact,sf,bf]{titlesec}
\titleformat{\section}[block]
  {\fontsize{15}{17}\bfseries\sffamily}
  {\thesection}
  {0.4em}{}
\titleformat{\subsection}[block]
  {\fontsize{12}{14}\bfseries\sffamily}
  {\thesubsection}
  {0.4em}{}
\titlespacing{\section}{0pt}{*5}{*1}
\titlespacing{\subsection}{0pt}{*2}{*0.2}

%% TITLE PAGE
\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}
        \placefig{2}{1.5}{width=5cm}{_extensions/numbats/wp/monash2}
        \placefig{16.9}{1.5}{width=2.1cm}{_extensions/numbats/wp/MBSportrait}
        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
        \begin{textblock}{7}(12.7,27.9)\hfill
        \includegraphics[height=0.7cm]{_extensions/numbats/wp/AACSB}~~~
        \includegraphics[height=0.7cm]{_extensions/numbats/wp/EQUIS}~~~
        \includegraphics[height=0.7cm]{_extensions/numbats/wp/AMBA}
        \end{textblock}
        \vspace*{2cm}
        \begin{center}\Large
        Department of Econometrics and Business Statistics\\[.5cm]
        \footnotesize http://monash.edu/business/ebs/research/publications
        \end{center}\vspace{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE\@author\end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}\vspace*{2cm}}}
\def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
\let\maketitle\titlepage
\makeatother

%% SPACING
\RequirePackage{setspace}
\spacing{1.5}

%% LINE AND PAGE BREAKING
\sloppy
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}

%% PARAGRAPH BREAKS
\setlength{\parskip}{1.4ex}
\setlength{\parindent}{0em}

%% HYPERLINKS
\RequirePackage{xcolor} % Needed for links
\definecolor{darkblue}{rgb}{0,0,.6}
\RequirePackage{url}

\makeatletter
\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
\makeatother
\hypersetup{
     citecolor=0 0 0,
     breaklinks=true,
     bookmarksopen=true,
     bookmarksnumbered=true,
     linkcolor=darkblue,
     urlcolor=blue,
     citecolor=darkblue,
     colorlinks=true}

%% KEYWORDS
\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}

%% ABSTRACT
\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
  {\end{minipage}}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage[showonlyrefs]{mathtools}
\usepackage[no-weekday]{eukdate}

%% BIBLIOGRAPHY

\makeatletter
\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
\makeatother
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false, date=year}

\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
% No dot before number of articles
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}

\makeatletter
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\makeatother
\renewcommand*{\finalnamedelim}{%
  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
  \addspace\&\space}

\wp{no/yr}
\jel{C10,C14,C22}

\RequirePackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}
\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}

\nocover

\author{Bahman~Rostami-Tabar, Rob J~Hyndman}
\addresses{\textbf{Bahman Rostami-Tabar}\newline
Cardiff Business School\newline
Wales CF10 3EU\newline
United Kingdom\newline
{Email: rostami-tabarb@cardiff.ac.uk}\newline Corresponding author\newline\\[0.5cm]
\textbf{Rob J Hyndman}\newline
Department of Econometrics \& Business Statistics\newline
Clayton VIC 3800\newline
Australia\newline
{Email: Rob.Hyndman@monash.edu}\\[0.5cm]
}

\date{\sf\Date~\Month~\Year}
\makeatletter
 \lfoot{\sf Rostami-Tabar, Hyndman: \@date}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
\makeatother
\makeatletter
\makeatother

% Adjust headwidth in case user has changed geometry in header-includes
\renewcommand{\headwidth}{\textwidth}

\begin{document}
\maketitle
\begin{abstract}
A brief summary of our ideas
\end{abstract}
\begin{keywords}
blah, blah
\end{keywords}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Not just about covid. How to handle forecasting when there has been a
period of major structural change.

Not interested in identifying structural breaks -- we know they have
occured. How to forecast in the presence of such change.

Covid has provided one example that is affecting many forecasters around
the world. But similar ideas needed in many other contexts.

Time series models usually assume that the future is similar to the
past, at least in how the data evolve. But a big event like covid makes
the future different from the past, at least in the short term.

\hypertarget{handling-covid-effects-when-forecasting}{%
\section{Handling covid effects when
forecasting}\label{handling-covid-effects-when-forecasting}}

Some options are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Use a highly adaptive model -- e.g., ETS with large smoothing
  parameters
\end{enumerate}

Advantages: simple, easy to implement, fast to compute.

Disadvantages: large prediction intervals, because the model is
downweighting the past data. Fails to remember things before the change.
Also, the model may not adapt quickly enough to the change.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Use a dynamic regression model with intervention covariates
\end{enumerate}

Advantages: retains full memory of the past, and allows the change
period to be effectively modelled provided you choose the covariates
well.

Disadvantages: requires a lot of thought to choose the covariates well.
Assumes that the post-pandemic period will be similar to the
pre-pandemic period.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Treat the covid period as missing and use a model that handles missing
  values
\end{enumerate}

Advantages: effectively ignores the change period

Disadvantages: throws away the covid-period data, and so starts the
post-covid period with no recent history. Therefore prediction intervals
will be relatively large.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Estimate what might have been and adjust the data
\end{enumerate}

Advantages: relatively simple and doesn't throw away the covid-period
data.

Disadvantages: need to have good estimates of covid-period data, and
that may be difficult to obtain. Forecasts are conditional on the
estimates being correct.

\hypertarget{example-ambulance-attendances}{%
\subsection{Example: Ambulance
attendances}\label{example-ambulance-attendances}}

\includegraphics{fits_files/figure-pdf/plot-data-1.pdf}

\includegraphics{fits_files/figure-pdf/plot-data-2.pdf}

\includegraphics{fits_files/figure-pdf/unnamed-chunk-3-1.pdf}

\includegraphics{fits_files/figure-pdf/forecast-1.pdf}

\includegraphics{fits_files/figure-pdf/forecast-2.pdf}

\includegraphics{fits_files/figure-pdf/forecast-3.pdf}

\printbibliography

\end{document}

### 2023-07-26 _targets.R

@@ -12,7 +12,7 @@ tar_option_set(
 tar_source()
 # source("other_functions.R") # Source other scripts as needed.
 
-# Replace the target list below with your own:
+# List of targets
 list(
   tar_target(df, create_tsibble()),
   tar_target(attended_plot, autoplot(df, Attended)),

### 2023-07-27 fits.qmd
updates

@@
   message: false
-  cache: true
+  cache: false
 ---
@@
 
-Not just about covid. How to handle forecasting when there has been a period of major structural change.
+Time series are sometimes interrupted by unusual events; for example, a natural disaster may occur, or a there may be a temporary policy change.
+Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns may have been severely disrupted due to lockdowns and other restrictions.
+In this paper, we consider the problem of forecasting after such an event has occurred.
 
-Not interested in identifying structural breaks -- we know they have occured. How to forecast in the presence of such change.
+This is a different problem from change point detection.
+In the situations we consider, we know that a change has occurred, and we want to forecast the future after the change.
+Change point detection is about identifying when a change has occurred.
 
-Covid has provided one example that is affecting many forecasters around the world. But similar ideas needed in many other contexts.
+The changes in the series as a result of the disruption may be relatively simple, for example a left shift at the start of the disruption, and another at the end of the disruption.
+Or they may be more complex, with changes to the seasonal patterns, and changes to the level, which evolve over time.
+In this paper, we consider a range of models that can be used to handle such changes, and compare their performance on some real data sets.
 
@@
 
-# Handling covid effects when forecasting
+# Handling interruptions when forecasting
 
-Some options are:
+We consider several possible ways to handle interruptions when forecasting.
 
-1. Use a highly adaptive model -- e.g., ETS with large smoothing parameters
+## Use a highly adaptive model
 
-Advantages: simple, easy to implement, fast to compute.
+Highly adaptive models can adjust to the interruption as it is happening, and will therefore be able to approximate the data generating process relatively well.
+For example, an ETS model with large smoothing parameters will be able to adjust to the interruption relatively quickly.
+This has the advantage of being a very simple solution that is easy to implement and fast to compute.
 
-Disadvantages: large prediction intervals, because the model is downweighting the past data. Fails to remember things before the change. Also, the model may not adapt quickly enough to the change.
+However, the prediction intervals will be large, because the model will have heavily discounted past data.
+In fact, the model will largely forget the past data other than the most recent observations, so there is no memory of the seasonal patterns and other dynamics that were present before the interruption.
+Consequently, the approach works best if there is no assumption that the post-interruption period will be similar to the pre-interruption period.
 
-2. Use a dynamic regression model with intervention covariates
+## Use a dynamic regression model with intervention covariates
+
+A dynamic regression model with intervention covariates can be used to model the interruption explicitly.
+For example, if the intervention involves a simple level shift, with a reverse level shift at the end of the intervention, we can use a dummy variable to indicate the interruption period, and allow the model to adjust to the interruption.
+More complicated interventions can be handled by using more covariates.
+
+This has the advantage of retaining the memory of the past, and so the seasonal patterns and other dynamics will be retained.
+However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be be too narrow.
 
@@
 
-3. Treat the covid period as missing and use a model that handles missing values
+## Treat the covid period as missing and use a model that handles missing values
 
@@
 
-4. Estimate what might have been and adjust the data
+## Estimate what might have been and adjust the data
 
@@
 Disadvantages: need to have good estimates of covid-period data, and that may be difficult to obtain. Forecasts are conditional on the estimates being correct.
+
+# Examples
 

### 2023-07-27 fits.tex
updates

@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
 \makeatother
@@
 
-Not just about covid. How to handle forecasting when there has been a
-period of major structural change.
+Time series are sometimes interrupted by unusual events; for example, a
+natural disaster may occur, or a there may be a temporary policy change.
+Many forecasters have faced this issue recently with the COVID-19
+pandemic, where historical patterns may have been severely disrupted due
+to lockdowns and other restrictions. In this paper, we consider the
+problem of forecasting after such an event has occurred.
 
-Not interested in identifying structural breaks -- we know they have
-occured. How to forecast in the presence of such change.
+This is a different problem from change point detection. In the
+situations we consider, we know that a change has occurred, and we want
+to forecast the future after the change. Change point detection is about
+identifying when a change has occurred.
 
-Covid has provided one example that is affecting many forecasters around
-the world. But similar ideas needed in many other contexts.
+The changes in the series as a result of the disruption may be
+relatively simple, for example a left shift at the start of the
+disruption, and another at the end of the disruption. Or they may be
+more complex, with changes to the seasonal patterns, and changes to the
+level, which evolve over time. In this paper, we consider a range of
+models that can be used to handle such changes, and compare their
+performance on some real data sets.
 
@@
 
-\hypertarget{handling-covid-effects-when-forecasting}{%
-\section{Handling covid effects when
-forecasting}\label{handling-covid-effects-when-forecasting}}
+\hypertarget{handling-interruptions-when-forecasting}{%
+\section{Handling interruptions when
+forecasting}\label{handling-interruptions-when-forecasting}}
 
-Some options are:
+We consider several possible ways to handle interruptions when
+forecasting.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Use a highly adaptive model -- e.g., ETS with large smoothing
-  parameters
-\end{enumerate}
+\hypertarget{use-a-highly-adaptive-model}{%
+\subsection{Use a highly adaptive
+model}\label{use-a-highly-adaptive-model}}
 
-Advantages: simple, easy to implement, fast to compute.
+Highly adaptive models can adjust to the interruption as it is
+happening, and will therefore be able to approximate the data generating
+process relatively well. For example, an ETS model with large smoothing
+parameters will be able to adjust to the interruption relatively
+quickly. This has the advantage of being a very simple solution that is
+easy to implement and fast to compute.
 
-Disadvantages: large prediction intervals, because the model is
-downweighting the past data. Fails to remember things before the change.
-Also, the model may not adapt quickly enough to the change.
+However, the prediction intervals will be large, because the model will
+have heavily discounted past data. In fact, the model will largely
+forget the past data other than the most recent observations, so there
+is no memory of the seasonal patterns and other dynamics that were
+present before the interruption. Consequently, the approach works best
+if there is no assumption that the post-interruption period will be
+similar to the pre-interruption period.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  Use a dynamic regression model with intervention covariates
-\end{enumerate}
+\hypertarget{use-a-dynamic-regression-model-with-intervention-covariates}{%
+\subsection{Use a dynamic regression model with intervention
+covariates}\label{use-a-dynamic-regression-model-with-intervention-covariates}}
+
+A dynamic regression model with intervention covariates can be used to
+model the interruption explicitly. For example, if the intervention
+involves a simple level shift, with a reverse level shift at the end of
+the intervention, we can use a dummy variable to indicate the
+interruption period, and allow the model to adjust to the interruption.
+More complicated interventions can be handled by using more covariates.
+
+This has the advantage of retaining the memory of the past, and so the
+seasonal patterns and other dynamics will be retained. However, the
+model will assume that the post-interruption period will be similar to
+the pre-interruption period, and so the prediction intervals may be be
+too narrow.
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  Treat the covid period as missing and use a model that handles missing
-  values
-\end{enumerate}
+\hypertarget{treat-the-covid-period-as-missing-and-use-a-model-that-handles-missing-values}{%
+\subsection{Treat the covid period as missing and use a model that
+handles missing
+values}\label{treat-the-covid-period-as-missing-and-use-a-model-that-handles-missing-values}}
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  Estimate what might have been and adjust the data
-\end{enumerate}
+\hypertarget{estimate-what-might-have-been-and-adjust-the-data}{%
+\subsection{Estimate what might have been and adjust the
+data}\label{estimate-what-might-have-been-and-adjust-the-data}}
 
@@
 estimates being correct.
+
+\hypertarget{examples}{%
+\section{Examples}\label{examples}}
 

### 2023-07-27 R/read.R

@@ -1,8 +1,4 @@
-create_tsibble <- function() {
-  # Read data
-  attendant <- readr::read_csv(here::here("data/attendent_incident.csv"))
-  verified <- readr::read_csv(here::here("data/verified_incident.csv"))
-
+create_tsibble <- function(attendant, verified) {
   # Create tsibble
   left_join(
     attendant |>

### 2023-07-27 _targets.R

@@ -14,7 +14,11 @@ tar_source()
 
 # List of targets
 list(
-  tar_target(df, create_tsibble()),
+  tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format="file"),
+  tar_target(verified_csv, here::here("data/verified_incident.csv"), format="file"),
+  tar_target(attendant, readr::read_csv(attendant_csv)),
+  tar_target(verified, readr::read_csv(verified_csv)),
+  tar_target(df, create_tsibble(attendant, verified)),
   tar_target(attended_plot, autoplot(df, Attended)),
   tar_target(stl, # STL Decomposition
     df |>

### 2023-07-27 fits.qmd
Removed Cairo

@@
 )
-theme_set(theme_get() + theme(text = element_text(family = "Fira Sans")))
+#theme_set(theme_get() + theme(text = element_text(family = "Fira Sans")))
 
@@
 
+## Australian tourism
+
 ## Example: Ambulance attendances

### 2023-07-27 fits.tex
Removed Cairo

@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 \makeatother
@@
 
+\hypertarget{australian-tourism}{%
+\subsection{Australian tourism}\label{australian-tourism}}
+
 \hypertarget{example-ambulance-attendances}{%

### 2023-07-27 _dependencies.R

@@ -3,6 +3,6 @@ library(tidyverse)
 library(codetools)
 library(languageserver)
 library(styler)
-library(Cairo)
 library(markdown)
 library(httpuv)
+library(quarto)

### 2023-07-27 fits.qmd
Added tourism example

@@
 biblio-style: authoryear-comp
-format:
-  wp-pdf:
-    knitr:
-      opts_chunk:
-        dev: "CairoPDF"
+format: wp-pdf
 fig-height: 5
@@
 
+```{r austa}
+tar_read(tourism_plot1)
+tar_read(tourism_plot2)
+```
+
 ## Example: Ambulance attendances

### 2023-07-27 fits.tex
Added tourism example

@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, enhanced, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
 \makeatother
@@
 
+\includegraphics{fits_files/figure-pdf/austa-1.pdf}
+
+\includegraphics{fits_files/figure-pdf/austa-2.pdf}
+
 \hypertarget{example-ambulance-attendances}{%

### 2023-07-27 R/tourism.R

@@ -1,0 +1,30 @@
+read_tourism <- function(file) {
+  readxl::read_excel(file, sheet = "Data1", skip = 9) |>
+    rename(date = `Series ID`, value = A85375847A) |>
+    select(date, value) |>
+    transmute(
+      Month = yearmonth(date),
+      Visitors = value / 1e3
+    ) |>
+    as_tsibble(index = Month) |>
+    filter(Month >= yearmonth("2000 Jan"))
+}
+
+tourism_plot <- function(austa) {
+  austa |>
+    as_tibble() |>
+    mutate(Month = as.Date(Month)) |>
+    ggplot(aes(x = Month, y = Visitors)) +
+    geom_line() +
+    labs(
+      x = "Month",
+      y = "Thousands of visitors",
+      title = "Total short-term visitors to Australia"
+    ) +
+    scale_x_date(
+      breaks = seq(as.Date("2000-01-01"), by = "5 years", l = 5),
+      labels = paste("Jan", seq(2000, 2020, by = 5)),
+      minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
+    ) 
+}
+

### 2023-07-27 _targets.R

@@ -14,6 +14,22 @@ tar_source()
 
 # List of targets
 list(
+  # Tourism example
+  tar_target(tourism_csv, here::here("data/340101.xlsx"), format="file"),
+  tar_target(austa, read_tourism(tourism_csv)),
+  # Fit ETS model
+  tar_target(fit_tourism,  
+    austa |>
+      filter(Month < yearmonth("2020 Jan")) |>
+      model(ETS(Visitors))),
+  tar_target(tourism_plot1, tourism_plot(austa)),
+  # ETS forecasts
+  tar_target(fc_tourism, forecast(fit_tourism, h = 48)),
+  tar_target(tourism_plot2, tourism_plot1 + 
+               autolayer(fc_tourism |> mutate(Month = as.Date(Month))) + 
+               geom_line()),
+                              
+  # Attended incidents example
   tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format="file"),
   tar_target(verified_csv, here::here("data/verified_incident.csv"), format="file"),
   tar_target(attendant, readr::read_csv(attendant_csv)),

### 2023-07-27 fits.qmd
updates

@@
 
-Advantages: effectively ignores the change period
+Many time series models will handle periods of missing values.
+So the problematic observations that occur during the period of disruption can be set to missing, and the model should continue to produce forecasts as if the interruption had not occurred.
+Of course, the forecasts will not be accurate for the period of disruption, but they can be interpreted as "what might have been".
+This solution requires a judgement to be made about when the disruption has begun, and when normality resumes.
 
-Disadvantages: throws away the covid-period data, and so starts the post-covid period with no recent history. Therefore prediction intervals will be relatively large.
+Because no information is retained during the disruption, the prediction intervals will become large during the disruption, and after the disruption they will remain large until the model has enough data to estimate the forecast distribution more accurately.
 
@@
 
-Advantages: relatively simple and doesn't throw away the covid-period data.
-
-Disadvantages: need to have good estimates of covid-period data, and that may be difficult to obtain. Forecasts are conditional on the estimates being correct.
+A fourth solution is to estimate what might have been during the period of disruption, and then use the adjusted data to fit a model.
+The estimates could be made using any convenient method.
+One approach would be to set the estimates to equal the forecasts made using only pre-interruption data.
+This then becomes almost equal to the previous solution except that the prediction intervals will be narrower because the estimation uncertainty has not been taken into account.
 
@@
 
+Let's consider some examples
+
 ## Australian tourism
 
-```{r austa}
+The Australian tourism data set is a monthly time series showing the number of short-term overseas visitors to Australia.
+The data are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1
+
+```{r}
+#| label: fig-tourism-plot1
+#| caption: Australian tourism data"
 tar_read(tourism_plot1)

### 2023-07-27 fits.qmd
More work on tourism example

@@
 The Australian tourism data set is a monthly time series showing the number of short-term overseas visitors to Australia.
-The data are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1
+The data are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1.
+As the borders closed in March 2020, the number of visitors to Australia dropped to near zero, and remained there until towards the end of 2021.
+The borders officially reopened on 21 February 2022, although it seems visitors began to arrive earlier than that.
+The data are available from the Australian Bureau of Statistics.
 
 ```{r}
-#| label: fig-tourism-plot1
+#| label: fig-tourism_plot1
 #| caption: Australian tourism data"
 tar_read(tourism_plot1)
-tar_read(tourism_plot2)
+```
+
+We will apply each of the approaches above to this data set, and compare their performance.
+
+```{r}
+#| label: fig-tsol1_plot
+#| fig-height: 8
+tar_read(tsol1_plot)
+```
+
+```{r}
+#| label: fig-tsol3_plot
+#| fig-height: 8
+tar_read(tsol3_plot)
 ```

### 2023-07-27 R/tourism.R

@@ -28,3 +28,46 @@ tourism_plot <- function(austa) {
     ) 
 }
 
+tourism_sol1 <- function(austa) {
+  austa |> 
+    stretch_tsibble(.step = 12, .init = 240) |> 
+    model(
+      ets = ETS(Visitors),
+      arima = ARIMA(Visitors)
+    ) |> 
+    forecast(h=12)
+}
+
+tourism_sol3 <- function(austa) {
+austa |> 
+  stretch_tsibble(.step = 12, .init = 240) |> 
+  mutate(
+    Visitors = if_else(Month >= yearmonth("2020 Mar") & Month <= yearmonth("2022 Nov"),
+                       NA_real_, Visitors)
+  ) |> 
+  model(
+    arima = ARIMA(Visitors)
+  ) |> 
+  forecast(h=12)
+}
+
+tourism_sol2 <- function(austa) {
+  austa_stretch <- austa |> 
+    mutate(
+      covid = as.numeric(Month >= yearmonth("2020 Mar")),
+      recovery = Month >= yearmonth("2022 Nov"),
+      recovery = recovery * (Month - yearmonth("2022 Oct"))
+    ) |> 
+    stretch_tsibble(.step = 12, .init = 240) 
+  fit1 <- austa_stretch |> 
+    model(arima = ARIMA(Visitors))
+  fit2 <- austa_stretch |>
+    model(arima = ARIMA(Visitors ~ covid))
+  fit3 <- austa_stretch |>
+    model(arima = ARIMA(Visitors ~ covid + recovery))
+  
+  #  newd <- new_data(austa, h = 48) 
+    
+  #  forecast(h=12)
+  #fc 
+}

### 2023-07-27 _dependencies.R

@@ -6,3 +6,4 @@ library(styler)
 library(markdown)
 library(httpuv)
 library(quarto)
+library(urca)

### 2023-07-27 _targets.R

@@ -28,7 +28,13 @@ list(
   tar_target(tourism_plot2, tourism_plot1 + 
                autolayer(fc_tourism |> mutate(Month = as.Date(Month))) + 
                geom_line()),
-                              
+  tar_target(tsol1, tourism_sol1(austa)),                            
+  tar_target(tsol3, tourism_sol3(austa)),                            
+  tar_target(tsol1_plot, autoplot(tsol1) + 
+     geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
+  tar_target(tsol3_plot, autoplot(tsol3) + 
+      geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
+
   # Attended incidents example
   tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format="file"),
   tar_target(verified_csv, here::here("data/verified_incident.csv"), format="file"),

### 2023-07-27 fits.qmd
Still working on tourism

@@
 biblio-style: authoryear-comp
-format: wp-pdf
+format: 
+  wp-pdf:
+    include-in-header: header.tex
+cite-method: biblatex
+number-sections: true
+keep-tex: true
 fig-height: 5
 fig-width: 8
-keep-tex: true
-number-sections: true
-cite-method: biblatex
 execute:
@@
 #| caption: Australian tourism data"
+#| fig-height: 3
 tar_read(tourism_plot1)
@@
 #| fig-height: 8
+#| fig-caption: Solution 1. ETS is highly adaptive.
 tar_read(tsol1_plot)
+```
+
+```{r}
+#| label: fig-tsol2_plot
+#| fig-height: 8
+#| fig-caption: Solution 2. We use a left shift in March 2020, and a ramp from November 2022.
+tar_read(tsol2_plot)
 ```
@@
 #| fig-height: 8
+#| fig-caption: Solution 3. The period from March 2020 to November 2022 is set to missing.
 tar_read(tsol3_plot)
 ```
+
+```{r}
+#| label: fig-tsol4_plot
+#| fig-height: 8
+#| fig-caption: Solution 4. The period from March 2020 to November 2022 is set to missing.
+tar_read(tsol4_plot)
+```
+
+\FloatBarrier
 

### 2023-07-27 fits.tex
Still working on tourism

--- 2023-07-27 fits.tex
@@
 
+\usepackage[section]{placeins}
 \makeatletter
@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, sharp corners, frame hidden, enhanced]}{\end{tcolorbox}}\fi
 \makeatother
@@
 
-Advantages: effectively ignores the change period
+Many time series models will handle periods of missing values. So the
+problematic observations that occur during the period of disruption can
+be set to missing, and the model should continue to produce forecasts as
+if the interruption had not occurred. Of course, the forecasts will not
+be accurate for the period of disruption, but they can be interpreted as
+``what might have been''. This solution requires a judgement to be made
+about when the disruption has begun, and when normality resumes.
 
-Disadvantages: throws away the covid-period data, and so starts the
-post-covid period with no recent history. Therefore prediction intervals
-will be relatively large.
+Because no information is retained during the disruption, the prediction
+intervals will become large during the disruption, and after the
+disruption they will remain large until the model has enough data to
+estimate the forecast distribution more accurately.
 
@@
 
-Advantages: relatively simple and doesn't throw away the covid-period
-data.
-
-Disadvantages: need to have good estimates of covid-period data, and
-that may be difficult to obtain. Forecasts are conditional on the
-estimates being correct.
+A fourth solution is to estimate what might have been during the period
+of disruption, and then use the adjusted data to fit a model. The
+estimates could be made using any convenient method. One approach would
+be to set the estimates to equal the forecasts made using only
+pre-interruption data. This then becomes almost equal to the previous
+solution except that the prediction intervals will be narrower because
+the estimation uncertainty has not been taken into account.
 
@@
 
+Let's consider some examples
+
 \hypertarget{australian-tourism}{%
 \subsection{Australian tourism}\label{australian-tourism}}
+
+The Australian tourism data set is a monthly time series showing the
+number of short-term overseas visitors to Australia. The data are
+available from January 2000 to May 2023, and are shown in
+Figure~\ref{fig-tourism_plot1}. As the borders closed in March 2020, the
+number of visitors to Australia dropped to near zero, and remained there
+until towards the end of 2021. The borders officially reopened on 21
+February 2022, although it seems visitors began to arrive earlier than
+that. The data are available from the Australian Bureau of Statistics.
+
+\begin{figure}
+
+{\centering \includegraphics{fits_files/figure-pdf/fig-tourism_plot1-1.pdf}
+
+}
+
+\caption{\label{fig-tourism_plot1}\textbf{?(caption)}}
+
+\end{figure}
+
+We will apply each of the approaches above to this data set, and compare
+their performance.
+
+\begin{figure}
+
+{\centering \includegraphics{fits_files/figure-pdf/fig-tsol1_plot-1.pdf}
+
+}
+
+\caption{\label{fig-tsol1_plot}\textbf{?(caption)}}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{fits_files/figure-pdf/fig-tsol2_plot-1.pdf}
+
+}
+
+\caption{\label{fig-tsol2_plot}\textbf{?(caption)}}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{fits_files/figure-pdf/fig-tsol3_plot-1.pdf}
+
+}
+
+\caption{\label{fig-tsol3_plot}\textbf{?(caption)}}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{fits_files/figure-pdf/fig-tsol4_plot-1.pdf}
+
+}
+
+\caption{\label{fig-tsol4_plot}\textbf{?(caption)}}
+
+\end{figure}
+
+\FloatBarrier
 
@@
 
-\includegraphics{fits_files/figure-pdf/unnamed-chunk-3-1.pdf}
+\includegraphics{fits_files/figure-pdf/unnamed-chunk-8-1.pdf}
 

### 2023-07-27 R/tourism.R

@@ -25,49 +25,110 @@ tourism_plot <- function(austa) {
       breaks = seq(as.Date("2000-01-01"), by = "5 years", l = 5),
       labels = paste("Jan", seq(2000, 2020, by = 5)),
       minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
-    ) 
+    )
 }
 
 tourism_sol1 <- function(austa) {
-  austa |> 
-    stretch_tsibble(.step = 12, .init = 240) |> 
+  austa |>
+    stretch_tsibble(.step = 12, .init = 240) |>
     model(
       ets = ETS(Visitors),
       arima = ARIMA(Visitors)
-    ) |> 
-    forecast(h=12)
+    ) |>
+    forecast(h = 12)
 }
 
-tourism_sol3 <- function(austa) {
-austa |> 
-  stretch_tsibble(.step = 12, .init = 240) |> 
-  mutate(
-    Visitors = if_else(Month >= yearmonth("2020 Mar") & Month <= yearmonth("2022 Nov"),
-                       NA_real_, Visitors)
-  ) |> 
-  model(
-    arima = ARIMA(Visitors)
-  ) |> 
-  forecast(h=12)
-}
 
 tourism_sol2 <- function(austa) {
-  austa_stretch <- austa |> 
+  austa_stretch <- austa |>
     mutate(
       covid = as.numeric(Month >= yearmonth("2020 Mar")),
       recovery = Month >= yearmonth("2022 Nov"),
       recovery = recovery * (Month - yearmonth("2022 Oct"))
-    ) |> 
-    stretch_tsibble(.step = 12, .init = 240) 
-  fit1 <- austa_stretch |> 
-    model(arima = ARIMA(Visitors))
+    ) |>
+    stretch_tsibble(.step = 12, .init = 240)
+  # Fit models using none, one or two covariates
+  fit1 <- austa_stretch |>
+    model(arima1 = ARIMA(Visitors))
   fit2 <- austa_stretch |>
-    model(arima = ARIMA(Visitors ~ covid))
+    model(arima2 = ARIMA(Visitors ~ covid)) |> 
+    suppressWarnings()
   fit3 <- austa_stretch |>
-    model(arima = ARIMA(Visitors ~ covid + recovery))
-  
-  #  newd <- new_data(austa, h = 48) 
-    
-  #  forecast(h=12)
-  #fc 
+    model(arima3 = ARIMA(Visitors ~ covid + recovery)) |> 
[…]

### 2023-07-27 _targets.R

@@ -29,12 +29,18 @@ list(
                autolayer(fc_tourism |> mutate(Month = as.Date(Month))) + 
                geom_line()),
   tar_target(tsol1, tourism_sol1(austa)),                            
+  tar_target(tsol2, tourism_sol2(austa)),                            
   tar_target(tsol3, tourism_sol3(austa)),                            
+  tar_target(tsol4, tourism_sol4(austa)),   
   tar_target(tsol1_plot, autoplot(tsol1) + 
+     geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
+  tar_target(tsol2_plot, autoplot(tsol2) + 
      geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
   tar_target(tsol3_plot, autoplot(tsol3) + 
       geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
-
+  tar_target(tsol4_plot, autoplot(tsol4) + 
+               geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
+  
   # Attended incidents example
   tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format="file"),
   tar_target(verified_csv, here::here("data/verified_incident.csv"), format="file"),

### 2023-07-28 fits.qmd
Added ABS ref
Adjustments to tourism examples

@@
 The Australian tourism data set is a monthly time series showing the number of short-term overseas visitors to Australia.
-The data are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1.
+The data [@tourismdata] are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1.
 As the borders closed in March 2020, the number of visitors to Australia dropped to near zero, and remained there until towards the end of 2021.
@@
 #| label: fig-tourism_plot1
-#| caption: Australian tourism data"
+#| fig-cap: "Short-term visitor arrivals to Australia (monthly): Jan 2000 -- May 2023."
 #| fig-height: 3
@@
 #| fig-height: 8
-#| fig-caption: Solution 1. ETS is highly adaptive.
+#| fig-cap: Solution 1. This should show that ETS is highly adaptive. However, for some reason, that is no longer the case, and I don't know why.
 tar_read(tsol1_plot)
@@
 #| fig-height: 8
-#| fig-caption: Solution 2. We use a left shift in March 2020, and a ramp from November 2022.
+#| fig-cap: Solution 2. We use a left shift in March 2020, and a ramp from November 2022 to June 2023.
 tar_read(tsol2_plot)
@@
 #| fig-height: 8
-#| fig-caption: Solution 3. The period from March 2020 to November 2022 is set to missing.
+#| fig-cap: Solution 3.
+The period from March 2020 to November 2022 is set to missing.
+So the first three years of forecasts show what might have been.
+The fourth set of forecasts uses observations from November and December 2022, and so the forecasts have been adjusted downwards.
 tar_read(tsol3_plot)
@@
 #| fig-height: 8
-#| fig-caption: Solution 4. The period from March 2020 to November 2022 is set to missing.
+#| fig-cap: Solution 4. We replace the observations from March 2020 to November 2022 with the average of the same month in the 3 preceding years. 
 tar_read(tsol4_plot)

### 2023-07-28 fits.tex
Adjustments to tourism examples

@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, sharp corners, frame hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, frame hidden, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 \makeatother
@@
 The Australian tourism data set is a monthly time series showing the
-number of short-term overseas visitors to Australia. The data are
-available from January 2000 to May 2023, and are shown in
-Figure~\ref{fig-tourism_plot1}. As the borders closed in March 2020, the
-number of visitors to Australia dropped to near zero, and remained there
-until towards the end of 2021. The borders officially reopened on 21
-February 2022, although it seems visitors began to arrive earlier than
-that. The data are available from the Australian Bureau of Statistics.
+number of short-term overseas visitors to Australia. The data
+\autocite{tourismdata} are available from January 2000 to May 2023, and
+are shown in Figure~\ref{fig-tourism_plot1}. As the borders closed in
+March 2020, the number of visitors to Australia dropped to near zero,
+and remained there until towards the end of 2021. The borders officially
+reopened on 21 February 2022, although it seems visitors began to arrive
+earlier than that. The data are available from the Australian Bureau of
+Statistics.
 
@@
 
-\caption{\label{fig-tourism_plot1}\textbf{?(caption)}}
+\caption{\label{fig-tourism_plot1}Short-term visitor arrivals to
+Australia (monthly): Jan 2000 -- May 2023.}
 
@@
 
-\caption{\label{fig-tsol1_plot}\textbf{?(caption)}}
+\caption{\label{fig-tsol1_plot}Solution 1. This should show that ETS is
+highly adaptive. However, for some reason, that is no longer the case,
+and I don't know why.}
 
@@
 
-\caption{\label{fig-tsol2_plot}\textbf{?(caption)}}
+\caption{\label{fig-tsol2_plot}Solution 2. We use a left shift in March
+2020, and a ramp from November 2022. We possibly need to stop the ramp
+around mid 2023.}
 
@@
 
-\caption{\label{fig-tsol3_plot}\textbf{?(caption)}}
+\caption{\label{fig-tsol3_plot}Solution 3. The period from March 2020 to
+November 2022 is set to missing. So the first three years of forecasts
+show what might have been. The fourth set of forecasts uses observations
+from November and December 2022, and so the forecasts have been adjusted
+downwards.}
 
@@
 
-\caption{\label{fig-tsol4_plot}\textbf{?(caption)}}
+\caption{\label{fig-tsol4_plot}Solution 4. We replace the observations
+from March 2020 to November 2022 with the average of the same month in
+the 3 preceding years.}
 

### 2023-07-28 R/tourism.R

@@ -35,15 +35,14 @@ tourism_sol1 <- function(austa) {
       ets = ETS(Visitors),
       arima = ARIMA(Visitors)
     ) |>
-    forecast(h = 12)
+    forecast(h = 12) 
 }
-
 
 tourism_sol2 <- function(austa) {
   austa_stretch <- austa |>
     mutate(
       covid = as.numeric(Month >= yearmonth("2020 Mar")),
-      recovery = Month >= yearmonth("2022 Nov"),
+      recovery = Month >= yearmonth("2022 Nov") & Month <= yearmonth("2023 Jun"),
       recovery = recovery * (Month - yearmonth("2022 Oct"))
     ) |>
     stretch_tsibble(.step = 12, .init = 240)
@@ -85,7 +84,7 @@ tourism_sol2_fc <- function(fit, h = h) {
   newd <- new_data(training_data, n = h) |>
     mutate(
       covid = as.numeric(Month >= yearmonth("2020 Mar")),
-      recovery = Month >= yearmonth("2022 Nov"),
+      recovery = Month >= yearmonth("2022 Nov") & Month <= yearmonth("2023 Jun"),
       recovery = recovery * (Month - yearmonth("2022 Oct"))
     )
   fit |>
@@ -120,15 +119,15 @@ tourism_sol4 <- function(austa) {
     mutate(month = month(Month)) |> 
     left_join(ave_3y, by="month") |> 
     mutate(
-      Vis = if_else(Month < yearmonth("2022 Nov") &
+      Visitors_adj = if_else(Month < yearmonth("2022 Nov") &
                          Month > yearmonth("2020 Feb"),
           ave, Visitors),
     )
   austa |> 
     stretch_tsibble(.step = 12, .init = 240) |>
     model(
-      ets = ETS(Visitors),
-      arima = ARIMA(Visitors)
+      ets = ETS(Visitors_adj),
+      arima = ARIMA(Visitors_adj)
     ) |>
     forecast(h = 12)
 }

### 2023-07-28 fits.qmd
More work on the tourism example

@@
 tar_config_set(store = here::here("_targets"))
+tar_load(c(austa))
 ```
@@
 The borders officially reopened on 21 February 2022, although it seems visitors began to arrive earlier than that.
-The data are available from the Australian Bureau of Statistics.
 
@@
 
-We will apply each of the approaches above to this data set, and compare their performance.
+We apply the four solutions discussed in the previous section to these data, making 12 month forecasts at the end of each year from 2019 to 2022.
+We fit ETS, ARIMA and dynamic ARIMA models to the data [@fpp3], first applying a log transformation to ensure the resulting forecasts are positive.
+This is possible, because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> filter(Visitors == min(Visitors)) |> pull(Month) |> format("%B %Y")`.
+For all forecasts, we also show 90% prediction intervals.
 
@@
 #| fig-height: 8
+#| fig-pos: "t"
 #| fig-cap: Solution 1. This should show that ETS is highly adaptive. However, for some reason, that is no longer the case, and I don't know why.
@@
 
+In @fig-tsol1_plot, we show forecasts by applying ETS and ARIMA models to the data.
+ETS, in particular, is well-known to be relatively adaptive to changes in the series, and this is evident in these forecasts.
+The forecasts for 2020 were made using data before COVID-19 had any affect, and so they show similar patterns to the past.
+The forecasts for 2021 were made after 9 months of very low levels of arrivals, and both models show forecasts consistent with the recent history.
+The use of logarithms is particularly important here as the variance is much smaller during 2020 than previously, but after taking logarithms, the variance is more stable over time.
+The forecasts made at the start of 2022 have struggled to detect the small increase in traffic at the end of 2021, and both models have forecast relatively flat trajectories as a result, although the prediction intervals are wide indicating model uncertainty.
+Finally, the forecasts made at the end of 2022 have captured the increasing trend, but ETS is much closer to the reality, adapting more quickly to the changing patterns.
+Again, the wide prediction intervals indicate a high level of uncertainty.
+It is possible to make ETS more adaptive to changes in the data by increasing the value of the smoothing parameters.
+For example, a high value of $\beta$ (the smoothing parameter for the slope), will result in changes of trend being incorporated into the forecasts more quickly, at the risk of over-reacting to noise in the data.
+
 ```{r}
@@
 #| fig-height: 8
-#| fig-cap: Solution 2. We use a left shift in March 2020, and a ramp from November 2022 to June 2023.
+#| fig-pos: "t"
+#| fig-cap: Solution 2. We use a left shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022. After October 2022, there are no non-zero predictors.
 tar_read(tsol2_plot)
 ```
+
+@fig-tsol2_plot shows forecasts obtained using a dynamic regression model (i.e., a regression with ARIMA errors) using two intervention variables: a level shift from March 2022 to November 2022, and a ramp from October 2021 to November 2022.
+It is evident that the level shift variable has worked well, giving relatively good forecasts for 2021.
+However, the forecasts for 2022 are particularly poor, because the ramp slope has been greatly overestimated, as it was based on only three observations (Oct -- Dec 2021).
+The forecasts for 2023 are much better, and the relatively large prediction intervals are appropriate given the uncertainty in the industry at the end of 2022.
 
@@
 #| fig-height: 8
+#| fig-pos: "t"
 #| fig-cap: Solution 3.
-The period from March 2020 to November 2022 is set to missing.
+The period from March 2020 to October 2022 is set to missing.
 So the first three years of forecasts show what might have been.
@@
 
+The third solution involved setting the observations during the disruption period to missing, and then fitting an ARIMA model to the series.
+Consequently, the first three years of forecasts show what might have been without the COVID-19 pandemic, based on the history to the end of 2019.
+The final forecasts for 2023 use the data to the end of 2019, and the two observations in November and December 2022.
+These are much better, but the prediction intervals are too narrow, as there hasn't been sufficient recent data.
+
 ```{r}
@@
 #| fig-height: 8
-#| fig-cap: Solution 4. We replace the observations from March 2020 to November 2022 with the average of the same month in the 3 preceding years. 
+#| fig-pos: "t"
+#| fig-cap: Solution 4. We replace the observations from March 2020 to October 2022 with the average of the same month in the three years prior to March 2020.
 tar_read(tsol4_plot)
 ```
+
+Finally, we show in @fig-tsol4_plot the forecasts obtained using solution 4.
+Here we have replaced observations between March 2020 and October 2022 with estimates based on the average of the same month in the three years prior to March 2020.
+The resulting forecasts are similar to those from solution 3, but with narrower prediction intervals because the model is (falsely) assuming that the "observations" during the disruption are real.
 

### 2023-07-28 fits.tex
More work on the tourism example

@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, frame hidden, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
 \makeatother
@@
 reopened on 21 February 2022, although it seems visitors began to arrive
-earlier than that. The data are available from the Australian Bureau of
-Statistics.
+earlier than that.
 
@@
 
-We will apply each of the approaches above to this data set, and compare
-their performance.
+We apply the four solutions discussed in the previous section to these
+data, making 12 month forecasts at the end of each year from 2019 to
+2022. We fit ETS, ARIMA and dynamic ARIMA models to the data
+\autocite{fpp3}, first applying a log transformation to ensure the
+resulting forecasts are positive. This is possible, because the
+observations never reach exactly zero, with the smallest number of
+visitors per month equal to 2250 in April 2020.
 
-\begin{figure}
+\begin{figure}[t]
 
@@
 
-\begin{figure}
+In Figure~\ref{fig-tsol1_plot}, we show forecasts by applying ETS and
+ARIMA models to the data. ETS, in particular, is well-known to be
+relatively adaptive to changes in the series, and this is evident in
+these forecasts. The forecasts for 2020 were made using data before
+COVID-19 had any affect, and so they show similar patterns to the past.
+The forecasts for 2021 were made after 9 months of very low levels of
+arrivals, and both models show forecasts consistent with the recent
+history. The use of logarithms is particularly important here as the
+variance is much smaller during 2020 than previously, but after taking
+logarithms, the variance is more stable over time. The forecasts made at
+the start of 2022 have struggled to detect the small increase in traffic
+at the end of 2021, and both models have forecast relatively flat
+trajectories as a result, although the prediction intervals are wide
+indicating model uncertainty. Finally, the forecasts made at the end of
+2022 have captured the increasing trend, but ETS is much closer to the
+reality, adapting more quickly to the changing patterns. Again, the wide
+prediction intervals indicate a high level of uncertainty. It is
+possible to make ETS more adaptive to changes in the data by increasing
+the value of the smoothing parameters. For example, a high value of
+\(\beta\) (the smoothing parameter for the slope), will result in
+changes of trend being incorporated into the forecasts more quickly, at
+the risk of over-reacting to noise in the data.
+
+\begin{figure}[t]
 
@@
 
-\caption{\label{fig-tsol2_plot}Solution 2. We use a left shift in March
-2020, and a ramp from November 2022. We possibly need to stop the ramp
-around mid 2023.}
+\caption{\label{fig-tsol2_plot}Solution 2. We use a left shift from
+March 2020 to October 2022, and a ramp from October 2021 to October
+2022. After October 2022, there are no non-zero predictors.}
 
@@
 
-\begin{figure}
+Figure~\ref{fig-tsol2_plot} shows forecasts obtained using a dynamic
+regression model (i.e., a regression with ARIMA errors) using two
+intervention variables: a level shift from March 2022 to November 2022,
+and a ramp from October 2021 to November 2022. It is evident that the
+level shift variable has worked well, giving relatively good forecasts
+for 2021. However, the forecasts for 2022 are particularly poor, because
+the ramp slope has been greatly overestimated, as it was based on only
+three observations (Oct -- Dec 2021). The forecasts for 2023 are much
+better, and the relatively large prediction intervals are appropriate
+given the uncertainty in the industry at the end of 2022.
+
+\begin{figure}[t]
 
@@
 \caption{\label{fig-tsol3_plot}Solution 3. The period from March 2020 to
-November 2022 is set to missing. So the first three years of forecasts
+October 2022 is set to missing. So the first three years of forecasts
 show what might have been. The fourth set of forecasts uses observations
@@
 
-\begin{figure}
+The third solution involved setting the observations during the
+disruption period to missing, and then fitting an ARIMA model to the
+series. Consequently, the first three years of forecasts show what might
+have been without the COVID-19 pandemic, based on the history to the end
+of 2019. The final forecasts for 2023 use the data to the end of 2019,
+and the two observations in November and December 2022. These are much
+better, but the prediction intervals are too narrow, as there hasn't
+been sufficient recent data.
+
+\begin{figure}[t]
 
@@
 \caption{\label{fig-tsol4_plot}Solution 4. We replace the observations
-from March 2020 to November 2022 with the average of the same month in
+from March 2020 to October 2022 with the average of the same month in
 the 3 preceding years.}
@@
 \end{figure}
+
+Finally, we show in Figure~\ref{fig-tsol4_plot} the forecasts obtained
+using solution 4. Here we have replaced observations between March 2020
+and October 2022 with estimates based on the average of the same month
+in the three preceding years. The resulting forecasts are similar to
+those from solution 3, but with narrower prediction intervals because
+the model is (falsely) assuming that the ``observations'' during the
+disruption are real.
 

### 2023-07-28 R/tourism.R

@@ -16,11 +16,6 @@ tourism_plot <- function(austa) {
     mutate(Month = as.Date(Month)) |>
     ggplot(aes(x = Month, y = Visitors)) +
     geom_line() +
-    labs(
-      x = "Month",
-      y = "Thousands of visitors",
-      title = "Total short-term visitors to Australia"
-    ) +
     scale_x_date(
       breaks = seq(as.Date("2000-01-01"), by = "5 years", l = 5),
       labels = paste("Jan", seq(2000, 2020, by = 5)),
@@ -32,28 +27,31 @@ tourism_sol1 <- function(austa) {
   austa |>
     stretch_tsibble(.step = 12, .init = 240) |>
     model(
-      ets = ETS(Visitors),
-      arima = ARIMA(Visitors)
+      ets = ETS(log(Visitors)),
+      arima = ARIMA(log(Visitors))
     ) |>
-    forecast(h = 12) 
+    forecast(h = 12) |>
+    mutate(.id = paste("Forecasts of", .id + 2019))
 }
 
 tourism_sol2 <- function(austa) {
   austa_stretch <- austa |>
     mutate(
-      covid = as.numeric(Month >= yearmonth("2020 Mar")),
-      recovery = Month >= yearmonth("2022 Nov") & Month <= yearmonth("2023 Jun"),
-      recovery = recovery * (Month - yearmonth("2022 Oct"))
+      covid = as.numeric(Month >= yearmonth("2020 Mar") & 
+                         Month <= yearmonth("2022 Oct")),
+      recovery = Month >= yearmonth("2021 Oct") & 
+                 Month <= yearmonth("2022 Oct"),
+      recovery = recovery * (Month - yearmonth("2021 Sep"))
     ) |>
     stretch_tsibble(.step = 12, .init = 240)
   # Fit models using none, one or two covariates
   fit1 <- austa_stretch |>
-    model(arima1 = ARIMA(Visitors))
+    model(arima1 = ARIMA(log(Visitors)))
   fit2 <- austa_stretch |>
-    model(arima2 = ARIMA(Visitors ~ covid)) |> 
+    model(arima2 = ARIMA(log(Visitors) ~ covid)) |>
     suppressWarnings()
   fit3 <- austa_stretch |>
-    model(arima3 = ARIMA(Visitors ~ covid + recovery)) |> 
+    model(arima3 = ARIMA(log(Visitors) ~ covid + recovery)) |>
     suppressWarnings()
[…]

### 2023-07-28 _targets.R

@@ -22,25 +22,35 @@ list(
     austa |>
       filter(Month < yearmonth("2020 Jan")) |>
       model(ETS(Visitors))),
-  tar_target(tourism_plot1, tourism_plot(austa)),
+  tar_target(tourism_plot1, tourism_plot(austa) + tourism_labels),
+  tar_target(tourism_labels, labs(
+    x = "Month",
+    y = "Thousands of visitors",
+    title = "Total short-term visitors to Australia"
+  )),
+  tar_target(tourism_history, 
+             geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), 
+                       aes(y=Visitors))
+  ),
   # ETS forecasts
   tar_target(fc_tourism, forecast(fit_tourism, h = 48)),
-  tar_target(tourism_plot2, tourism_plot1 + 
-               autolayer(fc_tourism |> mutate(Month = as.Date(Month))) + 
-               geom_line()),
   tar_target(tsol1, tourism_sol1(austa)),                            
   tar_target(tsol2, tourism_sol2(austa)),                            
   tar_target(tsol3, tourism_sol3(austa)),                            
   tar_target(tsol4, tourism_sol4(austa)),   
-  tar_target(tsol1_plot, autoplot(tsol1) + 
-     geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
-  tar_target(tsol2_plot, autoplot(tsol2) + 
-     geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
-  tar_target(tsol3_plot, autoplot(tsol3) + 
-      geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
-  tar_target(tsol4_plot, autoplot(tsol4) + 
-               geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), aes(y=Visitors))),
-  
+  tar_target(tsol1_plot, 
+    autoplot(tsol1, level=90) + tourism_history + tourism_labels
+  ),
+  tar_target(tsol2_plot, 
+    autoplot(tsol2, level=90) +  tourism_history + tourism_labels
+  ),
+  tar_target(tsol3_plot, 
+    autoplot(tsol3, level=90)  + tourism_history + tourism_labels
+  ),
+  tar_target(tsol4_plot, 
+    autoplot(tsol4, level=90) + tourism_history + tourism_labels
+  ),
+
[…]

### 2023-07-28 _targets_packages.R
Ran tar_renv()

@@ -1,0 +1,13 @@
+# Generated by targets::tar_renv(): do not edit by hand
+library(bs4Dash)
+library(clustermq)
+library(fpp3)
+library(future)
+library(gt)
+library(markdown)
+library(pingr)
+library(rstudioapi)
+library(shiny)
+library(shinybusy)
+library(shinyWidgets)
+library(visNetwork)

### 2023-07-28 fits.qmd
Added Monash affiliation

@@
     - Department of Econometrics & Business Statistics
+    - Monash University
     - Clayton VIC 3800
@@
 Time series are sometimes interrupted by unusual events; for example, a natural disaster may occur, or a there may be a temporary policy change.
-Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns may have been severely disrupted due to lockdowns and other restrictions.
-In this paper, we consider the problem of forecasting after such an event has occurred.
+Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to lockdowns and other restrictions.
+In this paper, we consider the problem of forecasting during such an event, and after it has occurred.
 
@@
 In the situations we consider, we know that a change has occurred, and we want to forecast the future after the change.
-Change point detection is about identifying when a change has occurred.
+Change point detection is about identifying *when* a change has occurred.
 
-The changes in the series as a result of the disruption may be relatively simple, for example a left shift at the start of the disruption, and another at the end of the disruption.
+The changes in the series as a result of the disruption may be relatively simple; for example a left shift at the start of the disruption, and another at the end of the disruption.
 Or they may be more complex, with changes to the seasonal patterns, and changes to the level, which evolve over time.
@@
 
-Time series models usually assume that the future is similar to the past, at least in how the data evolve.
-But a big event like covid makes the future different from the past, at least in the short term.
+Time series models usually assume that the data evolve in the future in a similar way to how they have evolved in the past.
+But a big event (such as COVID-19) can result in a future that is different from the past, at least in the short term.
 

### 2023-07-28 fits.qmd
Cleaned up some of tourism example

@@
 #| fig-pos: "t"
-#| fig-cap: Solution 1. This should show that ETS is highly adaptive. However, for some reason, that is no longer the case, and I don't know why.
+#| fig-cap: Solution 1. Forecasts from ETS and ARIMA models. Neither works particularly well for disruptions of this magnitude.
 tar_read(tsol1_plot)
@@
 It is possible to make ETS more adaptive to changes in the data by increasing the value of the smoothing parameters.
-For example, a high value of $\beta$ (the smoothing parameter for the slope), will result in changes of trend being incorporated into the forecasts more quickly, at the risk of over-reacting to noise in the data.
+For example, a high value of $\beta$ (the smoothing parameter for the slope), will result in changes of trend being incorporated into the forecasts more quickly, at the risk of over-reacting to noise in the data, and increasing the size of the prediction intervals even more.
 

### 2023-07-28 fits.tex
Cleaned up some of tourism example

@@
 Department of Econometrics \& Business Statistics\newline
+Monash University\newline
 Clayton VIC 3800\newline
@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
 \makeatother
@@
 Many forecasters have faced this issue recently with the COVID-19
-pandemic, where historical patterns may have been severely disrupted due
-to lockdowns and other restrictions. In this paper, we consider the
-problem of forecasting after such an event has occurred.
+pandemic, where historical patterns have been severely disrupted due to
+lockdowns and other restrictions. In this paper, we consider the problem
+of forecasting during such an event, and after it has occurred.
 
@@
 to forecast the future after the change. Change point detection is about
-identifying when a change has occurred.
+identifying \emph{when} a change has occurred.
 
 The changes in the series as a result of the disruption may be
-relatively simple, for example a left shift at the start of the
+relatively simple; for example a left shift at the start of the
 disruption, and another at the end of the disruption. Or they may be
@@
 
-Time series models usually assume that the future is similar to the
-past, at least in how the data evolve. But a big event like covid makes
-the future different from the past, at least in the short term.
+Time series models usually assume that the data evolve in the future in
+a similar way to how they have evolved in the past. But a big event
+(such as COVID-19) can result in a future that is different from the
+past, at least in the short term.
 
@@
 observations never reach exactly zero, with the smallest number of
-visitors per month equal to 2250 in April 2020.
+visitors per month equal to 2250 in April 2020. For all forecasts, we
+also show 90\% prediction intervals.
 
@@
 
-\caption{\label{fig-tsol1_plot}Solution 1. This should show that ETS is
-highly adaptive. However, for some reason, that is no longer the case,
-and I don't know why.}
+\caption{\label{fig-tsol1_plot}Solution 1. Forecasts from ETS and ARIMA
+models. Neither works particularly well for disruptions of this
+magnitude.}
 
@@
 changes of trend being incorporated into the forecasts more quickly, at
-the risk of over-reacting to noise in the data.
+the risk of over-reacting to noise in the data, and increasing the size
+of the prediction intervals even more.
 
@@
 from March 2020 to October 2022 with the average of the same month in
-the 3 preceding years.}
+the three years prior to March 2020.}
 
@@
 and October 2022 with estimates based on the average of the same month
-in the three preceding years. The resulting forecasts are similar to
-those from solution 3, but with narrower prediction intervals because
-the model is (falsely) assuming that the ``observations'' during the
-disruption are real.
+in the three years prior to March 2020. The resulting forecasts are
+similar to those from solution 3, but with narrower prediction intervals
+because the model is (falsely) assuming that the ``observations'' during
+the disruption are real.
 

### 2023-07-28 _targets.R

@@ -14,77 +14,86 @@ tar_source()
 
 # List of targets
 list(
-  # Tourism example
-  tar_target(tourism_csv, here::here("data/340101.xlsx"), format="file"),
+  # Tourism data ---------------------------------------------------------------
+  tar_target(tourism_csv, here::here("data/340101.xlsx"), format = "file"),
   tar_target(austa, read_tourism(tourism_csv)),
-  # Fit ETS model
-  tar_target(fit_tourism,  
-    austa |>
-      filter(Month < yearmonth("2020 Jan")) |>
-      model(ETS(Visitors))),
-  tar_target(tourism_plot1, tourism_plot(austa) + tourism_labels),
+  # Tourism time plot
   tar_target(tourism_labels, labs(
     x = "Month",
     y = "Thousands of visitors",
     title = "Total short-term visitors to Australia"
   )),
-  tar_target(tourism_history, 
-             geom_line(data = austa |> filter(Month >= yearmonth("2016 Jan")), 
-                       aes(y=Visitors))
+  tar_target(tourism_plot1, tourism_plot(austa) + tourism_labels),
+  # Forecasts
+  tar_target(tsol1, tourism_sol1(austa)),
+  tar_target(tsol2, tourism_sol2(austa)),
+  tar_target(tsol3, tourism_sol3(austa)),
+  tar_target(tsol4, tourism_sol4(austa)),
+  # Forecast plots
+  tar_target(
+    tourism_history,
+    geom_line(
+      data = austa |> filter(Month >= yearmonth("2016 Jan")),
+      aes(y = Visitors)
+    )
   ),
-  # ETS forecasts
-  tar_target(fc_tourism, forecast(fit_tourism, h = 48)),
-  tar_target(tsol1, tourism_sol1(austa)),                            
-  tar_target(tsol2, tourism_sol2(austa)),                            
-  tar_target(tsol3, tourism_sol3(austa)),                            
-  tar_target(tsol4, tourism_sol4(austa)),   
-  tar_target(tsol1_plot, 
-    autoplot(tsol1, level=90) + tourism_history + tourism_labels
+  tar_target(
+    tsol1_plot,
+    autoplot(tsol1, level = 90) + tourism_history + tourism_labels
   ),
-  tar_target(tsol2_plot, 
-    autoplot(tsol2, level=90) +  tourism_history + tourism_labels
+  tar_target(
+    tsol2_plot,
[…]

### 2023-07-29 fits.qmd
add over sampling and ensemble

@@
 
+## Over sampling (weighting)
+
+Oversampling is a technique used in machine learning to address class imbalance within a dataset.
+In this context, the goal is to ensure that the model is more influenced by the patterns observed before the disruption period while still accounting for the observations during the disruption period.
+
+Using this weighting approach, we can leverage the insights from the pre-disruptive period while still accounting for the disruptive period's influence.
+It allows the model to better capture the underlying patterns of the time series data, considering both periods, and potentially producing more accurate forecasts.
+
+Advantages:
+
+Preserves pre-event patterns: By assigning higher weights to the pre-disruption time series data, the model is encouraged to focus more on the patterns observed before the disruptive event.
+This can be beneficial in capturing the underlying dynamics of the time series that might be essential for accurate forecasting in the post-disruption era.
+
+Robustness to disruption impact: The weighting approach allows the model to still take the disruption period into account, even if it's considered less influential.
+This ensures that the model is not completely blind to the disruption's effects and can adapt to changes or shifts in the time series caused by the event.
+
+Easy to Implement: ??
+
+Interpretability?:
+
+Disadvantages:
+
+## Ensemble models
+
 # Examples

### 2024-01-02 fits.qmd
Fixed figure labels with underscores

@@
 The Australian tourism data set is a monthly time series showing the number of short-term overseas visitors to Australia.
-The data [@tourismdata] are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1.
+The data [@tourismdata] are available from January 2000 to May 2023, and are shown in @fig-tourism-plot1.
 As the borders closed in March 2020, the number of visitors to Australia dropped to near zero, and remained there until towards the end of 2021.
@@
 ```{r}
-#| label: fig-tourism_plot1
+#| label: fig-tourism-plot1
 #| fig-cap: "Short-term visitor arrivals to Australia (monthly): Jan 2000 -- May 2023."
@@
 ```{r}
-#| label: fig-tsol1_plot
+#| label: fig-tsol1-plot
 #| fig-height: 8
 #| fig-pos: "t"
-#| fig-cap: Solution 1. Forecasts from ETS and ARIMA models. Neither works particularly well for disruptions of this magnitude.
+#| fig-cap: "Solution 1. Forecasts from ETS and ARIMA models. Neither works particularly well for disruptions of this magnitude."
 tar_read(tsol1_plot)
@@
 
-In @fig-tsol1_plot, we show forecasts by applying ETS and ARIMA models to the data.
+In @fig-tsol1-plot, we show forecasts by applying ETS and ARIMA models to the data.
 ETS, in particular, is well-known to be relatively adaptive to changes in the series, and this is evident in these forecasts.
@@
 ```{r}
-#| label: fig-tsol2_plot
+#| label: fig-tsol2-plot
 #| fig-height: 8
@@
 
-@fig-tsol2_plot shows forecasts obtained using a dynamic regression model (i.e., a regression with ARIMA errors) using two intervention variables: a level shift from March 2022 to November 2022, and a ramp from October 2021 to November 2022.
+@fig-tsol2-plot shows forecasts obtained using a dynamic regression model (i.e., a regression with ARIMA errors) using two intervention variables: a level shift from March 2022 to November 2022, and a ramp from October 2021 to November 2022.
 It is evident that the level shift variable has worked well, giving relatively good forecasts for 2021.
@@
 ```{r}
-#| label: fig-tsol3_plot
+#| label: fig-tsol3-plot
 #| fig-height: 8
@@
 ```{r}
-#| label: fig-tsol4_plot
+#| label: fig-tsol4-plot
 #| fig-height: 8
@@
 
-Finally, we show in @fig-tsol4_plot the forecasts obtained using solution 4.
+Finally, we show in @fig-tsol4-plot the forecasts obtained using solution 4.
 Here we have replaced observations between March 2020 and October 2022 with estimates based on the average of the same month in the three years prior to March 2020.

### 2024-01-02 fits.tex
Fixed figure labels with underscores

@@
 \makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
-\makeatother
-\makeatletter
 \makeatother
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\hypertarget{handling-interruptions-when-forecasting}{%
 \section{Handling interruptions when
-forecasting}\label{handling-interruptions-when-forecasting}}
+forecasting}\label{handling-interruptions-when-forecasting}
 
@@
 
-\hypertarget{use-a-highly-adaptive-model}{%
 \subsection{Use a highly adaptive
-model}\label{use-a-highly-adaptive-model}}
+model}\label{use-a-highly-adaptive-model}
 
@@
 
-\hypertarget{use-a-dynamic-regression-model-with-intervention-covariates}{%
 \subsection{Use a dynamic regression model with intervention
-covariates}\label{use-a-dynamic-regression-model-with-intervention-covariates}}
+covariates}\label{use-a-dynamic-regression-model-with-intervention-covariates}
 
@@
 
-\hypertarget{treat-the-covid-period-as-missing-and-use-a-model-that-handles-missing-values}{%
 \subsection{Treat the covid period as missing and use a model that
 handles missing
-values}\label{treat-the-covid-period-as-missing-and-use-a-model-that-handles-missing-values}}
+values}\label{treat-the-covid-period-as-missing-and-use-a-model-that-handles-missing-values}
 
@@
 
-\hypertarget{estimate-what-might-have-been-and-adjust-the-data}{%
 \subsection{Estimate what might have been and adjust the
-data}\label{estimate-what-might-have-been-and-adjust-the-data}}
+data}\label{estimate-what-might-have-been-and-adjust-the-data}
 
@@
 
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
+\subsection{Over sampling (weighting)}\label{over-sampling-weighting}
+
+Oversampling is a technique used in machine learning to address class
+imbalance within a dataset. In this context, the goal is to ensure that
+the model is more influenced by the patterns observed before the
+disruption period while still accounting for the observations during the
+disruption period.
+
+Using this weighting approach, we can leverage the insights from the
+pre-disruptive period while still accounting for the disruptive period's
+influence. It allows the model to better capture the underlying patterns
+of the time series data, considering both periods, and potentially
+producing more accurate forecasts.
+
+Advantages:
+
+Preserves pre-event patterns: By assigning higher weights to the
+pre-disruption time series data, the model is encouraged to focus more
+on the patterns observed before the disruptive event. This can be
+beneficial in capturing the underlying dynamics of the time series that
+might be essential for accurate forecasting in the post-disruption era.
+
+Robustness to disruption impact: The weighting approach allows the model
+to still take the disruption period into account, even if it's
+considered less influential. This ensures that the model is not
+completely blind to the disruption's effects and can adapt to changes or
+shifts in the time series caused by the event.
+
+Easy to Implement: ??
+
+Interpretability?:
+
+Disadvantages:
+
+\subsection{Ensemble models}\label{ensemble-models}
+
+\section{Examples}\label{examples}
 
@@
 
-\hypertarget{australian-tourism}{%
-\subsection{Australian tourism}\label{australian-tourism}}
+\subsection{Australian tourism}\label{australian-tourism}
 
@@
 \autocite{tourismdata} are available from January 2000 to May 2023, and
-are shown in Figure~\ref{fig-tourism_plot1}. As the borders closed in
+are shown in Figure~\ref{fig-tourism-plot1}. As the borders closed in
 March 2020, the number of visitors to Australia dropped to near zero,
@@
 
-{\centering \includegraphics{fits_files/figure-pdf/fig-tourism_plot1-1.pdf}
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-tourism-plot1-1.pdf}
 
@@
 
-\caption{\label{fig-tourism_plot1}Short-term visitor arrivals to
+\caption{\label{fig-tourism-plot1}Short-term visitor arrivals to
 Australia (monthly): Jan 2000 -- May 2023.}
 
-\end{figure}
+\end{figure}%
 
@@
 
-{\centering \includegraphics{fits_files/figure-pdf/fig-tsol1_plot-1.pdf}
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-tsol1-plot-1.pdf}
 
@@
 
-\caption{\label{fig-tsol1_plot}Solution 1. Forecasts from ETS and ARIMA
+\caption{\label{fig-tsol1-plot}Solution 1. Forecasts from ETS and ARIMA
 models. Neither works particularly well for disruptions of this
@@
 
-\end{figure}
+\end{figure}%
 
-In Figure~\ref{fig-tsol1_plot}, we show forecasts by applying ETS and
+In Figure~\ref{fig-tsol1-plot}, we show forecasts by applying ETS and
 ARIMA models to the data. ETS, in particular, is well-known to be
@@
 
-{\centering \includegraphics{fits_files/figure-pdf/fig-tsol2_plot-1.pdf}
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-tsol2-plot-1.pdf}
 
@@
 
-\caption{\label{fig-tsol2_plot}Solution 2. We use a left shift from
+\caption{\label{fig-tsol2-plot}Solution 2. We use a left shift from
 March 2020 to October 2022, and a ramp from October 2021 to October
@@
 
-\end{figure}
+\end{figure}%
 
-Figure~\ref{fig-tsol2_plot} shows forecasts obtained using a dynamic
+Figure~\ref{fig-tsol2-plot} shows forecasts obtained using a dynamic
 regression model (i.e., a regression with ARIMA errors) using two
@@
 
-{\centering \includegraphics{fits_files/figure-pdf/fig-tsol3_plot-1.pdf}
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-tsol3-plot-1.pdf}
 
@@
 
-\caption{\label{fig-tsol3_plot}Solution 3. The period from March 2020 to
+\caption{\label{fig-tsol3-plot}Solution 3. The period from March 2020 to
 October 2022 is set to missing. So the first three years of forecasts
@@
 
-\end{figure}
+\end{figure}%
 
@@
 
-{\centering \includegraphics{fits_files/figure-pdf/fig-tsol4_plot-1.pdf}
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-tsol4-plot-1.pdf}
 
@@
 
-\caption{\label{fig-tsol4_plot}Solution 4. We replace the observations
+\caption{\label{fig-tsol4-plot}Solution 4. We replace the observations
 from March 2020 to October 2022 with the average of the same month in
@@
 
-\end{figure}
+\end{figure}%
 
-Finally, we show in Figure~\ref{fig-tsol4_plot} the forecasts obtained
+Finally, we show in Figure~\ref{fig-tsol4-plot} the forecasts obtained
 using solution 4. Here we have replaced observations between March 2020
@@
 
-\hypertarget{example-ambulance-attendances}{%
 \subsection{Example: Ambulance
-attendances}\label{example-ambulance-attendances}}
+attendances}\label{example-ambulance-attendances}
 

### 2024-01-04 pedestrians.R
Started daily example

@@ -1,0 +1,183 @@
+library(fpp3)
+library(tidyverse)
+library(rwalkr)
+
+init <- 21
+h <- 21
+
+fetch_walkers <- function(start = "2017-01-01", end = "2023-12-31") {
+  # Download hourly data from start to end
+  walkers <- melb_walk(from = as.Date(start), to = as.Date(end)) |> 
+    mutate(Count = parse_integer(Count))
+  # Compute daily totals by sensor
+  daily_walkers <- walkers |>
+    group_by(Date, Sensor) |>
+    summarise(Count = sum(Count, na.rm = TRUE)) |>
+    ungroup() |>
+    mutate(Count = if_else(Count == 0L, NA_integer_, Count)) |>
+    as_tsibble(index = Date, key = Sensor)
+
+  # Remove censors with large numbers of missing values
+  nmiss <- daily_walkers |>
+    as_tibble() |>
+    group_by(Sensor) |>
+    summarise(nmiss = sum(is.na(Count))) |>
+    filter(nmiss < 10)
+
+  # Interpolate missing values
+  daily_walkers <- daily_walkers |>
+    filter(Sensor %in% nmiss$Sensor) |>
+    mutate(Count = zoo::na.approx(Count))
+
+  # Average across sensors
+  average_daily_walkers <- daily_walkers |>
+    as_tibble() |>
+    group_by(Date) |>
+    summarise(Count = mean(Count) / 1e3) |>
+    tsibble::as_tsibble(index = Date)
+
+  return(average_daily_walkers)
+}
+
+average_daily_walkers <- fetch_walkers(end = "2017-01-30")
+
+average_daily_walkers |>
+  autoplot(Count) +
+  labs(
+    title = "Daily average number of pedestrians across Melbourne",
+    y = "Thousands of persons"
+  )
+
+# Solution 1
+
+fc1 <- average_daily_walkers |>
+  stretch_tsibble(.step = 7, .init = 21) |>
+  model(
+    ets = ETS(Count),
+    arima = ARIMA(sqrt(Count) ~ fourier(period = "year", K = 4) + PDQ(period = "week"))
+  ) |>
+  forecast(h = 14) |>
+  mutate(.id = paste("Forecasts of", .id + 2019))
+
+# Solution 2
+
+ped_sol2_fc <- function(fit, h = h) {
+  training_data <- fit$arima[[1]]$data
+  id <- fit$.id[1]
[…]

### 2024-01-05 fits.qmd
Updated pedestrians example

@@
 
-# Handling interruptions when forecasting
+# Handling interruptions when forecasting {#sec-methods}
 
@@
 
+## Example: Pedestrians
+
+```{r}
+#| include: false
+tar_load(c(raw_walkers, walkers2, lockdowns))
+nsensors <- length(unique(raw_walkers$Sensor))
+ngoodsensors <- length(unique(walkers2$Sensor))
+lockdowns <- lockdowns |>
+  mutate(Duration = End - Start + 1)
+```
+
+The pedestrian data set is a daily time series showing the number of pedestrians per day in Melbourne, Australia.
+The data were obtained from the Pedestrian Counting System maintained by the City of Melbourne [@pedestrians], and are based on automated hourly counts from sensors at `r nsensors` locations over the period from `r min(raw_walkers$Date)` to `r max(raw_walkers$Date)`.
+All sensors had some missing data, and we chose to use sensors with no more than `r 30*24` missing hours, equivalent to 30 days of missing data.
+This resulted in `r ngoodsensors` sensors being used in the analysis.
+The data were aggregated to daily counts for each sensor, and then the results were averaged across sensors to obtain a measure of pedestrian traffic per day.
+The resulting time series is shown in @fig-walkers.
+
+```{r}
+#| label: fig-walkers
+#| fig-height: 3.7
+#| fig-cap: "Pedestrian traffic in Melbourne, Australia, from 2018 to 2023. The gray shaded regions indicate periods of lockdown due to COVID-19."
+tar_read(walkers_plot)
+```
+
+We have also shown the periods of lockdown due to COVID-19.
+In Melbourne, there were six separate lockdowns ranging from `r as.numeric(min(lockdowns$Duration))` days to `r as.numeric(max(lockdowns$Duration))` days in length.
+The first official lockdown period was from 31 March 2020 to `r format(min(lockdowns$End), "%d %b %Y")`, and the last lockdown was from `r format(max(lockdowns$Start), "%d %b %Y")` to `r format(max(lockdowns$End), "%d %b %Y")`.
+However, the first period was preceded by over a week when most people elected to stay and work at home, so we have chosen to start the first period on `r format(min(lockdowns$Start), "%d %b %Y")`, to better reflect human behaviour.
+Otherwise, we have used the official lockdown dates.
+
+We apply the four solutions discussed in @sec-methods.
+We fit ETS, ARIMA and dynamic ARIMA models to the data, with no transformations used.
+We have evaluated the methods by using time series cross-validation, with the initial training set comprising the whole of 2019, and subsequent training sets growing by 1 week at a time, to the end of 2021.
+The test sets are always 3 weeks long.
+Thus, we have evaluated the forecasts during 2020 and 2021, covering all lockdown periods, and the start of the recovery period.
+The results are shown in @fig-walkers-plot1.
+
 ## Example: Ambulance attendances
@@
 ```
+
+# References

### 2024-01-05 R/pedestrians.R

@@ -1,0 +1,173 @@
+# Read data from API
+fetch_walkers <- function(start = "2019-01-01", end = "2023-12-31") {
+  # Download hourly data from start to end
+  walkers <- rwalkr::melb_walk(from = as.Date(start), to = as.Date(end)) 
+}
+
+clean_walkers <- function(walkers) {
+  # Remove first values of each day which appear to be sensor identifiers
+  walkers <- walkers |> 
+    arrange(Sensor, Date, Time) |> 
+    group_by(Sensor, Date) |> 
+    mutate(obs = row_number()-1) |> 
+    filter(obs > 0) |> 
+    select(-obs) |> 
+    mutate(Count = readr::parse_integer(Count))
+
+  # Find sensors that are used throughout the whole period
+  sensors_to_use <- walkers |> 
+    as_tsibble(index = Date_Time, key = Sensor) |> 
+    fill_gaps(.full = TRUE) |> 
+    as_tibble() |> 
+    group_by(Sensor) |> 
+    summarise(nmiss = sum(is.na(Count))) |> 
+    filter(nmiss < 30*24) |> 
+    pull(Sensor) 
+  
+  # Compute daily totals by sensor
+  walkers |>
+    filter(Sensor %in% sensors_to_use) |> 
+    group_by(Date, Sensor) |>
+    summarise(Count = sum(Count), .groups = "drop") |>
+    as_tsibble(index = Date, key = Sensor)
+}
+
+ave_walkers <- function(walkers) {
+  # Average across sensors
+  walkers |>
+    as_tibble() |>
+    group_by(Date) |>
+    summarise(Count = mean(Count, na.rm=TRUE) / 1e3) |>
+    tsibble::as_tsibble(index = Date)
+}
+
+lockdown_table <- function() {
+  # First lockdown officially started 31 March, but self-lockdown happened from 
+  # start of previous week. Other dates below are official periods as per
+  # https://en.wikipedia.org/wiki/COVID-19_pandemic_in_Victoria#Lockdown_statistics
+  # See also https://www.platinumaccounting.com.au/blog/melbourne-lockdown-dates
+  tribble(
+    ~Lockdown, ~Start, ~End,
+    1, "2020-03-23", "2020-05-12",
+    2, "2020-07-09", "2020-10-27",
+    3, "2021-02-13", "2021-02-17",
+    4, "2021-05-28", "2021-06-10",
[…]

### 2024-01-05 _targets.R

@@ -54,6 +54,21 @@ list(
     autoplot(tsol4, level = 90) + tourism_history + tourism_labels
   ),
 
+  # Pedestrians
+  tar_target(raw_walkers, fetch_walkers(start = "2019-01-01", end = "2021-12-31")),
+  tar_target(lockdowns, lockdown_table()),
+  tar_target(walkers2, clean_walkers(raw_walkers)),
+  tar_target(walkers3, ave_walkers(walkers2)),
+  tar_target(walkers, add_lockdowns(walkers3)),
+  tar_target(walkers_plot, pedestrian_plot(walkers)),
+  tar_target(init, 365),
+  tar_target(step, 7),
+  tar_target(h, 21),
+  tar_target(walkers_sol1, pedestrian_sol1(walkers, step = step, init = init, h = h)),
+  tar_target(walkers_sol2, pedestrian_sol2(walkers, step = step, init = init, h = h)),
+  tar_target(walkers_sol3, pedestrian_sol3(walkers, step = step, init = init, h = h)),
+  tar_target(walkers_sol4, pedestrian_sol4(walkers, step = step, init = init, h = h)),
+
   # Attended incidents example
   tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format = "file"),
   tar_target(verified_csv, here::here("data/verified_incident.csv"), format = "file"),

### 2024-01-05 pedestrians.R

@@ -1,183 +1,0 @@
-library(fpp3)
-library(tidyverse)
-library(rwalkr)
-
-init <- 21
-h <- 21
-
-fetch_walkers <- function(start = "2017-01-01", end = "2023-12-31") {
-  # Download hourly data from start to end
-  walkers <- melb_walk(from = as.Date(start), to = as.Date(end)) |> 
-    mutate(Count = parse_integer(Count))
-  # Compute daily totals by sensor
-  daily_walkers <- walkers |>
-    group_by(Date, Sensor) |>
-    summarise(Count = sum(Count, na.rm = TRUE)) |>
-    ungroup() |>
-    mutate(Count = if_else(Count == 0L, NA_integer_, Count)) |>
-    as_tsibble(index = Date, key = Sensor)
-
-  # Remove censors with large numbers of missing values
-  nmiss <- daily_walkers |>
-    as_tibble() |>
-    group_by(Sensor) |>
-    summarise(nmiss = sum(is.na(Count))) |>
-    filter(nmiss < 10)
-
-  # Interpolate missing values
-  daily_walkers <- daily_walkers |>
-    filter(Sensor %in% nmiss$Sensor) |>
-    mutate(Count = zoo::na.approx(Count))
-
-  # Average across sensors
-  average_daily_walkers <- daily_walkers |>
-    as_tibble() |>
-    group_by(Date) |>
-    summarise(Count = mean(Count) / 1e3) |>
-    tsibble::as_tsibble(index = Date)
-
-  return(average_daily_walkers)
-}
-
-average_daily_walkers <- fetch_walkers(end = "2017-01-30")
-
-average_daily_walkers |>
-  autoplot(Count) +
-  labs(
-    title = "Daily average number of pedestrians across Melbourne",
-    y = "Thousands of persons"
-  )
-
-# Solution 1
-
-fc1 <- average_daily_walkers |>
-  stretch_tsibble(.step = 7, .init = 21) |>
-  model(
-    ets = ETS(Count),
-    arima = ARIMA(sqrt(Count) ~ fourier(period = "year", K = 4) + PDQ(period = "week"))
-  ) |>
-  forecast(h = 14) |>
-  mutate(.id = paste("Forecasts of", .id + 2019))
-
-# Solution 2
-
-ped_sol2_fc <- function(fit, h = h) {
-  training_data <- fit$arima[[1]]$data
-  id <- fit$.id[1]
[…]

### 2024-01-05 fits.tex
More work on pedestrians

@@
 
-\section{Handling interruptions when
-forecasting}\label{handling-interruptions-when-forecasting}
+\section{Handling interruptions when forecasting}\label{sec-methods}
 
@@
 
+\subsection{Example: Pedestrians}\label{example-pedestrians}
+
+The pedestrian data set is a daily time series showing the number of
+pedestrians per day in Melbourne, Australia. The data were obtained from
+the Pedestrian Counting System maintained by the City of Melbourne
+\autocite{pedestrians}, and are based on automated hourly counts from
+sensors at 66 locations over the period from 2019-01-01 to 2021-12-31.
+All sensors had some missing data, and we chose to use sensors with no
+more than 720 missing hours, equivalent to 30 days of missing data. This
+resulted in 25 sensors being used in the analysis. The data were
+aggregated to daily counts for each sensor, and then the results were
+averaged across sensors to obtain a measure of pedestrian traffic per
+day. The resulting time series is shown in Figure~\ref{fig-walkers}.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-1.pdf}
+
+}
+
+\caption{\label{fig-walkers}Pedestrian traffic in Melbourne, Australia,
+from 2018 to 2023. The gray shaded regions indicate periods of lockdown
+due to COVID-19.}
+
+\end{figure}%
+
+We have also shown the periods of lockdown due to COVID-19. In
+Melbourne, there were six separate lockdowns ranging from 5 days to 111
+days in length. The first official lockdown period was from 31 March
+2020 to 12 May 2020, and the last lockdown was from 05 Aug 2021 to 21
+Oct 2021. However, the first period was preceded by over a week when
+most people elected to stay and work at home, so we have chosen to start
+the first period on 23 Mar 2020, to better reflect human behaviour.
+Otherwise, we have used the official lockdown dates.
+
+We apply the four solutions discussed in Section~\ref{sec-methods}. We
+fit ETS, ARIMA and dynamic ARIMA models to the data, with no
+transformations used. We have evaluated the methods by using time series
+cross-validation, with the initial training set comprising the whole of
+2019, and subsequent training sets growing by 1 week at a time, to the
+end of 2021. The test sets are always 3 weeks long. Thus, we have
+evaluated the forecasts during 2020 and 2021, covering all lockdown
+periods, and the start of the recovery period. The results are shown in
+\textbf{?@fig-walkers-plot1}.
+
 \subsection{Example: Ambulance
@@
 
-\includegraphics{fits_files/figure-pdf/unnamed-chunk-8-1.pdf}
+\includegraphics{fits_files/figure-pdf/unnamed-chunk-10-1.pdf}
 
@@
 
-\printbibliography
+\printbibliography[title=References]
 

### 2024-01-05 R/pedestrians.R

@@ -16,6 +16,8 @@ clean_walkers <- function(walkers) {
 
   # Find sensors that are used throughout the whole period
   sensors_to_use <- walkers |> 
+    group_by(Sensor, Date_Time) |>
+    summarise(Count = sum(Count), .groups = "drop") |>
     as_tsibble(index = Date_Time, key = Sensor) |> 
     fill_gaps(.full = TRUE) |> 
     as_tibble() |> 
@@ -95,10 +97,7 @@ pedestrian_plot <- function(data) {
 pedestrian_sol1 <- function(walkers, step, init, h) {
   walkers |>
     stretch_tsibble(.step = step, .init = init) |>
-    model(
-      ets = ETS(Count),
-      arima = ARIMA(Count ~ PDQ(period = "week"))
-    ) |>
+    model(arima = ARIMA(Count ~ PDQ(period = "week"))) |>
     forecast(h = h) 
 }
 
@@ -165,9 +164,6 @@ pedestrian_sol4 <- function(walkers, step, init, h) {
     interpolate(new_data = tmp) 
   new_walkers |>
     stretch_tsibble(.step = step, .init = init) |>
-    model(
-      ets = ETS(Count),
-      arima = ARIMA(Count ~ PDQ(period = "week"))
-    ) |>
+    model(arima = ARIMA(Count ~ PDQ(period = "week"))) |>
     forecast(h = h) 
 }

### 2024-01-06 R/pedestrians.R
Bug fix

@@ -66,8 +66,8 @@ add_lockdowns <- function(walkers) {
   lockdowns <- lockdown_table()  
   walkers <- walkers |>
     mutate(Lockdown_period = 0)
-  for(i in seq(nrow(lockdowns))) {
-    awalkers <- walkers |>
+  for (i in seq(nrow(lockdowns))) {
+    walkers <- walkers |>
       mutate(
         Lockdown_period = if_else(Date >= lockdowns$Start[i] & Date <= lockdowns$End[i], i, Lockdown_period)
       )

### 2024-01-06 R/pedestrians.R
Code styling

@@ -1,34 +1,34 @@
 # Read data from API
 fetch_walkers <- function(start = "2019-01-01", end = "2023-12-31") {
   # Download hourly data from start to end
-  walkers <- rwalkr::melb_walk(from = as.Date(start), to = as.Date(end)) 
+  walkers <- rwalkr::melb_walk(from = as.Date(start), to = as.Date(end))
 }
 
 clean_walkers <- function(walkers) {
   # Remove first values of each day which appear to be sensor identifiers
-  walkers <- walkers |> 
-    arrange(Sensor, Date, Time) |> 
-    group_by(Sensor, Date) |> 
-    mutate(obs = row_number()-1) |> 
-    filter(obs > 0) |> 
-    select(-obs) |> 
+  walkers <- walkers |>
+    arrange(Sensor, Date, Time) |>
+    group_by(Sensor, Date) |>
+    mutate(obs = row_number() - 1) |>
+    filter(obs > 0) |>
+    select(-obs) |>
     mutate(Count = readr::parse_integer(Count))
 
   # Find sensors that are used throughout the whole period
-  sensors_to_use <- walkers |> 
+  sensors_to_use <- walkers |>
     group_by(Sensor, Date_Time) |>
     summarise(Count = sum(Count), .groups = "drop") |>
-    as_tsibble(index = Date_Time, key = Sensor) |> 
-    fill_gaps(.full = TRUE) |> 
-    as_tibble() |> 
-    group_by(Sensor) |> 
-    summarise(nmiss = sum(is.na(Count))) |> 
-    filter(nmiss < 30*24) |> 
-    pull(Sensor) 
-  
+    as_tsibble(index = Date_Time, key = Sensor) |>
+    fill_gaps(.full = TRUE) |>
+    as_tibble() |>
+    group_by(Sensor) |>
+    summarise(nmiss = sum(is.na(Count))) |>
+    filter(nmiss < 30 * 24) |>
+    pull(Sensor)
+
   # Compute daily totals by sensor
   walkers |>
-    filter(Sensor %in% sensors_to_use) |> 
+    filter(Sensor %in% sensors_to_use) |>
     group_by(Date, Sensor) |>
     summarise(Count = sum(Count), .groups = "drop") |>
     as_tsibble(index = Date, key = Sensor)
@@ -39,12 +39,12 @@ ave_walkers <- function(walkers) {
   walkers |>
     as_tibble() |>
     group_by(Date) |>
[…]

### 2024-01-06 fits.qmd
Added pedestrian plot

@@
 We apply the four solutions discussed in @sec-methods.
-We fit ETS, ARIMA and dynamic ARIMA models to the data, with no transformations used.
+We fit ARIMA and dynamic ARIMA models to the data, with no transformations used.
 We have evaluated the methods by using time series cross-validation, with the initial training set comprising the whole of 2019, and subsequent training sets growing by 1 week at a time, to the end of 2021.
@@
 Thus, we have evaluated the forecasts during 2020 and 2021, covering all lockdown periods, and the start of the recovery period.
-The results are shown in @fig-walkers-plot1.
+The results are shown in Figures [-@fig-walkers-plot1]--[-@fit-walkers-plot4]
+
+```{r}
+#| label: fig-walkers-plot1
+#| fig-height: 3
+#| fig-cap: "Solution 1: the forecasts do not take into account the lockdown periods, but the ARIMA model is able to adjust to the changing level of the series."
+tar_read(walkers_plot1)
+```
+
+```{r}
+#| label: fig-walkers-plot2
+#| fig-height: 3
+#| fig-cap: "Solution 2: the forecasts are based on an ARIMA model with a dummy variable indicating when the lockdown periods occurred."
+tar_read(walkers_plot2)
+```
+
+```{r}
+#| label: fig-walkers-plot3
+#| fig-height: 3
+#| fig-cap: "Solution 3: the lockdown periods are set to missing, and the ARIMA model uses the remaining data to produce forecasts"
+tar_read(walkers_plot3)
+```
+
+```{r}
+#| label: fig-walkers-plot4
+#| fig-height: 3
+#| fig-cap: "Solution 4: the lockdown periods are replaced with estimated counts based on an ARIMA model applied to the remaining data.
+Then a new ARIMA model is fitted to the whole data set, and used to produce forecasts."
+tar_read(walkers_plot4)
+```
 

### 2024-01-06 R/pedestrians.R

@@ -172,4 +172,17 @@ pedestrian_sol4 <- function(walkers, step, init, h) {
     model(arima = ARIMA(Count ~ PDQ(period = "week"))) |>
     forecast(h = h)
 }
+
+pedestrian_fc_plot <- function(fc, data) {
+  fc |>
+    group_by(.id) |>
+    slice(1:7) |>
+    mutate(.id = as.character(.id)) |>
+    ggplot(aes(x = Date)) +
+    geom_line(aes(y = .mean, group = .id), col = "blue") +
+    geom_line(
+      data = data |> filter(Date >= "2020-01-01"),
+      mapping = aes(y = Count)
+    ) +
+    guides(col = "none")
 }

### 2024-01-06 _targets.R

@@ -68,7 +68,11 @@ list(
   tar_target(walkers_sol2, pedestrian_sol2(walkers, step = step, init = init, h = h)),
   tar_target(walkers_sol3, pedestrian_sol3(walkers, step = step, init = init, h = h)),
   tar_target(walkers_sol4, pedestrian_sol4(walkers, step = step, init = init, h = h)),
-
+  tar_target(walkers_plot1, pedestrian_fc_plot(walkers_sol1, walkers)),
+  tar_target(walkers_plot2, pedestrian_fc_plot(walkers_sol2, walkers)),
+  tar_target(walkers_plot3, pedestrian_fc_plot(walkers_sol3, walkers)),
+  tar_target(walkers_plot4, pedestrian_fc_plot(walkers_sol4, walkers)),
+  
   # Attended incidents example
   tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format = "file"),
   tar_target(verified_csv, here::here("data/verified_incident.csv"), format = "file"),

### 2024-01-06 fits.tex
Recompiled

@@
 We apply the four solutions discussed in Section~\ref{sec-methods}. We
-fit ETS, ARIMA and dynamic ARIMA models to the data, with no
-transformations used. We have evaluated the methods by using time series
+fit ARIMA and dynamic ARIMA models to the data, with no transformations
+used. We have evaluated the methods by using time series
 cross-validation, with the initial training set comprising the whole of
 2019, and subsequent training sets growing by 1 week at a time, to the
-end of 2021. The test sets are always 3 weeks long. Thus, we have
+end of 2021. The test sets are always 1 week long. Thus, we have
 evaluated the forecasts during 2020 and 2021, covering all lockdown
 periods, and the start of the recovery period. The results are shown in
-\textbf{?@fig-walkers-plot1}.
+Figures \ref{fig-walkers-plot1}--\ref{fig-walkers-plot4}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot1-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot1}Solution 1: the forecasts do not take
+into account the lockdown periods, but the ARIMA model is able to adjust
+to the changing level of the series.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot2-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot2}Solution 2: the forecasts are based on
+an ARIMA model with a dummy variable indicating when the lockdown
+periods occurred.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot3-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot3}Solution 3: the lockdown periods are
+set to missing, and the ARIMA model uses the remaining data to produce
+forecasts}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot4-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot4}Solution 4: the lockdown periods are
+replaced with estimated counts based on an ARIMA model applied to the
+remaining data. Then a new ARIMA model is fitted to the whole data set,
+and used to produce forecasts.}
+
+\end{figure}%
 
@@
 
-\includegraphics{fits_files/figure-pdf/unnamed-chunk-10-1.pdf}
+\includegraphics{fits_files/figure-pdf/unnamed-chunk-14-1.pdf}
 

### 2024-01-08 fits.qmd
Recompiled
Updated methods section

@@
 
-The changes in the series as a result of the disruption may be relatively simple; for example a left shift at the start of the disruption, and another at the end of the disruption.
+Time series models usually assume that the data evolve in the future in a similar way to how they have evolved in the past.
+But a big event (such as COVID-19) can result in a future that is different from the past, at least in the short term.
+
+The disruption may result in relatively simple changes in the series; for example a level shift at the start of the disruption, and another at the end of the disruption.
 Or they may be more complex, with changes to the seasonal patterns, and changes to the level, which evolve over time.
@@
 
-Time series models usually assume that the data evolve in the future in a similar way to how they have evolved in the past.
-But a big event (such as COVID-19) can result in a future that is different from the past, at least in the short term.
+We introduce several approaches to handling interruptions when forecasting in @sec-methods.
+We then apply these approaches to two real data sets in @sec-sec-examples.
+Finally, in @sec-conclusions, we discuss some of the advantages and disadvantages of each approach, and provide some recommendations for practitioners.
 
@@
 This has the advantage of being a very simple solution that is easy to implement and fast to compute.
+There is no need to explicitly model the interruption, and so the model can be used for forecasting even if the timing or effect of the interruption is unknown.
 
@@
 
-## Use a dynamic regression model with intervention covariates
+## Use an intervention model
 
@@
 This has the advantage of retaining the memory of the past, and so the seasonal patterns and other dynamics will be retained.
-However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be be too narrow.
+This allows the change period to be effectively modelled provided the intervention variables are chosen well.
+However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be be too narrow, especially if there is a lasting effect beyond the end of the interruption.
 
-Advantages: retains full memory of the past, and allows the change period to be effectively modelled provided you choose the covariates well.
-
-Disadvantages: requires a lot of thought to choose the covariates well. Assumes that the post-pandemic period will be similar to the pre-pandemic period.
-
-## Treat the covid period as missing and use a model that handles missing values
+## Treat the interruption period as missing
 
@@
 
-## Estimate what might have been and adjust the data
+## Estimate what might have been
 
@@
 This then becomes almost equal to the previous solution except that the prediction intervals will be narrower because the estimation uncertainty has not been taken into account.
+On the other hand, it is more flexible than the previous approach because models that do not handle missing values can then be used.
 
-## Over sampling (weighting)
+An alternative would be to take the model estimated under the previous solution (setting the observations during the disruption to missing), and use it to estimate what might have been during the disruption.
+Then the model can be re-estimated using the adjusted data.
+Note that forecasts obtained in this way during the disruption will not be true forecasts, because they will have used data from the future in computing the adjusted data during the disruption.
+But post-disruption forecasts will be true forecasts, and should be almost the same as those obtained using the previous solution.
 
-Oversampling is a technique used in machine learning to address class imbalance within a dataset.
-In this context, the goal is to ensure that the model is more influenced by the patterns observed before the disruption period while still accounting for the observations during the disruption period.
+## Downweight the interruption period
+
+Intuitively, we want the model to be more influenced by the patterns observed before the disruption period than those during the disruption, but we still want to take some account of the observations during the disruption period.
 
 Using this weighting approach, we can leverage the insights from the pre-disruptive period while still accounting for the disruptive period's influence.
-It allows the model to better capture the underlying patterns of the time series data, considering both periods, and potentially producing more accurate forecasts.
+It allows the model to better capture the underlying dynamics of the time series in non-disrupted periods, which may be essential for accurate forecasting in the post-disruption era.
 
-Advantages:
+If the weights during the disruption are set to zero, and the weights at other times are set to 1, this becomes equivalent to the missing value solution above.
+Rather than ignoring the data during the disruption, the weighting approach allows the model to not be completely blind to the disruption's effects, and adapt to changes or shifts in the time series caused by the event.
 
-Preserves pre-event patterns: By assigning higher weights to the pre-disruption time series data, the model is encouraged to focus more on the patterns observed before the disruptive event.
-This can be beneficial in capturing the underlying dynamics of the time series that might be essential for accurate forecasting in the post-disruption era.
-
-Robustness to disruption impact: The weighting approach allows the model to still take the disruption period into account, even if it's considered less influential.
-This ensures that the model is not completely blind to the disruption's effects and can adapt to changes or shifts in the time series caused by the event.
-
-Easy to Implement: ??
-
-Interpretability?:
-
-Disadvantages:
+While conceptually simple to implement, in practice, this approach may require more work than the other methods discussed here as most forecasting software does not allow for the explicit use of weights.
 
@@
 
-# Examples
+As with may forecasting problems, taken an ensemble approach often leads to more accurate forecasts [@combinations].
+Here, we could combine some or all four of the approaches discussed above to obtain an ensemble forecast.
+In fact, if we were unable to implement the weighted approach, by using an ensemble of the other approaches we are effectively downweighting the observations during the disruption period as they are only explicitly used in the first two approaches (using a highly adaptive model, or using an intervention model).
 
-Let's consider some examples
+However, a disadvantage of this approach is that it is averaging forecasts that aim to achieve different objectives.
+For example, the highly adaptive model and the intervention model are trying to forecast what happened during the disruption, while the missing value and estimation approaches are trying to forecast what might have happened.
+So the ensemble approach will lie somewhere between these two objectives.
 
-## Australian tourism
+# Examples {#sec-examples}
+
+Let's consider some examples to data observed during the last few years, where the effect of the COVID-19 pandemic has been particularly evident.
+
+## Australian tourism {#sec-tourism}
 
@@
 We have evaluated the methods by using time series cross-validation, with the initial training set comprising the whole of 2019, and subsequent training sets growing by 1 week at a time, to the end of 2021.
-The test sets are always 3 weeks long.
+The test sets are always 1 week long.
 Thus, we have evaluated the forecasts during 2020 and 2021, covering all lockdown periods, and the start of the recovery period.
-The results are shown in Figures [-@fig-walkers-plot1]--[-@fit-walkers-plot4]
+The results are shown in Figures [-@fig-walkers-plot1]--[-@fig-walkers-plot4]
 

### 2024-01-08 fits.qmd
Updated tourism analysis

@@
 
-We apply the four solutions discussed in the previous section to these data, making 12 month forecasts at the end of each year from 2019 to 2022.
+We apply the first four solutions discussed in the previous section to these data, making 12 month forecasts at the end of each year from 2019 to 2022.
 We fit ETS, ARIMA and dynamic ARIMA models to the data [@fpp3], first applying a log transformation to ensure the resulting forecasts are positive.
@@
 #| fig-height: 8
-#| fig-pos: "t"
-#| fig-cap: Solution 2. We use a left shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022. After October 2022, there are no non-zero predictors.
+#| fig-pos: "b"
+#| fig-cap: Solution 2. We use a level shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022. After October 2022, all intervention variables are set to zero.
 tar_read(tsol2_plot)
@@
 #| fig-height: 8
-#| fig-pos: "t"
+#| fig-pos: "b"
 #| fig-cap: Solution 3.
@@
 
-The third solution involved setting the observations during the disruption period to missing, and then fitting an ARIMA model to the series.
+The third solution (@fig-tsol3-plot) involved setting the observations during the disruption period to missing, and then fitting an ARIMA model to the series.
 Consequently, the first three years of forecasts show what might have been without the COVID-19 pandemic, based on the history to the end of 2019.
@@
 #| fig-height: 8
-#| fig-pos: "t"
+#| fig-pos: "!b"
 #| fig-cap: Solution 4. We replace the observations from March 2020 to October 2022 with the average of the same month in the three years prior to March 2020.
@@
 
-Finally, we show in @fig-tsol4-plot the forecasts obtained using solution 4.
+We show in @fig-tsol4-plot the forecasts obtained using solution 4.
 Here we have replaced observations between March 2020 and October 2022 with estimates based on the average of the same month in the three years prior to March 2020.
 The resulting forecasts are similar to those from solution 3, but with narrower prediction intervals because the model is (falsely) assuming that the "observations" during the disruption are real.
+
+Finally, in @fig-tensemble-plot, we show the ensemble forecasts obtained by averaging the forecasts from the four solutions discussed above.
+The first and fourth solutions carry double weight, because both the ETS and ARIMA forecasts were included in the ensemble.
+This time, no prediction intervals have been computed.
+The first set of forecasts have not taken any account of the disruption, and so are almost identical to those obtained in the previous plots.
+The last set of forecasts are reasonably good, showing what is obtained by combining the forecasting methods used to handle the disruption period.
+However, the forecasts for 2021 and 2022 are particularly poor, because they are averaging forecasts that aim to achieve different objectives, and are affected by the poor forecasts obtained using the dynamic regression model.
+
+```{r}
+#| label: fig-tensemble-plot
+#| fig-height: 8
+#| fig-pos: "t"
+#| fig-cap: Ensemble solution, combining the previous four approaches shown in Figures [-@fig-tsol1-plot]--[-@fig-tsol4-plot].
+tar_read(tensemble_plot)
+```
 

### 2024-01-08 R/tourism.R

@@ -138,3 +138,22 @@ tourism_sol4 <- function(austa) {
     forecast(h = 12) |>
     mutate(.id = paste("Forecasts of", .id + 2019))
 }
+
+# Create ensemble from list of fable objects
+# Assumes the key structures are identical other than .model
+tourism_ensemble <- function(object) {
+  lapply(object, function(x) {
+      x |> as_tibble() |> mutate(.model = NULL)
+    }) |> 
+    bind_rows() |> 
+    group_by(Month) |> 
+    summarise(
+      .id = unique(.id),
+      .mean = mean(.mean)
+    ) |> 
+    mutate(Visitors = distributional::dist_degenerate(.mean)) |> 
+    as_fable(index = Month, key = .id, response = "Visitors", distribution = Visitors)
+}
+
+
+    

### 2024-01-08 _targets.R

@@ -29,6 +29,8 @@ list(
   tar_target(tsol2, tourism_sol2(austa)),
   tar_target(tsol3, tourism_sol3(austa)),
   tar_target(tsol4, tourism_sol4(austa)),
+  tar_target(tensemble, tourism_ensemble(list(tsol1,tsol2,tsol3,tsol4))),
+  
   # Forecast plots
   tar_target(
     tourism_history,
@@ -53,7 +55,11 @@ list(
     tsol4_plot,
     autoplot(tsol4, level = 90) + tourism_history + tourism_labels
   ),
-
+  tar_target(
+    tensemble_plot,
+    autoplot(tensemble, level = NULL) + tourism_history + tourism_labels
+  ),
+  
   # Pedestrians
   tar_target(raw_walkers, fetch_walkers(start = "2019-01-01", end = "2021-12-31")),
   tar_target(lockdowns, lockdown_table()),

### 2024-01-08 fits.qmd
Updated pedestrian analysis

@@
 However, the first period was preceded by over a week when most people elected to stay and work at home, so we have chosen to start the first period on `r format(min(lockdowns$Start), "%d %b %Y")`, to better reflect human behaviour.
-Otherwise, we have used the official lockdown dates.
+Otherwise, we have used the official lockdown dates [@MelbourneLockdowns].
 
-We apply the four solutions discussed in @sec-methods.
-We fit ARIMA and dynamic ARIMA models to the data, with no transformations used.
-We have evaluated the methods by using time series cross-validation, with the initial training set comprising the whole of 2019, and subsequent training sets growing by 1 week at a time, to the end of 2021.
+We evaluate the solutions discussed in @sec-methods by using time series cross-validation, with the initial training set comprising the whole of 2019, and subsequent training sets growing by 1 week at a time, to the end of 2021.
 The test sets are always 1 week long.
 Thus, we have evaluated the forecasts during 2020 and 2021, covering all lockdown periods, and the start of the recovery period.
-The results are shown in Figures [-@fig-walkers-plot1]--[-@fig-walkers-plot4]
+No transformations of the data have been used.
+The results are shown in Figures [-@fig-walkers-plot1]--[-@fig-walkers-plot4].
+Prediction intervals have not been shown to avoid cluttering the plots.
 
@@
 #| fig-height: 3
-#| fig-cap: "Solution 1: the forecasts do not take into account the lockdown periods, but the ARIMA model is able to adjust to the changing level of the series."
+#| fig-cap: "Solution 1: the forecasts do not take into account the lockdown periods, but the ARIMA model is largely able to adjust to the changing level of the series."
 tar_read(walkers_plot1)
 ```
+
+We use ARIMA models in @fig-walkers-plot1, which have mostly been able to adapt to the changing level of the series, apart from during the first few weeks of the first lockdown, and the first day or two of subsequent lockdowns.
+
+In @fig-walkers-plot2, we have used an intervention model, with a dummy variable indicating the lockdown periods.
+This has worked a little better than the ARIMA model used in @fig-walkers-plot1, apart from around the first lockdown period.
+Many people were in self-imposed lockdown before the start of the first lockdown period, and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
 
@@
 #| fig-height: 3
-#| fig-cap: "Solution 2: the forecasts are based on an ARIMA model with a dummy variable indicating when the lockdown periods occurred."
+#| fig-cap: "Solution 2: the forecasts are based on an ARIMA model with a dummy variable indicating when the lockdown periods occurred.
+Many people were in self-imposed lockdown before the start of the first lockdown period, and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
+The model was able to adjust to the changing level of the series during subsequent lockdowns."
 tar_read(walkers_plot2)
@@
 #| fig-height: 3
-#| fig-cap: "Solution 3: the lockdown periods are set to missing, and the ARIMA model uses the remaining data to produce forecasts"
+#| fig-cap: "Solution 3: the lockdown periods are set to missing, and the ARIMA model uses the remaining data to produce forecasts.
+Thus, the forecasts are based on only previous non-lockdown data.
+Naturally, the forecasts during the lockdown periods do not reflect reality, but they can be interpreted as what might have been."
 tar_read(walkers_plot3)
@@
 #| fig-cap: "Solution 4: the lockdown periods are replaced with estimated counts based on an ARIMA model applied to the remaining data.
-Then a new ARIMA model is fitted to the whole data set, and used to produce forecasts."
+Then a new ARIMA model is fitted to the whole data set, and used to produce forecasts.
+As a result, the forecasts during the lockdown periods are not true forecasts, as they use non-lockdown data from the future."
 tar_read(walkers_plot4)
 ```
+
+```{r}
+#| label: fig-walkers-ensemble-plot
+#| fig-height: 3
+#| fig-cap: "Ensemble forecasts based on the first two solutions shown in Figures [-@fig-walkers-plot1] and [-@fig-walkers-plot2].
+Other than having some difficulty around the period of the first lockdown, the results are relatively good."
+tar_read(walkers_ensemble_plot)
+```
+
+@fig-walkers-plot3 shows the results obtained by setting the observations during the lockdown periods to missing, and then fitting an ARIMA model to the remaining data.
+The forecasts during lockdowns show what might have been if that particular lockdown hadn't occurred, based on previous non-lockdown data.
+
+The forecasts shown in @fig-walkers-plot4 are obtained by replacing the observations during the lockdown periods with estimates based on an ARIMA model applied to the remaining data.
+Then a new ARIMA model is fitted to the whole data set, and used to produce forecasts.
+As a result, the "forecasts" appear to interpolate across the lockdown periods, reflecting neither the true lockdown pattern, or the pattern that might have occurred without the lockdowns.
+
+Finally, in @fig-walkers-ensemble-plot, we show the ensemble forecasts obtained by averaging the forecasts from the first two solutions discussed above.
+The results are relatively good, other than having some difficulty around the period of the first lockdown.
 

### 2024-01-08 R/pedestrians.R

@@ -174,15 +174,29 @@ pedestrian_sol4 <- function(walkers, step, init, h) {
 }
 
 pedestrian_fc_plot <- function(fc, data) {
-  fc |>
+  # Plot of data showing lockdown periods
+  p <- pedestrian_plot(data |> filter(Date >= "2020-01-01"))
+  fc <- fc |>
     group_by(.id) |>
     slice(1:7) |>
-    mutate(.id = as.character(.id)) |>
-    ggplot(aes(x = Date)) +
-    geom_line(aes(y = .mean, group = .id), col = "blue") +
-    geom_line(
-      data = data |> filter(Date >= "2020-01-01"),
-      mapping = aes(y = Count)
-    ) +
-    guides(col = "none")
+    mutate(.id = as.character(.id))
+  p + 
+    geom_line(data = fc, mapping = aes(x = Date, y = .mean, group = .id), col = "blue") 
+}
+
+
+# Create ensemble from list of fable objects
+# Assumes the key structures are identical other than .model
+pedestrian_ensemble <- function(object) {
+  lapply(object, function(x) {
+    x |> as_tibble() |> mutate(.model = NULL) |> group_by(.id) |> slice(1:7)
+  }) |> 
+    bind_rows() |> 
+    group_by(Date) |> 
+    summarise(
+      .id = unique(.id),
+      .mean = mean(.mean)
+    ) |> 
+    mutate(Count = distributional::dist_degenerate(.mean)) |> 
+    as_fable(index = Date, key = .id, response = "Count", distribution = Count)
 }

### 2024-01-08 _targets.R

@@ -74,10 +74,12 @@ list(
   tar_target(walkers_sol2, pedestrian_sol2(walkers, step = step, init = init, h = h)),
   tar_target(walkers_sol3, pedestrian_sol3(walkers, step = step, init = init, h = h)),
   tar_target(walkers_sol4, pedestrian_sol4(walkers, step = step, init = init, h = h)),
+  tar_target(wensemble, pedestrian_ensemble(list(walkers_sol1, walkers_sol2))),
   tar_target(walkers_plot1, pedestrian_fc_plot(walkers_sol1, walkers)),
   tar_target(walkers_plot2, pedestrian_fc_plot(walkers_sol2, walkers)),
   tar_target(walkers_plot3, pedestrian_fc_plot(walkers_sol3, walkers)),
   tar_target(walkers_plot4, pedestrian_fc_plot(walkers_sol4, walkers)),
+  tar_target(walkers_ensemble_plot, pedestrian_fc_plot(wensemble, walkers)),
   
   # Attended incidents example
   tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format = "file"),

### 2024-01-08 fits.qmd
Removed ambulance example

@@
 
-## Example: Ambulance attendances
-
-```{r}
-#| label: plot-data
-tar_read(attended_plot)
-tar_read(stl_plot)
-```
-
-```{r}
-tar_load(fit)
-fit |> filter(Area == "BC") |> select(ets) |> components() |> autoplot()
-```
-
-```{r}
-#| label: forecast
-tar_load(fc)
-tar_load(df)
-fc |> filter(Area == "BC", .model=="ets") |> autoplot(df)
-fc |> filter(Area == "BC", .model=="stl_ets") |> autoplot(df)
-fc |> filter(Area == "BC", .model=="arima") |> autoplot(df)
-```
-
 # References

### 2024-01-08 fits.qmd
Added discussion

@@
 
+# Discussion
+
+We have presented several approaches to forecasting in the presence of an interruption.
+Each has its advantages and disadvantages, and the choice of which to use will depend on the situation.
+In fact, we have used almost all of these approaches in our own consulting work.
+An intervention model is often a good solution, provided the intervention can be modelled relatively simply.
+However, if the intervention is complex, then a highly adaptive method is often better.
+The highly adaptive model is also useful when the timing of the interruption is unknown, or when the post-interruption period is expected to be very different from the pre-interruption period.
+The missing value approach is particularly useful when only post-interruption forecasts are required, and forecasts during the interruption period are not needed.
+
+We have not previously used the weighted approach in practice, and software to implement it is not currently available.
+However, it seems to be a good compromise between the highly adaptive model and the intervention model, and we expect it to be useful in many situations, once software is available to implement it easily.
+
+Finally, the ensemble approach is useful when there is uncertainty about which approach to use, and will often result in greater accuracy due the power of averaging.
+
+We have not discussed probabilistic forecasting in any detail in this paper, other than producing some prediction intervals for some of the examples.
+One would expect forecast uncertainty to increase during, and to some extent after, a major disruption.
+The highly adaptive model and innovation model should produce relatively good prediction intervals, although the variance is often underestimated.
+Because the missing value approach is producing counterfactual ("what might have been") forecasts, the prediction intervals will often not include the actual observations during the disruption period.
+Prediction intervals for ensemble forecasts can be obtained by averaging the end points of the component prediction intervals [@lichtendahl2013better], for those models where the prediction intervals seem reasonable.
+
 # References

### 2024-01-08 fits.tex
Added discussion

--- 2024-01-05 fits.tex
@@
 
-The changes in the series as a result of the disruption may be
-relatively simple; for example a left shift at the start of the
-disruption, and another at the end of the disruption. Or they may be
-more complex, with changes to the seasonal patterns, and changes to the
-level, which evolve over time. In this paper, we consider a range of
-models that can be used to handle such changes, and compare their
-performance on some real data sets.
-
 Time series models usually assume that the data evolve in the future in
@@
 past, at least in the short term.
+
+The disruption may result in relatively simple changes in the series;
+for example a level shift at the start of the disruption, and another at
+the end of the disruption. Or they may be more complex, with changes to
+the seasonal patterns, and changes to the level, which evolve over time.
+In this paper, we consider a range of models that can be used to handle
+such changes, and compare their performance on some real data sets.
+
+We introduce several approaches to handling interruptions when
+forecasting in Section~\ref{sec-methods}. We then apply these approaches
+to two real data sets in \textbf{?@sec-sec-examples}. Finally, in
+\textbf{?@sec-conclusions}, we discuss some of the advantages and
+disadvantages of each approach, and provide some recommendations for
+practitioners.
 
@@
 quickly. This has the advantage of being a very simple solution that is
-easy to implement and fast to compute.
+easy to implement and fast to compute. There is no need to explicitly
+model the interruption, and so the model can be used for forecasting
+even if the timing or effect of the interruption is unknown.
 
@@
 
-\subsection{Use a dynamic regression model with intervention
-covariates}\label{use-a-dynamic-regression-model-with-intervention-covariates}
+\subsection{Use an intervention model}\label{use-an-intervention-model}
 
@@
 This has the advantage of retaining the memory of the past, and so the
-seasonal patterns and other dynamics will be retained. However, the
-model will assume that the post-interruption period will be similar to
-the pre-interruption period, and so the prediction intervals may be be
-too narrow.
+seasonal patterns and other dynamics will be retained. This allows the
+change period to be effectively modelled provided the intervention
+variables are chosen well. However, the model will assume that the
+post-interruption period will be similar to the pre-interruption period,
+and so the prediction intervals may be be too narrow, especially if
+there is a lasting effect beyond the end of the interruption.
 
-Advantages: retains full memory of the past, and allows the change
-period to be effectively modelled provided you choose the covariates
-well.
-
-Disadvantages: requires a lot of thought to choose the covariates well.
-Assumes that the post-pandemic period will be similar to the
-pre-pandemic period.
-
-\subsection{Treat the covid period as missing and use a model that
-handles missing
-values}\label{treat-the-covid-period-as-missing-and-use-a-model-that-handles-missing-values}
+\subsection{Treat the interruption period as
+missing}\label{treat-the-interruption-period-as-missing}
 
@@
 
-\subsection{Estimate what might have been and adjust the
-data}\label{estimate-what-might-have-been-and-adjust-the-data}
+\subsection{Estimate what might have
+been}\label{estimate-what-might-have-been}
 
@@
 solution except that the prediction intervals will be narrower because
-the estimation uncertainty has not been taken into account.
+the estimation uncertainty has not been taken into account. On the other
+hand, it is more flexible than the previous approach because models that
+do not handle missing values can then be used.
 
-\subsection{Over sampling (weighting)}\label{over-sampling-weighting}
+An alternative would be to take the model estimated under the previous
+solution (setting the observations during the disruption to missing),
+and use it to estimate what might have been during the disruption. Then
+the model can be re-estimated using the adjusted data. Note that
+forecasts obtained in this way during the disruption will not be true
+forecasts, because they will have used data from the future in computing
+the adjusted data during the disruption. But post-disruption forecasts
+will be true forecasts, and should be almost the same as those obtained
+using the previous solution.
 
-Oversampling is a technique used in machine learning to address class
-imbalance within a dataset. In this context, the goal is to ensure that
-the model is more influenced by the patterns observed before the
-disruption period while still accounting for the observations during the
+\subsection{Downweight the interruption
+period}\label{downweight-the-interruption-period}
+
+Intuitively, we want the model to be more influenced by the patterns
+observed before the disruption period than those during the disruption,
+but we still want to take some account of the observations during the
 disruption period.
@@
 pre-disruptive period while still accounting for the disruptive period's
-influence. It allows the model to better capture the underlying patterns
-of the time series data, considering both periods, and potentially
-producing more accurate forecasts.
+influence. It allows the model to better capture the underlying dynamics
+of the time series in non-disrupted periods, which may be essential for
+accurate forecasting in the post-disruption era.
 
-Advantages:
+If the weights during the disruption are set to zero, and the weights at
+other times are set to 1, this becomes equivalent to the missing value
+solution above. Rather than ignoring the data during the disruption, the
+weighting approach allows the model to not be completely blind to the
+disruption's effects, and adapt to changes or shifts in the time series
+caused by the event.
 
-Preserves pre-event patterns: By assigning higher weights to the
-pre-disruption time series data, the model is encouraged to focus more
-on the patterns observed before the disruptive event. This can be
-beneficial in capturing the underlying dynamics of the time series that
-might be essential for accurate forecasting in the post-disruption era.
-
-Robustness to disruption impact: The weighting approach allows the model
-to still take the disruption period into account, even if it's
-considered less influential. This ensures that the model is not
-completely blind to the disruption's effects and can adapt to changes or
-shifts in the time series caused by the event.
-
-Easy to Implement: ??
-
-Interpretability?:
-
-Disadvantages:
+While conceptually simple to implement, in practice, this approach may
+require more work than the other methods discussed here as most
+forecasting software does not allow for the explicit use of weights.
 
@@
 
-\section{Examples}\label{examples}
+As with may forecasting problems, taken an ensemble approach often leads
+to more accurate forecasts \autocite{combinations}. Here, we could
+combine some or all four of the approaches discussed above to obtain an
+ensemble forecast. In fact, if we were unable to implement the weighted
+approach, by using an ensemble of the other approaches we are
+effectively downweighting the observations during the disruption period
+as they are only explicitly used in the first two approaches (using a
+highly adaptive model, or using an intervention model).
 
-Let's consider some examples
+However, a disadvantage of this approach is that it is averaging
+forecasts that aim to achieve different objectives. For example, the
+highly adaptive model and the intervention model are trying to forecast
+what happened during the disruption, while the missing value and
+estimation approaches are trying to forecast what might have happened.
+So the ensemble approach will lie somewhere between these two
+objectives.
 
-\subsection{Australian tourism}\label{australian-tourism}
+\section{Examples}\label{sec-examples}
+
+Let's consider some examples to data observed during the last few years,
+where the effect of the COVID-19 pandemic has been particularly evident.
+
+\subsection{Australian tourism}\label{sec-tourism}
 
@@
 
-We apply the four solutions discussed in the previous section to these
-data, making 12 month forecasts at the end of each year from 2019 to
-2022. We fit ETS, ARIMA and dynamic ARIMA models to the data
+We apply the first four solutions discussed in the previous section to
+these data, making 12 month forecasts at the end of each year from 2019
+to 2022. We fit ETS, ARIMA and dynamic ARIMA models to the data
 \autocite{fpp3}, first applying a log transformation to ensure the
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 
@@
 
-\caption{\label{fig-tsol2-plot}Solution 2. We use a left shift from
+\caption{\label{fig-tsol2-plot}Solution 2. We use a level shift from
 March 2020 to October 2022, and a ramp from October 2021 to October
-2022. After October 2022, there are no non-zero predictors.}
+2022. After October 2022, all intervention variables are set to zero.}
 
@@
 
-The third solution involved setting the observations during the
-disruption period to missing, and then fitting an ARIMA model to the
-series. Consequently, the first three years of forecasts show what might
-have been without the COVID-19 pandemic, based on the history to the end
-of 2019. The final forecasts for 2023 use the data to the end of 2019,
-and the two observations in November and December 2022. These are much
-better, but the prediction intervals are too narrow, as there hasn't
-been sufficient recent data.
+The third solution (Figure~\ref{fig-tsol3-plot}) involved setting the
+observations during the disruption period to missing, and then fitting
+an ARIMA model to the series. Consequently, the first three years of
+forecasts show what might have been without the COVID-19 pandemic, based
+on the history to the end of 2019. The final forecasts for 2023 use the
+data to the end of 2019, and the two observations in November and
+December 2022. These are much better, but the prediction intervals are
+too narrow, as there hasn't been sufficient recent data.
 
-\begin{figure}[t]
+\begin{figure}[!b]
 
@@
 
-Finally, we show in Figure~\ref{fig-tsol4-plot} the forecasts obtained
-using solution 4. Here we have replaced observations between March 2020
-and October 2022 with estimates based on the average of the same month
-in the three years prior to March 2020. The resulting forecasts are
-similar to those from solution 3, but with narrower prediction intervals
-because the model is (falsely) assuming that the ``observations'' during
-the disruption are real.
+We show in Figure~\ref{fig-tsol4-plot} the forecasts obtained using
+solution 4. Here we have replaced observations between March 2020 and
+October 2022 with estimates based on the average of the same month in
+the three years prior to March 2020. The resulting forecasts are similar
+to those from solution 3, but with narrower prediction intervals because
+the model is (falsely) assuming that the ``observations'' during the
+disruption are real.
+
+Finally, in Figure~\ref{fig-tensemble-plot}, we show the ensemble
+forecasts obtained by averaging the forecasts from the four solutions
+discussed above. The first and fourth solutions carry double weight,
+because both the ETS and ARIMA forecasts were included in the ensemble.
+This time, no prediction intervals have been computed. The first set of
+forecasts have not taken any account of the disruption, and so are
+almost identical to those obtained in the previous plots. The last set
+of forecasts are reasonably good, showing what is obtained by combining
+the forecasting methods used to handle the disruption period. However,
+the forecasts for 2021 and 2022 are particularly poor, because they are
+averaging forecasts that aim to achieve different objectives, and are
+affected by the poor forecasts obtained using the dynamic regression
+model.
+
+\begin{figure}[t]
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-tensemble-plot-1.pdf}
+
+}
+
+\caption{\label{fig-tensemble-plot}Ensemble solution, combining the
+previous four approaches shown in Figures
+\ref{fig-tsol1-plot}--\ref{fig-tsol4-plot}.}
+
+\end{figure}%
 
@@
 the first period on 23 Mar 2020, to better reflect human behaviour.
-Otherwise, we have used the official lockdown dates.
+Otherwise, we have used the official lockdown dates
+\autocite{MelbourneLockdowns}.
 
-We apply the four solutions discussed in Section~\ref{sec-methods}. We
-fit ETS, ARIMA and dynamic ARIMA models to the data, with no
-transformations used. We have evaluated the methods by using time series
-cross-validation, with the initial training set comprising the whole of
-2019, and subsequent training sets growing by 1 week at a time, to the
-end of 2021. The test sets are always 3 weeks long. Thus, we have
-evaluated the forecasts during 2020 and 2021, covering all lockdown
-periods, and the start of the recovery period. The results are shown in
-\textbf{?@fig-walkers-plot1}.
+We evaluate the solutions discussed in Section~\ref{sec-methods} by
+using time series cross-validation, with the initial training set
+comprising the whole of 2019, and subsequent training sets growing by 1
+week at a time, to the end of 2021. The test sets are always 1 week
+long. Thus, we have evaluated the forecasts during 2020 and 2021,
+covering all lockdown periods, and the start of the recovery period. No
+transformations of the data have been used. The results are shown in
+Figures \ref{fig-walkers-plot1}--\ref{fig-walkers-plot4}. Prediction
+intervals have not been shown to avoid cluttering the plots.
 
-\subsection{Example: Ambulance
-attendances}\label{example-ambulance-attendances}
+\begin{figure}
 
-\includegraphics{fits_files/figure-pdf/plot-data-1.pdf}
+\centering{
 
-\includegraphics{fits_files/figure-pdf/plot-data-2.pdf}
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot1-1.pdf}
 
-\includegraphics{fits_files/figure-pdf/unnamed-chunk-10-1.pdf}
+}
 
-\includegraphics{fits_files/figure-pdf/forecast-1.pdf}
+\caption{\label{fig-walkers-plot1}Solution 1: the forecasts do not take
+into account the lockdown periods, but the ARIMA model is largely able
+to adjust to the changing level of the series.}
 
-\includegraphics{fits_files/figure-pdf/forecast-2.pdf}
+\end{figure}%
 
-\includegraphics{fits_files/figure-pdf/forecast-3.pdf}
+We use ARIMA models in Figure~\ref{fig-walkers-plot1}, which have mostly
+been able to adapt to the changing level of the series, apart from
+during the first few weeks of the first lockdown, and the first day or
+two of subsequent lockdowns.
+
+In Figure~\ref{fig-walkers-plot2}, we have used an intervention model,
+with a dummy variable indicating the lockdown periods. This has worked a
+little better than the ARIMA model used in
+Figure~\ref{fig-walkers-plot1}, apart from around the first lockdown
+period. Many people were in self-imposed lockdown before the start of
+the first lockdown period, and were reluctant to return to the city
+after it ended, resulting in poor forecasts on either side of the first
+lockdown.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot2-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot2}Solution 2: the forecasts are based on
+an ARIMA model with a dummy variable indicating when the lockdown
+periods occurred. Many people were in self-imposed lockdown before the
+start of the first lockdown period, and were reluctant to return to the
+city after it ended, resulting in poor forecasts on either side of the
+first lockdown. The model was able to adjust to the changing level of
+the series during subsequent lockdowns.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot3-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot3}Solution 3: the lockdown periods are
+set to missing, and the ARIMA model uses the remaining data to produce
+forecasts. Thus, the forecasts are based on only previous non-lockdown
+data. Naturally, the forecasts during the lockdown periods do not
+reflect reality, but they can be interpreted as what might have been.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot4-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot4}Solution 4: the lockdown periods are
+replaced with estimated counts based on an ARIMA model applied to the
+remaining data. Then a new ARIMA model is fitted to the whole data set,
+and used to produce forecasts. As a result, the forecasts during the
+lockdown periods are not true forecasts, as they use non-lockdown data
+from the future.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-ensemble-plot-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-ensemble-plot}Ensemble forecasts based on
+the first two solutions shown in Figures \ref{fig-walkers-plot1} and
+\ref{fig-walkers-plot2}. Other than having some difficulty around the
+period of the first lockdown, the results are relatively good.}
+
+\end{figure}%
+
+Figure~\ref{fig-walkers-plot3} shows the results obtained by setting the
+observations during the lockdown periods to missing, and then fitting an
+ARIMA model to the remaining data. The forecasts during lockdowns show
+what might have been if that particular lockdown hadn't occurred, based
+on previous non-lockdown data.
+
+The forecasts shown in Figure~\ref{fig-walkers-plot4} are obtained by
+replacing the observations during the lockdown periods with estimates
+based on an ARIMA model applied to the remaining data. Then a new ARIMA
+model is fitted to the whole data set, and used to produce forecasts. As
+a result, the ``forecasts'' appear to interpolate across the lockdown
+periods, reflecting neither the true lockdown pattern, or the pattern
+that might have occurred without the lockdowns.
+
+Finally, in Figure~\ref{fig-walkers-ensemble-plot}, we show the ensemble
+forecasts obtained by averaging the forecasts from the first two
+solutions discussed above. The results are relatively good, other than
+having some difficulty around the period of the first lockdown.
+
+\section{Discussion}\label{discussion}
+
+We have presented several approaches to forecasting in the presence of
+an interruption. Each has its advantages and disadvantages, and the
+choice of which to use will depend on the situation. In fact, we have
+used almost all of these approaches in our own consulting work. An
+intervention model is often a good solution, provided the intervention
+can be modelled relatively simply. However, if the intervention is
+complex, then a highly adaptive method is often better. The highly
+adaptive model is also useful when the timing of the interruption is
+unknown, or when the post-interruption period is expected to be very
+different from the pre-interruption period. The missing value approach
+is particularly useful when only post-interruption forecasts are
+required, and forecasts during the interruption period are not needed.
+
+We have not previously used the weighted approach in practice, and
+software to implement it is not currently available. However, it seems
+to be a good compromise between the highly adaptive model and the
+intervention model, and we expect it to be useful in many situations,
+once software is available to implement it easily.
+
+Finally, the ensemble approach is useful when there is uncertainty about
+which approach to use, and will often result in greater accuracy due the
+power of averaging.
+
+We have not discussed probabilistic forecasting in any detail in this
+paper, other than producing some prediction intervals for some of the
+examples. One would expect forecast uncertainty to increase during, and
+to some extent after, a major disruption. The highly adaptive model and
+innovation model should produce relatively good prediction intervals,
+although the variance is often underestimated. Because the missing value
+approach is producing counterfactual (``what might have been'')
+forecasts, the prediction intervals will often not include the actual
+observations during the disruption period. Prediction intervals for
+ensemble forecasts can be obtained by averaging the end points of the
+component prediction intervals, for those models where the prediction
+intervals seem reasonable.
 

### 2024-01-08 fits.qmd
Improved naming of methods

@@
 author:
-- familyname: Rostami-Tabar
-  othernames: Bahman
-  address:
-    - Cardiff Business School
-    - Wales CF10 3EU
-    - United Kingdom
-  email: rostami-tabarb@cardiff.ac.uk
-  correspondingauthor: true
 - familyname: Hyndman
@@
   email: Rob.Hyndman@monash.edu
+  correspondingauthor: true
+- familyname: Rostami-Tabar
+  othernames: Bahman
+  address:
+    - Cardiff Business School
+    - Wales CF10 3EU
+    - United Kingdom
+  email: rostami-tabarb@cardiff.ac.uk
 abstract: "A brief summary of our ideas"
@@
 
-## Treat the interruption period as missing
+## Set to missing
 
@@
 #| fig-height: 8
-#| fig-pos: "t"
-#| fig-cap: "Solution 1. Forecasts from ETS and ARIMA models. Neither works particularly well for disruptions of this magnitude."
+#| fig-pos: "!b"
+#| fig-cap: "**Highly adaptive models**. Forecasts from ETS and ARIMA models. Neither works particularly well for disruptions of this magnitude."
 tar_read(tsol1_plot)
@@
 #| fig-height: 8
-#| fig-pos: "b"
-#| fig-cap: Solution 2. We use a level shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022. After October 2022, all intervention variables are set to zero.
+#| fig-pos: "!b"
+#| fig-cap: "**Intervention model**.
+We use a level shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022.
+After October 2022, all intervention variables are set to zero."
 tar_read(tsol2_plot)
@@
 #| fig-height: 8
-#| fig-pos: "b"
-#| fig-cap: Solution 3.
+#| fig-pos: "!b"
+#| fig-cap: "**Set to missing**.
 The period from March 2020 to October 2022 is set to missing.
 So the first three years of forecasts show what might have been.
-The fourth set of forecasts uses observations from November and December 2022, and so the forecasts have been adjusted downwards.
+The fourth set of forecasts uses observations from November and December 2022, and so the forecasts have been adjusted downwards."
 tar_read(tsol3_plot)
@@
 #| fig-pos: "!b"
-#| fig-cap: Solution 4. We replace the observations from March 2020 to October 2022 with the average of the same month in the three years prior to March 2020.
+#| fig-cap: "**Estimate what might have been**. We replace the observations from March 2020 to October 2022 with the average of the same month in the three years prior to March 2020."
 tar_read(tsol4_plot)
@@
 #| fig-height: 8
-#| fig-pos: "t"
-#| fig-cap: Ensemble solution, combining the previous four approaches shown in Figures [-@fig-tsol1-plot]--[-@fig-tsol4-plot].
+#| fig-pos: "!t"
+#| fig-cap: "**Ensemble approach**, combining the previous four approaches shown in Figures [-@fig-tsol1-plot]--[-@fig-tsol4-plot]."
 tar_read(tensemble_plot)
@@
 #| fig-height: 3
-#| fig-cap: "Solution 1: the forecasts do not take into account the lockdown periods, but the ARIMA model is largely able to adjust to the changing level of the series."
+#| fig-cap: "**Highly adaptive model**: the forecasts do not take into account the lockdown periods, but the ARIMA model is largely able to adjust to the changing level of the series."
 tar_read(walkers_plot1)
@@
 #| fig-height: 3
-#| fig-cap: "Solution 2: the forecasts are based on an ARIMA model with a dummy variable indicating when the lockdown periods occurred.
+#| fig-cap: "**Intervention model**: the forecasts are based on an ARIMA model with a dummy variable indicating when the lockdown periods occurred.
 Many people were in self-imposed lockdown before the start of the first lockdown period, and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
@@
 #| fig-height: 3
-#| fig-cap: "Solution 3: the lockdown periods are set to missing, and the ARIMA model uses the remaining data to produce forecasts.
+#| fig-cap: "**Set to missing**: the lockdown periods are set to missing, and the ARIMA model uses the remaining data to produce forecasts.
 Thus, the forecasts are based on only previous non-lockdown data.
@@
 #| fig-height: 3
-#| fig-cap: "Solution 4: the lockdown periods are replaced with estimated counts based on an ARIMA model applied to the remaining data.
+#| fig-cap: "**Estimate what might have been**: the lockdown periods are replaced with estimated counts based on an ARIMA model applied to the remaining data.
 Then a new ARIMA model is fitted to the whole data set, and used to produce forecasts.
@@
 #| fig-height: 3
-#| fig-cap: "Ensemble forecasts based on the first two solutions shown in Figures [-@fig-walkers-plot1] and [-@fig-walkers-plot2].
+#| fig-cap: "**Ensemble forecasts** based on the first two solutions shown in Figures [-@fig-walkers-plot1] and [-@fig-walkers-plot2].
 Other than having some difficulty around the period of the first lockdown, the results are relatively good."

### 2024-01-08 fits.tex
Improved naming of methods

@@
 
-\author{Bahman~Rostami-Tabar, Rob J~Hyndman}
-\addresses{\textbf{Bahman Rostami-Tabar}\newline
-Cardiff Business School\newline
-Wales CF10 3EU\newline
-United Kingdom\newline
-{Email: rostami-tabarb@cardiff.ac.uk}\newline Corresponding author\newline\\[0.5cm]
-\textbf{Rob J Hyndman}\newline
+\author{Rob J~Hyndman, Bahman~Rostami-Tabar}
+\addresses{\textbf{Rob J Hyndman}\newline
 Department of Econometrics \& Business Statistics\newline
@@
 Australia\newline
-{Email: Rob.Hyndman@monash.edu}\\[0.5cm]
+{Email: Rob.Hyndman@monash.edu}\newline Corresponding author\newline\\[0.5cm]
+\textbf{Bahman Rostami-Tabar}\newline
+Cardiff Business School\newline
+Wales CF10 3EU\newline
+United Kingdom\newline
+{Email: rostami-tabarb@cardiff.ac.uk}\\[0.5cm]
 }
@@
 \makeatletter
- \lfoot{\sf Rostami-Tabar, Hyndman: \@date}
+ \lfoot{\sf Hyndman, Rostami-Tabar: \@date}
 \makeatother
@@
 
-\subsection{Treat the interruption period as
-missing}\label{treat-the-interruption-period-as-missing}
+\subsection{Set to missing}\label{set-to-missing}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[!b]
 
@@
 
-\caption{\label{fig-tsol1-plot}Solution 1. Forecasts from ETS and ARIMA
-models. Neither works particularly well for disruptions of this
-magnitude.}
+\caption{\label{fig-tsol1-plot}\textbf{Highly adaptive models}.
+Forecasts from ETS and ARIMA models. Neither works particularly well for
+disruptions of this magnitude.}
 
@@
 
-\begin{figure}[b]
+\begin{figure}[!b]
 
@@
 
-\caption{\label{fig-tsol2-plot}Solution 2. We use a level shift from
-March 2020 to October 2022, and a ramp from October 2021 to October
-2022. After October 2022, all intervention variables are set to zero.}
+\caption{\label{fig-tsol2-plot}\textbf{Intervention model}. We use a
+level shift from March 2020 to October 2022, and a ramp from October
+2021 to October 2022. After October 2022, all intervention variables are
+set to zero.}
 
@@
 
-\caption{\label{fig-tsol3-plot}Solution 3. The period from March 2020 to
-October 2022 is set to missing. So the first three years of forecasts
-show what might have been. The fourth set of forecasts uses observations
-from November and December 2022, and so the forecasts have been adjusted
-downwards.}
+\caption{\label{fig-tsol3-plot}\textbf{Set to missing}. The period from
+March 2020 to October 2022 is set to missing. So the first three years
+of forecasts show what might have been. The fourth set of forecasts uses
+observations from November and December 2022, and so the forecasts have
+been adjusted downwards.}
 
@@
 
-\caption{\label{fig-tsol4-plot}Solution 4. We replace the observations
-from March 2020 to October 2022 with the average of the same month in
-the three years prior to March 2020.}
+\caption{\label{fig-tsol4-plot}\textbf{Estimate what might have been}.
+We replace the observations from March 2020 to October 2022 with the
+average of the same month in the three years prior to March 2020.}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[!t]
 
@@
 
-\caption{\label{fig-tensemble-plot}Ensemble solution, combining the
-previous four approaches shown in Figures
+\caption{\label{fig-tensemble-plot}\textbf{Ensemble approach}, combining
+the previous four approaches shown in Figures
 \ref{fig-tsol1-plot}--\ref{fig-tsol4-plot}.}
@@
 
-\caption{\label{fig-walkers-plot1}Solution 1: the forecasts do not take
-into account the lockdown periods, but the ARIMA model is largely able
-to adjust to the changing level of the series.}
+\caption{\label{fig-walkers-plot1}\textbf{Highly adaptive model}: the
+forecasts do not take into account the lockdown periods, but the ARIMA
+model is largely able to adjust to the changing level of the series.}
 
@@
 
-\caption{\label{fig-walkers-plot2}Solution 2: the forecasts are based on
-an ARIMA model with a dummy variable indicating when the lockdown
-periods occurred. Many people were in self-imposed lockdown before the
-start of the first lockdown period, and were reluctant to return to the
-city after it ended, resulting in poor forecasts on either side of the
-first lockdown. The model was able to adjust to the changing level of
-the series during subsequent lockdowns.}
+\caption{\label{fig-walkers-plot2}\textbf{Intervention model}: the
+forecasts are based on an ARIMA model with a dummy variable indicating
+when the lockdown periods occurred. Many people were in self-imposed
+lockdown before the start of the first lockdown period, and were
+reluctant to return to the city after it ended, resulting in poor
+forecasts on either side of the first lockdown. The model was able to
+adjust to the changing level of the series during subsequent lockdowns.}
 
@@
 
-\caption{\label{fig-walkers-plot3}Solution 3: the lockdown periods are
-set to missing, and the ARIMA model uses the remaining data to produce
-forecasts. Thus, the forecasts are based on only previous non-lockdown
-data. Naturally, the forecasts during the lockdown periods do not
-reflect reality, but they can be interpreted as what might have been.}
+\caption{\label{fig-walkers-plot3}\textbf{Set to missing}: the lockdown
+periods are set to missing, and the ARIMA model uses the remaining data
+to produce forecasts. Thus, the forecasts are based on only previous
+non-lockdown data. Naturally, the forecasts during the lockdown periods
+do not reflect reality, but they can be interpreted as what might have
+been.}
 
@@
 
-\caption{\label{fig-walkers-plot4}Solution 4: the lockdown periods are
-replaced with estimated counts based on an ARIMA model applied to the
-remaining data. Then a new ARIMA model is fitted to the whole data set,
-and used to produce forecasts. As a result, the forecasts during the
-lockdown periods are not true forecasts, as they use non-lockdown data
-from the future.}
+\caption{\label{fig-walkers-plot4}\textbf{Estimate what might have
+been}: the lockdown periods are replaced with estimated counts based on
+an ARIMA model applied to the remaining data. Then a new ARIMA model is
+fitted to the whole data set, and used to produce forecasts. As a
+result, the forecasts during the lockdown periods are not true
+forecasts, as they use non-lockdown data from the future.}
 
@@
 
-\caption{\label{fig-walkers-ensemble-plot}Ensemble forecasts based on
-the first two solutions shown in Figures \ref{fig-walkers-plot1} and
-\ref{fig-walkers-plot2}. Other than having some difficulty around the
-period of the first lockdown, the results are relatively good.}
+\caption{\label{fig-walkers-ensemble-plot}\textbf{Ensemble forecasts}
+based on the first two solutions shown in Figures
+\ref{fig-walkers-plot1} and \ref{fig-walkers-plot2}. Other than having
+some difficulty around the period of the first lockdown, the results are
+relatively good.}
 
@@
 ensemble forecasts can be obtained by averaging the end points of the
-component prediction intervals, for those models where the prediction
-intervals seem reasonable.
+component prediction intervals \autocite{lichtendahl2013better}, for
+those models where the prediction intervals seem reasonable.
 

### 2024-02-14 fits.qmd
add some texts to the introduction

@@
 
-Time series are sometimes interrupted by unusual events; for example, a natural disaster may occur, or a there may be a temporary policy change.
+Time series forecasting models use historical data to forecast future values based on consistent patterns observed in the past.
+However, time series are sometimes interrupted by unusual events, that jeopardise the regularity of the time series pattern.
+Such interruptions may take a variety of forms including from a temporary policy change to a natural disaster; a few examples are provided below:
+
+- Outbreaks, epidemics, and pandemics may all have a substantial influence on population behaviour, causing fluctuations and affecting time series data in a wide range of areas such as hospital admissions, pharmaceutical sales, transportation, retail sales, tourism, transportation, and traffic.
+
+- Supply chain interruptions, such as strikes, trade disputes, or shortages, may cause inconsistencies in data, including sales, inventory levels, production, and deliveries .
+Firms may need to shut down for repair, overhaul, or maintenance purposes.
+These events force production to halt, leaving a gap in the data.
+
+- Equipment failures in transportation: Vehicle, aeroplane, or rail malfunctions or breakdowns may result in missing or incorrect data points in transportation-related datasets .
+
+- Cyberattacks: Cybersecurity breaches may jeopardise data integrity, resulting in incorrect readings in datasets including financial transactions, network traffic, and user activity.
+
+The occurrence of such events presents a tremendous challenge to time series forecasting methods, limiting their capacity to automatically and accurately capture systematic information and forecast future values.
+Therefore, practitioners face a critical question: how to effectively handle time series data impacted by such disruptive events?
 Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to lockdowns and other restrictions.
-In this paper, we consider the problem of forecasting during such an event, and after it has occurred.
+This study is motivated by numerous conversations with academics and practitioners, which demonstrate widespread concern about dealing with data affected by events like the COVID-19 pandemic.
 
-This is a different problem from change point detection.
-In the situations we consider, we know that a change has occurred, and we want to forecast the future after the change.
-Change point detection is about identifying *when* a change has occurred.
-
-Time series models usually assume that the data evolve in the future in a similar way to how they have evolved in the past.
+Time series models usually assume that the data will evolve in the future in a similar way to how it has evolved in the past.
 But a big event (such as COVID-19) can result in a future that is different from the past, at least in the short term.
@@
 
+<!-- Time series are sometimes interrupted by unusual events; for example, a natural disaster may occur, or a there may be a temporary policy change.
+Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to lockdowns and other restrictions.
+-->
+
+It is important to note that in this study, we consider the problem of forecasting both during and after the event.
+This is a separate problem from change point detection.
+In the situations we investigate, we know that a change has occurred and want to forecast what will happen next.
+Change point detection refers to determining *when* a change occurred.
+
+The paper proposes and evaluates multiple forecasting strategies designed to handle interruptions in time series data caused by disruptive events.
+These approaches include a range of models capable of capturing different types of changes, from simple level shifts to more complex alterations in seasonal patterns and trends.
+By comparing the performance of these models on two real data sets, the paper offers valuable insights into the effectiveness and practical implications of each approach.The paper offers practical guidance for practitioners seeking to improve the accuracy and reliability of time series forecasts in the presence of disruptive events.
+By discussing the advantages and disadvantages of each strategy and providing recommendations based on the findings, the paper equips practitioners with actionable insights for addressing the challenges associated with forecasting during and after disruptive events.
+Furthermore, by adhering to reproducibility principles and providing both data and R codes for the forecasting models, the research not only contributes to transparency but also promotes the generalizability of the suggested strategies to a variety of domains.
+This openness allows for the replication of results, ensuring the reliability of findings, and enhancing the accessibility and utility of study outputs for the larger research community.
+
 We introduce several approaches to handling interruptions when forecasting in @sec-methods.
@@
 Finally, in @sec-conclusions, we discuss some of the advantages and disadvantages of each approach, and provide some recommendations for practitioners.
+
+# Related work
+
+Stock out, maintenancem sensors
 
@@
 
+# Conclusion
+
 # References

### 2024-02-17 fits.qmd
first complete drfat

--- 2024-01-08 fits.qmd
@@
   email: rostami-tabarb@cardiff.ac.uk
-abstract: "A brief summary of our ideas"
-keywords: "blah, blah"
+abstract: "Forecasting interrupted time series data is a major challenge for forecasting teams, especially in light of events such as the COVID-19 pandemic.
+This paper investigates several strategies for dealing with interruptions in time series forecasting, including highly adaptable models, intervention models, marking interrupted periods as missing, forecasting what may have been, downweighting the interruption period, and ensemble models.
+Each approach offers specific advantages and disadvantages, such as adaptability, memory retention, data integrity, flexibility, and accuracy.
+We evaluate the effectiveness of these strategies using two actual datasets that were interrupted by COVID-19, and we provide recommendations for how to handle these interruptions.
+This work contributes to the literature on time series forecasting, offering insights for academics and practitioners dealing with interrupted data in numerous domains."
+keywords: "Forecasting, interrupted time series, disruptive events, COVID-19"
 wpnumber: no/yr
@@
 #theme_set(theme_get() + theme(text = element_text(family = "Fira Sans")))
-
+ggplot2::theme_set(ggthemes::theme_few())
 # Load main data objects from targets folder
@@
 
-Time series are sometimes interrupted by unusual events; for example, a natural disaster may occur, or a there may be a temporary policy change.
-Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to lockdowns and other restrictions.
-In this paper, we consider the problem of forecasting during such an event, and after it has occurred.
+Time series forecasting models use historical data to estimate future values [@fildes2008forecasting] based on consistent patterns observed in the past.
+However, time series are sometimes disrupted by unusual events that jeopardise the regularity of the time series pattern.
+This can be defined as an *interrupted time series*.
+Time series data may be disrupted by a variety of reasons, ranging from temporary policy changes to natural disasters.
+Outbreaks, epidemics, and pandemics, for example, may have a considerable impact on population behavior, causing discrepancies and interruptions of data in a variety of industries such as healthcare, pharmaceuticals, transportation, retail, tourism, and traffic management.
+Similarly, supply chain disruptions induced by factors like strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps.
+Transportation-related data may be interrupted due to equipment failures, such as malfunctions or breakdowns in vehicles, aircraft, or rail systems, resulting in missing or inaccurate data points.
+Furthermore, cybersecurity breaches data quality by possibly skewing readings in datasets relating to financial transactions, network traffic, and user activity, complicating time series analysis.
 
-This is a different problem from change point detection.
-In the situations we consider, we know that a change has occurred, and we want to forecast the future after the change.
-Change point detection is about identifying *when* a change has occurred.
+The disruptions may result in relatively simple changes in the series; for example a level shift at the start of the disruption, and another at the end of the disruption.
+Or they may be more complex, with changes to the seasonal patterns, and changes to the level, which evolve over time.
 
-Time series models usually assume that the data evolve in the future in a similar way to how they have evolved in the past.
+The presence of such disrupted events poses a significant challenge to time series forecasting approaches, restricting their ability to reliably capture systematic information and forecast future values.
+The implications of these temporal interruptions are far-reaching: time series forecasting techniques, which are a key component of the forecaster's toolbox, may no longer be viable options since time series models usually assume that the data will evolve in the future in a similar way to how it has evolved in the past.
 But a big event (such as COVID-19) can result in a future that is different from the past, at least in the short term.
+Therefore, practitioners face a critical question: how to effectively forecast time series data that are impacted by such disruptive events?
+Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to policy interventions, lockdowns, and other restrictions.
+This study is motivated by numerous conversations with academics and practitioners, which demonstrate widespread concern about dealing with data affected by events like the COVID-19 pandemic.
+In this paper, we consider a range of models that can be used to forecast time series influenced by disrupted events, and compare their performance on some real data sets.
 
-The disruption may result in relatively simple changes in the series; for example a level shift at the start of the disruption, and another at the end of the disruption.
-Or they may be more complex, with changes to the seasonal patterns, and changes to the level, which evolve over time.
-In this paper, we consider a range of models that can be used to handle such changes, and compare their performance on some real data sets.
+It is important to note that in this study, we consider the problem of forecasting interrupted time series both during and after the event.
+This is a separate problem from change point detection [@truong2020selective] and anomaly detection [@talagala2020anomaly].
+In the situations we investigate, we know that a change has occurred that causes unusual data points and want to forecast what will happen next.
+Change point detection refers to determining *when* a change occurred [@blazquez2021review], and anomaly detection aims to identify unusual observations with the intention of either minimising their value, or paying particular attention to them, and doing a root-cause investigation.
+Our study focus is also slightly differs from *intervention modeling*, also refereed to as *interrupted time series analysis and modelling* [@mcdowall2019interrupted], which is used when we have data about an outcome over time and the aim is to assess the impact of an intervention, policy, or program adopted for the whole population at a specific time point.
+This is a special case of what we are looking at at this paper.
 
-We introduce several approaches to handling interruptions when forecasting in @sec-methods.
+To our knowledge, this is the first study to offer strategies for forecasting interrupted time series, such as COVID-19 affected data.
+The study proposes and evaluates multiple forecasting strategies designed to handle interruptions in time series forecasting caused by disruptive events.
+These approaches include a range of models capable of capturing different types of changes, from simple level shifts to more complex alterations in seasonal patterns and trends.
+By comparing the performance of these models on two real data sets affected by COVID-19 pandemic, the paper offers valuable insights into the effectiveness and practical implications of each approach.The paper offers practical guidance for practitioners seeking to improve the accuracy and reliability of time series forecasts in the presence of disruptive events.
+By discussing the advantages and disadvantages of each strategy and providing recommendations based on the findings, the paper equips practitioners with actionable insights for addressing the challenges associated with forecasting during and after disruptive events.
+Furthermore, by adhering to reproducibility principles and providing both data and R codes for the forecasting models, the research not only contributes to transparency but also promotes the generalizability of the suggested strategies to a variety of domains.
+This openness allows for the replication of results, ensuring the reliability of findings, and enhancing the accessibility and utility of study outputs for the larger research community.
+
+We introduce several strategies to handling interruptions when forecasting in @sec-methods.
 We then apply these approaches to two real data sets in @sec-sec-examples.
@@
 
+# Related works
+
+The presence of unusual observations in time series analysis is a significant concern that has received attention in numerous studies.
+In this section, we provide a brief overview of some of these relevant publications.
+
+One modelling approach that has gained popularity is the intervention model, also known as interrupted time series (ITS) modelling, introduced @box1975intervention.
+This approach focuses on understanding how and whether outcomes change following the implementation of an intervention, policy, or programme at a specific time point.
+ITS models provide a robust tool for analysing policy and programme evaluation across diverse domains [@bernal2017interrupted;@mcdowall2019interrupted].
+These models can forecast future values by developing models that incorporate a range of components, such as pre-intervention levels, trends, seasonality, covariates related to the intervention time, and other external factors that might affect the outcome variable.
+Another related approach is the multiple regimes model, which is particularly useful in forecasting due to its ability to capture structural changes and transitions observed in interrupted time series data [@Koopg2007].
+
+Statistical models developed to detect changes in time series data and handle unusual observations have also proven beneficial in similar circumstances.
+Methods for detecting changing points in time series data may include simple location shift and scale shift models[@aue2013structural], Bayesian [@barry1993bayesian], and supervised and supervised machine learning [@aminikhanghahi2017survey].
+Anomaly detection seeks to find points or subsequences in time series data that have unexpected or anomalous values.
+Facebook's Prophet is an example of a time series forecasting system that uses changepoint detection and automatically adjusts model parameters to catch the most recent trends in a time series.
+
+There are also statistical models that are used to detect changes in the time series data and that can model unusual observation and have proven effective in similar situations.
+A range of methods such as simple location shift and scale shift models [@aue2013structural], Bayesian [@barry1993bayesian], and supervised and supervised machine learning [@aminikhanghahi2017survey].
+Facebook’s Prophet [@taylor2018forecasting] is an example of a time series forecasting method that can incorporate changepoint detection (automatically or manually), and adjust model parameters to capture the newest trends in a time series.
+Anomaly detection is another related area in time series analysis.
+In the context of anomaly detection, time series anomalies are defined as points or subsequences with unexpected or abnormal values, and the detection goal is to identify such anomalies [@blazquez2021review].
+
+In the context of forecasting disruptions like the COVID-19 pandemic, Bayesian approaches have been proposed for detecting changing points in outbreaks and forecasting case numbers under counterfactual scenarios [@dehning2020inferring].
+Counterfactual analysis examines what would have happened to the forecast variable of interest if the disruptive event had not occurred.
+In line with this, @athanasopoulos2023probabilistic proposed a variant of this approach for forecasting tourism recovery from COVID-19, combining forecast reconciliation and forecast combinations applied to historical data to generate robust COVID-free counterfactual forecasts that set the expected future paths as if the pandemic never occurred.
+Comparing those estimations to produced probabilistic forecasts allowed the tourism sector to estimate the difference between scenario-based probabilistic forecasts and forecasts prepared as if the pandemic had never occurred.
+
+Inventory stockouts are a common cause of data interruption in the supply chain.
+To deal with these interruptions, a popular technique is to provide an estimate for the observation corresponding to the stock-out period as if there were no stock-outs, and then generate projections appropriately.
+@Bell2000 presented an adjustment strategy for stock-out periods that involves smoothing demand volatility and correcting for stockouts using predicted variance conditioned on the observed stockout.
+@trapero2023demand also used the Tobit Kalman filter (TKF) for models presented in a state space framework for forecasting purposes.
+This method can efficiently deal with trends, seasonality, and exogenous influences in censored data, as long as the forecasting model functions within a state space framework.
+
+Certain models offer flexibility in adjusting to changes in time series characteristics over time.
+Exponential Smoothing State Space models (ETS) represent an adaptive method that has demonstrated competitiveness in time series forecasting [@gardner2006exponential].
+These approaches allow for the adaptation of model parameters over time to accommodate shifts in time series characteristics [@hyndman2002state].
+Similarly, time-varying parameter models exhibit high adaptability [@harvey2006forecasting], enabling forecasting models to better accommodate interruptions in data over time.
+Time-varying parameters are frequently utilized in dynamic forecasting models, such as state-space models or Bayesian structural time series models [@talih2005structural].
+Some models employ approaches to reduce the influence of affected observations during model training by assigning different weights.
+The utilization of weights extends beyond merely including or excluding observations and can also balance the degree of influence each observation has on the forecasting model [@khoshgoftaar2007empirical].
+
+Disruptive events may potentially alter time series by causing gaps, or sometimes observations can be converted into missing data.
+While many time series forecasting models struggle with missing data, several approaches can internally handle missing data [@twala2008good].
+@7157837 suggested a forecasting approach based on the Least Squares Support Vector Machine (LSSVM) that was particularly designed to handle time series forecasting with missing data.
+A different study by @tang2020joint focused on local and global temporal dynamics for multivariate time series forecasting with missing data.
+Their proposed framework uses a memory network to capture global temporal patterns with local data as key components.
+
+Although several approaches exist, the literature lacks recommendations and comparisons of alternative strategies for forecasting time series data influenced by interruptions such as the COVID-19 pandemic.
+This paper aims to close this gap by proposing several methodologies for forecasting time series influenced by such disruptions.
+
 # Handling interruptions when forecasting {#sec-methods}
 
-We consider several possible ways to handle interruptions when forecasting.
+In this section, we describe several possible strategies to handle interruptions when forecasting time series data.
 
@@
 
-Highly adaptive models can adjust to the interruption as it is happening, and will therefore be able to approximate the data generating process relatively well.
+Highly adaptive models can adjust to the interruption as it is happening and will therefore be able to approximate the data generating process relatively well.
 For example, an ETS model with large smoothing parameters will be able to adjust to the interruption relatively quickly.
@@
 
-However, the prediction intervals will be large, because the model will have heavily discounted past data.
+However, the prediction intervals will be large because the model will have heavily discounted past data.
 In fact, the model will largely forget the past data other than the most recent observations, so there is no memory of the seasonal patterns and other dynamics that were present before the interruption.
@@
 A dynamic regression model with intervention covariates can be used to model the interruption explicitly.
-For example, if the intervention involves a simple level shift, with a reverse level shift at the end of the intervention, we can use a dummy variable to indicate the interruption period, and allow the model to adjust to the interruption.
+For example, if the intervention involves a simple level shift with a reverse level shift at the end of the intervention, we can use a dummy variable to indicate the interruption period and allow the model to adjust to the interruption.
 More complicated interventions can be handled by using more covariates.
@@
 This has the advantage of retaining the memory of the past, and so the seasonal patterns and other dynamics will be retained.
-This allows the change period to be effectively modelled provided the intervention variables are chosen well.
-However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be be too narrow, especially if there is a lasting effect beyond the end of the interruption.
+This allows the change period to be effectively modelled, provided the intervention variables are chosen well.
+However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.
 
@@
 
-Because no information is retained during the disruption, the prediction intervals will become large during the disruption, and after the disruption they will remain large until the model has enough data to estimate the forecast distribution more accurately.
+Because no information is retained during the disruption, the prediction intervals will become large during the disruption, and after the disruption, they will remain large until the model has enough data to estimate the forecast distribution more accurately.
 
@@
 
-A fourth solution is to estimate what might have been during the period of disruption, and then use the adjusted data to fit a model.
+A fourth solution is to estimate what might have been during the period of disruption and then use the adjusted data to fit a model.
 The estimates could be made using any convenient method.
 One approach would be to set the estimates to equal the forecasts made using only pre-interruption data.
-This then becomes almost equal to the previous solution except that the prediction intervals will be narrower because the estimation uncertainty has not been taken into account.
+This then becomes almost equal to the previous solution, except that the prediction intervals will be narrower because the estimation uncertainty has not been taken into account.
 On the other hand, it is more flexible than the previous approach because models that do not handle missing values can then be used.
 
-An alternative would be to take the model estimated under the previous solution (setting the observations during the disruption to missing), and use it to estimate what might have been during the disruption.
+An alternative would be to take the model estimated under the previous solution (setting the observations during the disruption to missing) and use it to estimate what might have been during the disruption.
 Then the model can be re-estimated using the adjusted data.
-Note that forecasts obtained in this way during the disruption will not be true forecasts, because they will have used data from the future in computing the adjusted data during the disruption.
-But post-disruption forecasts will be true forecasts, and should be almost the same as those obtained using the previous solution.
+Note that forecasts obtained in this way during the disruption will not be true forecasts because they will have used data from the future in computing the adjusted data during the disruption.
+But post-disruption forecasts will be true forecasts and should be almost the same as those obtained using the previous solution.
 
@@
 
-If the weights during the disruption are set to zero, and the weights at other times are set to 1, this becomes equivalent to the missing value solution above.
-Rather than ignoring the data during the disruption, the weighting approach allows the model to not be completely blind to the disruption's effects, and adapt to changes or shifts in the time series caused by the event.
+If the weights during the disruption are set to zero and the weights at other times are set to 1, this becomes equivalent to the missing value solution above.
+Rather than ignoring the data during the disruption, the weighting approach allows the model to not be completely blind to the disruption's effects and adapt to changes or shifts in the time series caused by the event.
 
-While conceptually simple to implement, in practice, this approach may require more work than the other methods discussed here as most forecasting software does not allow for the explicit use of weights.
+While conceptually simple to implement, in practice, this approach may require more work than the other methods discussed here, as most forecasting software does not allow for the explicit use of weights.
 
@@
 
-As with may forecasting problems, taken an ensemble approach often leads to more accurate forecasts [@combinations].
+As with many forecasting problems, taking an ensemble approach often leads to more accurate forecasts [@combinations].
 Here, we could combine some or all four of the approaches discussed above to obtain an ensemble forecast.
-In fact, if we were unable to implement the weighted approach, by using an ensemble of the other approaches we are effectively downweighting the observations during the disruption period as they are only explicitly used in the first two approaches (using a highly adaptive model, or using an intervention model).
+In fact, if we were unable to implement the weighted approach, by using an ensemble of the other approaches, we are effectively downweighting the observations during the disruption period, as they are only explicitly used in the first two approaches (using a highly adaptive model or using an intervention model).
 
-However, a disadvantage of this approach is that it is averaging forecasts that aim to achieve different objectives.
+However, a disadvantage of this approach is that it is based on averaging forecasts that aim to achieve different objectives.
 For example, the highly adaptive model and the intervention model are trying to forecast what happened during the disruption, while the missing value and estimation approaches are trying to forecast what might have happened.
@@
 
-Let's consider some examples to data observed during the last few years, where the effect of the COVID-19 pandemic has been particularly evident.
+In this section, we implement and evaluate the strategies described in @sec-methods using two daily and monthly datasets collected over the last several years, where the impact of the COVID-19 pandemic has been particularly evident.
 
@@
 The data [@tourismdata] are available from January 2000 to May 2023, and are shown in @fig-tourism-plot1.
-As the borders closed in March 2020, the number of visitors to Australia dropped to near zero, and remained there until towards the end of 2021.
-The borders officially reopened on 21 February 2022, although it seems visitors began to arrive earlier than that.
+As the borders closed in March 2020, the number of visitors to Australia dropped to near zero and remained there until towards the end of 2021.
+The borders officially reopened on February 21, 2022, although it seems visitors began to arrive earlier than that.
 
@@
 We apply the first four solutions discussed in the previous section to these data, making 12 month forecasts at the end of each year from 2019 to 2022.
-We fit ETS, ARIMA and dynamic ARIMA models to the data [@fpp3], first applying a log transformation to ensure the resulting forecasts are positive.
-This is possible, because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> filter(Visitors == min(Visitors)) |> pull(Month) |> format("%B %Y")`.
+We fit ETS, ARIMA, and dynamic ARIMA models to the data [@fpp3], first applying a log transformation to ensure the resulting forecasts are positive.
+This is possible because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> filter(Visitors == min(Visitors)) |> pull(Month) |> format("%B %Y")`.
 For all forecasts, we also show 90% prediction intervals.
@@
 ETS, in particular, is well-known to be relatively adaptive to changes in the series, and this is evident in these forecasts.
-The forecasts for 2020 were made using data before COVID-19 had any affect, and so they show similar patterns to the past.
-The forecasts for 2021 were made after 9 months of very low levels of arrivals, and both models show forecasts consistent with the recent history.
+The forecasts for 2020 were made using data before COVID-19 had any effect, and so they show similar patterns to the past.
+The forecasts for 2021 were made after 9 months of very low levels of arrivals, and both models show forecasts consistent with recent history.
 The use of logarithms is particularly important here as the variance is much smaller during 2020 than previously, but after taking logarithms, the variance is more stable over time.
-The forecasts made at the start of 2022 have struggled to detect the small increase in traffic at the end of 2021, and both models have forecast relatively flat trajectories as a result, although the prediction intervals are wide indicating model uncertainty.
-Finally, the forecasts made at the end of 2022 have captured the increasing trend, but ETS is much closer to the reality, adapting more quickly to the changing patterns.
+The forecasts made at the start of 2022 have struggled to detect the small increase in traffic at the end of 2021, and both models have forecast relatively flat trajectories as a result, although the prediction intervals are wide, indicating model uncertainty.
+Finally, the forecasts made at the end of 2022 have captured the increasing trend, but ETS is much closer to reality, adapting more quickly to the changing patterns.
 Again, the wide prediction intervals indicate a high level of uncertainty.
 It is possible to make ETS more adaptive to changes in the data by increasing the value of the smoothing parameters.
-For example, a high value of $\beta$ (the smoothing parameter for the slope), will result in changes of trend being incorporated into the forecasts more quickly, at the risk of over-reacting to noise in the data, and increasing the size of the prediction intervals even more.
+For example, a high value of $\beta$ (the smoothing parameter for the slope) will result in changes in trend being incorporated into the forecasts more quickly, at the risk of overreacting to noise in the data, and increasing the size of the prediction intervals even more.
 
@@
 It is evident that the level shift variable has worked well, giving relatively good forecasts for 2021.
-However, the forecasts for 2022 are particularly poor, because the ramp slope has been greatly overestimated, as it was based on only three observations (Oct -- Dec 2021).
+However, the forecasts for 2022 are particularly poor because the ramp slope has been greatly overestimated, as it was based on only three observations (Oct -- Dec 2021).
 The forecasts for 2023 are much better, and the relatively large prediction intervals are appropriate given the uncertainty in the industry at the end of 2022.
@@
 
-The third solution (@fig-tsol3-plot) involved setting the observations during the disruption period to missing, and then fitting an ARIMA model to the series.
+The third solution (@fig-tsol3-plot) involved setting the observations during the disruption period to missing and then fitting an ARIMA model to the series.
 Consequently, the first three years of forecasts show what might have been without the COVID-19 pandemic, based on the history to the end of 2019.
-The final forecasts for 2023 use the data to the end of 2019, and the two observations in November and December 2022.
+The final forecasts for 2023 use the data to the end of 2019 and the two observations in November and December 2022.
 These are much better, but the prediction intervals are too narrow, as there hasn't been sufficient recent data.
@@
 This time, no prediction intervals have been computed.
-The first set of forecasts have not taken any account of the disruption, and so are almost identical to those obtained in the previous plots.
-The last set of forecasts are reasonably good, showing what is obtained by combining the forecasting methods used to handle the disruption period.
+The first set of forecasts has not taken any account of the disruption, and so they are almost identical to those obtained in the previous plots.
+The last set of forecasts is reasonably good, showing what is obtained by combining the forecasting methods used to handle the disruption period.
 However, the forecasts for 2021 and 2022 are particularly poor, because they are averaging forecasts that aim to achieve different objectives, and are affected by the poor forecasts obtained using the dynamic regression model.
@@
 The pedestrian data set is a daily time series showing the number of pedestrians per day in Melbourne, Australia.
-The data were obtained from the Pedestrian Counting System maintained by the City of Melbourne [@pedestrians], and are based on automated hourly counts from sensors at `r nsensors` locations over the period from `r min(raw_walkers$Date)` to `r max(raw_walkers$Date)`.
+The data were obtained from the Pedestrian Counting System maintained by the City of Melbourne [@pedestrians] and are based on automated hourly counts from sensors at `r nsensors` locations over the period from `r min(raw_walkers$Date)` to `r max(raw_walkers$Date)`.
 All sensors had some missing data, and we chose to use sensors with no more than `r 30*24` missing hours, equivalent to 30 days of missing data.
@@
 In Melbourne, there were six separate lockdowns ranging from `r as.numeric(min(lockdowns$Duration))` days to `r as.numeric(max(lockdowns$Duration))` days in length.
-The first official lockdown period was from 31 March 2020 to `r format(min(lockdowns$End), "%d %b %Y")`, and the last lockdown was from `r format(max(lockdowns$Start), "%d %b %Y")` to `r format(max(lockdowns$End), "%d %b %Y")`.
+The first official lockdown period was from March 31, 2020, to `r format(min(lockdowns$End), "%d %b %Y")`, and the last lockdown was from `r format(max(lockdowns$Start), "%d %b %Y")` to `r format(max(lockdowns$End), "%d %b %Y")`.
 However, the first period was preceded by over a week when most people elected to stay and work at home, so we have chosen to start the first period on `r format(min(lockdowns$Start), "%d %b %Y")`, to better reflect human behaviour.
@@
 We evaluate the solutions discussed in @sec-methods by using time series cross-validation, with the initial training set comprising the whole of 2019, and subsequent training sets growing by 1 week at a time, to the end of 2021.
-The test sets are always 1 week long.
-Thus, we have evaluated the forecasts during 2020 and 2021, covering all lockdown periods, and the start of the recovery period.
+The test sets are always one week long.
+Thus, we have evaluated the forecasts for 2020 and 2021, covering all lockdown periods and the start of the recovery period.
 No transformations of the data have been used.
@@
 
-In @fig-walkers-plot2, we have used an intervention model, with a dummy variable indicating the lockdown periods.
+In @fig-walkers-plot2, we have used an intervention model with a dummy variable indicating the lockdown periods.
 This has worked a little better than the ARIMA model used in @fig-walkers-plot1, apart from around the first lockdown period.
-Many people were in self-imposed lockdown before the start of the first lockdown period, and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
+Many people were in self-imposed lockdown before the start of the first lockdown period and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
 
@@
 
-@fig-walkers-plot3 shows the results obtained by setting the observations during the lockdown periods to missing, and then fitting an ARIMA model to the remaining data.
+@fig-walkers-plot3 shows the results obtained by setting the observations during the lockdown periods to missing and then fitting an ARIMA model to the remaining data.
 The forecasts during lockdowns show what might have been if that particular lockdown hadn't occurred, based on previous non-lockdown data.
@@
 Then a new ARIMA model is fitted to the whole data set, and used to produce forecasts.
-As a result, the "forecasts" appear to interpolate across the lockdown periods, reflecting neither the true lockdown pattern, or the pattern that might have occurred without the lockdowns.
+As a result, the "forecasts" appear to interpolate across the lockdown periods, reflecting neither the true lockdown pattern, nor the pattern that might have occurred without the lockdowns.
 
@@
 However, if the intervention is complex, then a highly adaptive method is often better.
-The highly adaptive model is also useful when the timing of the interruption is unknown, or when the post-interruption period is expected to be very different from the pre-interruption period.
-The missing value approach is particularly useful when only post-interruption forecasts are required, and forecasts during the interruption period are not needed.
+The highly adaptive model is also useful when the timing of the interruption is unknown or when the post-interruption period is expected to be very different from the pre-interruption period.
+The missing value approach is particularly useful when only post-interruption forecasts are required and forecasts during the interruption period are not needed.
 
 We have not previously used the weighted approach in practice, and software to implement it is not currently available.
-However, it seems to be a good compromise between the highly adaptive model and the intervention model, and we expect it to be useful in many situations, once software is available to implement it easily.
+However, it seems to be a good compromise between the highly adaptive model and the intervention model, and we expect it to be useful in many situations once software is available to implement it easily.
 
-Finally, the ensemble approach is useful when there is uncertainty about which approach to use, and will often result in greater accuracy due the power of averaging.
+Finally, the ensemble approach is useful when there is uncertainty about which approach to use, and will often result in greater accuracy due to the power of averaging.
 
@@
 
+# Conclusion
+
+It is not uncommon for modellers to experience interruptions in time series data.
+Whether you're in charge of developing forecasting models or a manager who relies on forecasts to inform decisions and policies, you've undoubtedly recently faced the issue of effectively forecasting time series data influenced by the COVID-19 pandemic.
+While this can stand as a critical challenge for many forecasting teams, it is important to recognise that similar challenges could arise in circumstances other than COVID-19.
+Time series can be disrupted by various factors, such as natural disasters, policy changes, price fluctuations, definition changes, sensor failures in the IoT (Internet of Things), maintenance, and breakdowns in factories,  just to mention a few.
+
+The aim of this study is not to compile a comprehensive list of forecasting models applicable to interrupted time series data.
+Rather, we intend to propose overarching strategies to address such situations, with each strategy allowing for a variety of forecasting methodologies.
+As we provide publicly accessible R codes, readers are encouraged to adapt the study by employing other forecasting methods tailored to their specific contexts.
+
+The proposed strategies to forecast interrupted time series offer various advantages and disadvantages.
+Highly adaptive models offer swift adjustments to interruptions with ease of implementation and speedy computation.
+However, this simplicity comes at the expense of wider prediction intervals and the loss of historical pattern memory, limiting their effectiveness in capturing pre-interruption dynamics.
+Intervention models explicitly account for disruptions, retaining past patterns, yet may yield overly narrow prediction intervals by assuming post-interruption similarity without proper calibration.
+Setting interrupted periods to missing values preserves data integrity but may result in inaccurate forecasts during interruptions.
+Estimating what might have been provides flexibility but may underestimate uncertainty.
+Downweighting the interruption period balances pre- and post-disruption data but may necessitate additional implementation complexity.
+Ensemble models combine various approaches for enhanced accuracy but may introduce challenges in interpretation.
+
+When determining a suitable approach to forecasting interrupted time series, several factors should be considered.
+To begin, the complexity of the interruption itself is critical to approach selection.
+Intervention models are a suitable approach for interruptions with known features or intricate patterns since they can represent the disruption.
+In contrast, highly adaptable models give a simpler option for fast adaptations to interruptions without explicit modelling, which is especially useful when the time or effect of the interruption is unknown.
+Balancing simplicity and flexibility is critical to ensuring that the selected technique is consistent with the features of the interruption and the required forecasting accuracy.
+Furthermore, the width of prediction intervals should be carefully assessed to determine forecast uncertainty, particularly during and after interruptions.
+Finally, for greater robustness and accuracy, investigating ensemble models that integrate various forecasting methods might be beneficial.
+
+Future research might focus on enhancing probabilistic forecasting methods that are designed specifically for interrupted time series.
+Large-scale empirical research evaluating the effectiveness of different strategies in real-world settings might also provide useful insights.
+Comparative studies that assess the accuracy, robustness, and computing efficiency of different techniques across several datasets and disruption situations can help discover the most effective approaches for a range of practical applications.
+
+# Reproducibility {.unnumbered}
+
+To enhance reproducibility and facilitate the adoption of proposed strategies, we provide the data and R codes, as well as the entire paper written in R using [Quarto](https://quarto.org/) and the targets R package [@targetsr2021].
+All materials required to reproduce this paper will be accessible via a public GitHub repository once the paper is accepted for publication.
+
 # References

### 2024-02-18 fits.qmd
Updated quarto template

@@
 author:
-- familyname: Hyndman
-  othernames: Rob J
-  address:
-    - Department of Econometrics & Business Statistics
-    - Monash University
-    - Clayton VIC 3800
-    - Australia
+- name: Rob J Hyndman
+  affiliations:
+    - department: Department of Econometrics & Business Statistics
+    - name: Monash University
+    - city: Clayton VIC
+    - country: Australia
+    - postal-code: 3800
   email: Rob.Hyndman@monash.edu
-  correspondingauthor: true
-- familyname: Rostami-Tabar
-  othernames: Bahman
-  address:
-    - Cardiff Business School
-    - Wales CF10 3EU
-    - United Kingdom
+  corresponding: true
+- name: Bahman Rostami-Tabar
+  affiliations:
+    - name: Cardiff Business School
+    - postal-code: Wales CF10 3EU
+    - country: United Kingdom
   email: rostami-tabarb@cardiff.ac.uk
@@
 keywords: "Forecasting, interrupted time series, disruptive events, COVID-19"
+bibliography: references.bib
 wpnumber: no/yr
@@
 cover: false
-toc: false
-bibliography: references.bib
-biblio-style: authoryear-comp
+linestretch: 1.5
 format:
@@
     include-in-header: header.tex
-cite-method: biblatex
-number-sections: true
     keep-tex: true
-fig-height: 5
-fig-width: 8
-execute:
-  echo: false
-  warning: false
-  message: false
-  cache: false
 ---

### 2024-02-18 fits.qmd
Polishing the wording

@@
 Outbreaks, epidemics, and pandemics, for example, may have a considerable impact on population behavior, causing discrepancies and interruptions of data in a variety of industries such as healthcare, pharmaceuticals, transportation, retail, tourism, and traffic management.
-Similarly, supply chain disruptions induced by factors like strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps.
+Similarly, supply chain disruptions induced by factors such as strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps.
 Transportation-related data may be interrupted due to equipment failures, such as malfunctions or breakdowns in vehicles, aircraft, or rail systems, resulting in missing or inaccurate data points.
@@
 This is a separate problem from change point detection [@truong2020selective] and anomaly detection [@talagala2020anomaly].
-In the situations we investigate, we know that a change has occurred that causes unusual data points and want to forecast what will happen next.
+In the situations we investigate, we know that a change has occurred which has caused some unusual observations, and we want to forecast what will happen next.
 Change point detection refers to determining *when* a change occurred [@blazquez2021review], and anomaly detection aims to identify unusual observations with the intention of either minimising their value, or paying particular attention to them, and doing a root-cause investigation.
-Our study focus is also slightly differs from *intervention modeling*, also refereed to as *interrupted time series analysis and modelling* [@mcdowall2019interrupted], which is used when we have data about an outcome over time and the aim is to assess the impact of an intervention, policy, or program adopted for the whole population at a specific time point.
-This is a special case of what we are looking at at this paper.
+Our study focus also differs from *intervention modeling*, sometimes referred to as *interrupted time series analysis and modelling* [@mcdowall2019interrupted], which is used when we have data about an outcome over time and the aim is to assess the impact of an intervention, policy, or program implemented at a specific time point.
+We are not trying to measure the impact of the interruption; rather we are trying to produce sensible forecasts during and after the interruption.
 
-To our knowledge, this is the first study to offer strategies for forecasting interrupted time series, such as COVID-19 affected data.
+To our knowledge, this is the first study to describe and compare general strategies for forecasting interrupted time series, such as COVID-19 affected data.
 The study proposes and evaluates multiple forecasting strategies designed to handle interruptions in time series forecasting caused by disruptive events.
 These approaches include a range of models capable of capturing different types of changes, from simple level shifts to more complex alterations in seasonal patterns and trends.
-By comparing the performance of these models on two real data sets affected by COVID-19 pandemic, the paper offers valuable insights into the effectiveness and practical implications of each approach.The paper offers practical guidance for practitioners seeking to improve the accuracy and reliability of time series forecasts in the presence of disruptive events.
-By discussing the advantages and disadvantages of each strategy and providing recommendations based on the findings, the paper equips practitioners with actionable insights for addressing the challenges associated with forecasting during and after disruptive events.
-Furthermore, by adhering to reproducibility principles and providing both data and R codes for the forecasting models, the research not only contributes to transparency but also promotes the generalizability of the suggested strategies to a variety of domains.
+By comparing the performance of these models on two real data sets affected by the COVID-19 pandemic, the paper offers valuable insights into the effectiveness and practical implications of each approach.
+By discussing the advantages and disadvantages of each strategy, and providing recommendations based on the findings, the paper equips practitioners with actionable insights for addressing the challenges associated with forecasting during and after disruptive events.
+Further, by adhering to reproducibility principles and providing both data and R code for the forecasting models, the research contributes to transparency and promotes the generalizability of the suggested strategies to a variety of domains.
 This openness allows for the replication of results, ensuring the reliability of findings, and enhancing the accessibility and utility of study outputs for the larger research community.
@@
 
-One modelling approach that has gained popularity is the intervention model, also known as interrupted time series (ITS) modelling, introduced @box1975intervention.
+One modelling approach that has gained popularity is the intervention model, also known as interrupted time series (ITS) modelling, introduced by @box1975intervention.
 This approach focuses on understanding how and whether outcomes change following the implementation of an intervention, policy, or programme at a specific time point.
@@
 These models can forecast future values by developing models that incorporate a range of components, such as pre-intervention levels, trends, seasonality, covariates related to the intervention time, and other external factors that might affect the outcome variable.
-Another related approach is the multiple regimes model, which is particularly useful in forecasting due to its ability to capture structural changes and transitions observed in interrupted time series data [@Koopg2007].
-
-Statistical models developed to detect changes in time series data and handle unusual observations have also proven beneficial in similar circumstances.
-Methods for detecting changing points in time series data may include simple location shift and scale shift models[@aue2013structural], Bayesian [@barry1993bayesian], and supervised and supervised machine learning [@aminikhanghahi2017survey].
-Anomaly detection seeks to find points or subsequences in time series data that have unexpected or anomalous values.
-Facebook's Prophet is an example of a time series forecasting system that uses changepoint detection and automatically adjusts model parameters to catch the most recent trends in a time series.
+A related approach is the multiple regimes model, which is particularly useful in forecasting due to its ability to capture structural changes and transitions observed in interrupted time series data [@Koopg2007].
 
@@
 
-In the context of forecasting disruptions like the COVID-19 pandemic, Bayesian approaches have been proposed for detecting changing points in outbreaks and forecasting case numbers under counterfactual scenarios [@dehning2020inferring].
+In the context of forecasting disruptions such as the COVID-19 pandemic, Bayesian approaches have been proposed for detecting changing points in outbreaks and forecasting case numbers under counterfactual scenarios [@dehning2020inferring].
 Counterfactual analysis examines what would have happened to the forecast variable of interest if the disruptive event had not occurred.
-In line with this, @athanasopoulos2023probabilistic proposed a variant of this approach for forecasting tourism recovery from COVID-19, combining forecast reconciliation and forecast combinations applied to historical data to generate robust COVID-free counterfactual forecasts that set the expected future paths as if the pandemic never occurred.
-Comparing those estimations to produced probabilistic forecasts allowed the tourism sector to estimate the difference between scenario-based probabilistic forecasts and forecasts prepared as if the pandemic had never occurred.
+@athanasopoulos2023probabilistic proposed a variant of this approach for forecasting tourism recovery from COVID-19, combining forecast reconciliation and forecast combinations applied to historical data, to generate COVID-free counterfactual forecasts of what might have been if the pandemic never occurred.
+Then scenario-based judgemental probabilistic forecasts were compared with the counterfactual forecasts, to better understand the future recovery of the tourism sector from the pandemic.
 
@@
 
-An alternative would be to take the model estimated under the previous solution (setting the observations during the disruption to missing) and use it to estimate what might have been during the disruption.
+An alternative would be to take the model estimated under the previous solution (setting the observations during the disruption to missing), and use it to estimate what might have been during the disruption.
 Then the model can be re-estimated using the adjusted data.
@@
 As the borders closed in March 2020, the number of visitors to Australia dropped to near zero and remained there until towards the end of 2021.
-The borders officially reopened on February 21, 2022, although it seems visitors began to arrive earlier than that.
+The borders officially reopened on 21 February 2022, although it seems visitors began to arrive earlier than that.
 
@@
 In Melbourne, there were six separate lockdowns ranging from `r as.numeric(min(lockdowns$Duration))` days to `r as.numeric(max(lockdowns$Duration))` days in length.
-The first official lockdown period was from March 31, 2020, to `r format(min(lockdowns$End), "%d %b %Y")`, and the last lockdown was from `r format(max(lockdowns$Start), "%d %b %Y")` to `r format(max(lockdowns$End), "%d %b %Y")`.
+The first official lockdown period was from 31 March 2020, to `r format(min(lockdowns$End), "%d %b %Y")`, and the last lockdown was from `r format(max(lockdowns$Start), "%d %b %Y")` to `r format(max(lockdowns$End), "%d %b %Y")`.
 However, the first period was preceded by over a week when most people elected to stay and work at home, so we have chosen to start the first period on `r format(min(lockdowns$Start), "%d %b %Y")`, to better reflect human behaviour.
@@
 
-We have not previously used the weighted approach in practice, and software to implement it is not currently available.
+We have not previously used the weighted approach in practice, and open-source software to implement it is not currently available.
 However, it seems to be a good compromise between the highly adaptive model and the intervention model, and we expect it to be useful in many situations once software is available to implement it easily.
@@
 It is not uncommon for modellers to experience interruptions in time series data.
-Whether you're in charge of developing forecasting models or a manager who relies on forecasts to inform decisions and policies, you've undoubtedly recently faced the issue of effectively forecasting time series data influenced by the COVID-19 pandemic.
+Both those in charge of developing forecasting models, and managers relying on forecasts to inform decisions and policies, have recently faced the issue of effectively forecasting time series data influenced by the COVID-19 pandemic.
 While this can stand as a critical challenge for many forecasting teams, it is important to recognise that similar challenges could arise in circumstances other than COVID-19.
@@
 Rather, we intend to propose overarching strategies to address such situations, with each strategy allowing for a variety of forecasting methodologies.
-As we provide publicly accessible R codes, readers are encouraged to adapt the study by employing other forecasting methods tailored to their specific contexts.
+As we provide publicly accessible R code, readers are encouraged to adapt the study by employing other forecasting methods tailored to their specific contexts.
 
@@
 To begin, the complexity of the interruption itself is critical to approach selection.
-Intervention models are a suitable approach for interruptions with known features or intricate patterns since they can represent the disruption.
+Intervention models are a suitable approach for interruptions with known features since they can represent the disruption explicitly.
 In contrast, highly adaptable models give a simpler option for fast adaptations to interruptions without explicit modelling, which is especially useful when the time or effect of the interruption is unknown.
 Balancing simplicity and flexibility is critical to ensuring that the selected technique is consistent with the features of the interruption and the required forecasting accuracy.
-Furthermore, the width of prediction intervals should be carefully assessed to determine forecast uncertainty, particularly during and after interruptions.
+Further, the width of prediction intervals should be carefully assessed to determine forecast uncertainty, particularly during and after interruptions.
 Finally, for greater robustness and accuracy, investigating ensemble models that integrate various forecasting methods might be beneficial.
@@
 
-To enhance reproducibility and facilitate the adoption of proposed strategies, we provide the data and R codes, as well as the entire paper written in R using [Quarto](https://quarto.org/) and the targets R package [@targetsr2021].
+To enhance reproducibility and facilitate the adoption of proposed strategies, we provide the data and R code, as well as the entire paper written in R using [Quarto](https://quarto.org/) and the targets package for R [@targetsr2021].
 All materials required to reproduce this paper will be accessible via a public GitHub repository once the paper is accepted for publication.

### 2024-02-18 fits.tex
Recompiled

@@
-\documentclass[11pt,a4paper,]{article}
-\usepackage{lmodern}
-
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  11pt,
+  a4paper,
+]{article}
+
+\usepackage{amsmath,amssymb}
+\usepackage{setspace}
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{
-            pdftitle={Forecasting interrupted time series},
-            pdfkeywords={blah, blah},
-            colorlinks=true,
-            linkcolor=blue,
-            citecolor=Blue,
-            urlcolor=Blue,
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{geometry}
-\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
-\usepackage[style=authoryear-comp,]{biblatex}
-\addbibresource{references.bib}
-\usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
-\usepackage{graphicx,grffile}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[top=2.5cm,bottom=2.5cm,left=2.5cm,right=2.5cm]{geometry}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{2}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-
-% set default figure placement to htbp
+% Set default figure placement to htbp
 \makeatletter
@@
 
-\title{Forecasting interrupted time series}
-
-%% MONASH STUFF
-
-%% CAPTIONS
-\RequirePackage{caption}
-\DeclareCaptionStyle{italic}[justification=centering]
- {labelfont={bf},textfont={it},labelsep=colon}
-\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
-\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
-
-%% FONT
-\RequirePackage{bera}
-\RequirePackage[charter,expert]{mathdesign}
-\RequirePackage[scale=0.9]{sourcecodepro}
-\RequirePackage[lf,t]{FiraSans}
-
-%% HEADERS AND FOOTERS
-\RequirePackage{fancyhdr}
-\pagestyle{fancy}
-\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
-\makeatletter
-\lhead{\textsf{\expandafter{\@title}}}
-\makeatother
-\rhead{}
-\cfoot{}
-\setlength{\headheight}{15pt}
-\renewcommand{\headrulewidth}{0.4pt}
-\renewcommand{\footrulewidth}{0.4pt}
-\fancypagestyle{plain}{%
-\fancyhf{} % clear all header and footer fields
-\fancyfoot[C]{\sffamily\thepage} % except the center
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrulewidth}{0pt}}
-
-%% MATHS
-\RequirePackage{bm,amsmath}
-\allowdisplaybreaks
-
-%% GRAPHICS
-\RequirePackage{graphicx}
-\setcounter{topnumber}{2}
-\setcounter{bottomnumber}{2}
-\setcounter{totalnumber}{4}
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\bottomfraction}{0.85}
-\renewcommand{\textfraction}{0.15}
-\renewcommand{\floatpagefraction}{0.8}
-
-%\RequirePackage[section]{placeins}
-
-%% SECTION TITLES
-\RequirePackage[compact,sf,bf]{titlesec}
-\titleformat{\section}[block]
-  {\fontsize{15}{17}\bfseries\sffamily}
-  {\thesection}
-  {0.4em}{}
-\titleformat{\subsection}[block]
-  {\fontsize{12}{14}\bfseries\sffamily}
-  {\thesubsection}
-  {0.4em}{}
-\titlespacing{\section}{0pt}{*5}{*1}
-\titlespacing{\subsection}{0pt}{*2}{*0.2}
-
-%% TITLE PAGE
-\def\Date{\number\day}
-\def\Month{\ifcase\month\or
- January\or February\or March\or April\or May\or June\or
- July\or August\or September\or October\or November\or December\fi}
-\def\Year{\number\year}
-
-\makeatletter
-\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
-\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
-\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
-\def\nojel{\def\showjel{}}
-\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
-\def\cover{{\sffamily\setcounter{page}{0}
-        \thispagestyle{empty}
-        \placefig{2}{1.5}{width=5cm}{_extensions/numbats/wp/monash2}
-        \placefig{16.9}{1.5}{width=2.1cm}{_extensions/numbats/wp/MBSportrait}
-        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
-        \begin{textblock}{7}(12.7,27.9)\hfill
-        \includegraphics[height=0.7cm]{_extensions/numbats/wp/AACSB}~~~
-        \includegraphics[height=0.7cm]{_extensions/numbats/wp/EQUIS}~~~
-        \includegraphics[height=0.7cm]{_extensions/numbats/wp/AMBA}
-        \end{textblock}
-        \vspace*{2cm}
-        \begin{center}\Large
-        Department of Econometrics and Business Statistics\\[.5cm]
-        \footnotesize http://monash.edu/business/ebs/research/publications
-        \end{center}\vspace{2cm}
-        \begin{center}
-        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
-                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-                \LARGE\@author\end{onehalfspace}
-        }}
-        \end{center}
-        \vfill
-                \begin{center}\Large
-                \Month~\Year\\[1cm]
-                Working Paper \@wp
-        \end{center}\vspace*{2cm}}}
-\def\pageone{{\sffamily\setstretch{1}%
-        \thispagestyle{empty}%
-        \vbox to \textheight{%
-        \raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
-        \vspace{2cm}\par
-        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
-        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
-        \hspace{1cm}\showjel\vss}}}
-\def\blindtitle{{\sffamily
-     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-        }}
-\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}
-
-\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
-\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
-\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
-\let\maketitle\titlepage
-\makeatother
-
-%% SPACING
-\RequirePackage{setspace}
-\spacing{1.5}
-
-%% LINE AND PAGE BREAKING
-\sloppy
-\clubpenalty = 10000
-\widowpenalty = 10000
-\brokenpenalty = 10000
-\RequirePackage{microtype}
-
-%% PARAGRAPH BREAKS
-\setlength{\parskip}{1.4ex}
-\setlength{\parindent}{0em}
-
-%% HYPERLINKS
-\RequirePackage{xcolor} % Needed for links
-\definecolor{darkblue}{rgb}{0,0,.6}
-\RequirePackage{url}
-
-\makeatletter
-\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
-\makeatother
-\hypersetup{
-     citecolor=0 0 0,
-     breaklinks=true,
-     bookmarksopen=true,
-     bookmarksnumbered=true,
-     linkcolor=darkblue,
-     urlcolor=blue,
-     citecolor=darkblue,
-     colorlinks=true}
-
-%% KEYWORDS
-\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}
-
-%% ABSTRACT
-\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
-\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
-  {\end{minipage}}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\usepackage[showonlyrefs]{mathtools}
-\usepackage[no-weekday]{eukdate}
-
-%% BIBLIOGRAPHY
-
-\makeatletter
-\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
-\makeatother
-\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false, date=year}
-
-\DeclareFieldFormat{url}{\texttt{\url{#1}}}
-\DeclareFieldFormat[article]{pages}{#1}
-\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
-\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
-\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
-\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
-\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
-\DeclareFieldFormat[inproceedings]{title}{#1}
-\DeclareFieldFormat{shorthandwidth}{#1}
-% No dot before number of articles
-\usepackage{xpatch}
-\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
-% Remove In: for an article.
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{%
-  \printtext{\bibstring{in}\intitlepunct}}}
-
-\makeatletter
-\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
-\makeatother
-\renewcommand*{\finalnamedelim}{%
-  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
-  \addspace\&\space}
-
-\wp{no/yr}
-\jel{C10,C14,C22}
-
-\RequirePackage[absolute,overlay]{textpos}
-\setlength{\TPHorizModule}{1cm}
-\setlength{\TPVertModule}{1cm}
-\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}
-
-\nocover
-
-\author{Rob J~Hyndman, Bahman~Rostami-Tabar}
-\addresses{\textbf{Rob J Hyndman}\newline
-Department of Econometrics \& Business Statistics\newline
-Monash University\newline
-Clayton VIC 3800\newline
-Australia\newline
-{Email: Rob.Hyndman@monash.edu}\newline Corresponding author\newline\\[0.5cm]
-\textbf{Bahman Rostami-Tabar}\newline
-Cardiff Business School\newline
-Wales CF10 3EU\newline
-United Kingdom\newline
-{Email: rostami-tabarb@cardiff.ac.uk}\\[0.5cm]
-}
-
-\date{\sf\Date~\Month~\Year}
-\makeatletter
- \lfoot{\sf Hyndman, Rostami-Tabar: \@date}
-\makeatother
-
 \usepackage[section]{placeins}
+\usepackage{orcidlink}
+\definecolor{mypink}{RGB}{219, 48, 122}
 \makeatletter
@@
 \makeatother
-
-% Adjust headwidth in case user has changed geometry in header-includes
-\renewcommand{\headwidth}{\textwidth}
-
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage[style=authoryear-comp,]{biblatex}
+\addbibresource{references.bib}
+\usepackage{bookmark}
+
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Forecasting interrupted time series},
+  pdfauthor={Rob J Hyndman; Bahman Rostami-Tabar},
+  pdfkeywords={Forecasting, interrupted time series, disruptive events,
+COVID-19},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
+
+%% CAPTIONS
+\usepackage{caption}
+\DeclareCaptionStyle{italic}[justification=centering]
+ {labelfont={bf},textfont={it},labelsep=colon}
+\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
+\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
+
+%% FONT
+\usepackage{bera}
+\usepackage[charter]{mathdesign}
+\usepackage[scale=0.9]{sourcecodepro}
+\usepackage[lf,t]{FiraSans}
+
+%% HEADERS AND FOOTERS
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
+\makeatletter
+\lhead{\textsf{\expandafter{\@title}}}
+\makeatother
+\rhead{}
+\cfoot{}
+\setlength{\headheight}{15pt}
+\renewcommand{\headrulewidth}{0.4pt}
+\renewcommand{\footrulewidth}{0.4pt}
+\fancypagestyle{plain}{%
+\fancyhf{} % clear all header and footer fields
+\fancyfoot[C]{\sffamily\thepage} % except the center
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrulewidth}{0pt}}
+
+%% MATHS
+\usepackage{bm,amsmath}
+\allowdisplaybreaks
+
+%% GRAPHICS
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setcounter{topnumber}{2}
+\setcounter{bottomnumber}{2}
+\setcounter{totalnumber}{4}
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\bottomfraction}{0.85}
+\renewcommand{\textfraction}{0.15}
+\renewcommand{\floatpagefraction}{0.8}
+
+%% SECTION TITLES
+\usepackage[compact,sf,bf]{titlesec}
+\titleformat{\section}[block]
+  {\fontsize{15}{17}\bfseries\sffamily}
+  {\thesection}
+  {0.4em}{}
+\titleformat{\subsection}[block]
+  {\fontsize{12}{14}\bfseries\sffamily}
+  {\thesubsection}
+  {0.4em}{}
+\titlespacing{\section}{0pt}{*5}{*1}
+\titlespacing{\subsection}{0pt}{*2}{*0.2}
+
+%% BIBLIOGRAPHY.
+
+\makeatletter
+\@ifpackageloaded{biblatex}{
+\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}
+\DeclareFieldFormat{url}{\texttt{\url{#1}}}
+\DeclareFieldFormat[article]{pages}{#1}
+\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
+\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
+\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
+\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
+\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
+\DeclareFieldFormat[article]{url}{}
+\DeclareFieldFormat[inproceedings]{title}{#1}
+\DeclareFieldFormat{shorthandwidth}{#1}
+\usepackage{xpatch}
+\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
+% Remove In: for an article.
+\renewbibmacro{in:}{%
+  \ifentrytype{article}{}{%
+  \printtext{\bibstring{in}\intitlepunct}}}
+\AtEveryBibitem{\clearfield{month}}
+\AtEveryCitekey{\clearfield{month}}
+\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
+\renewcommand*{\finalnamedelim}{\addspace\&\space}
+}{}
+\makeatother% Placement of logos
+
+\RequirePackage[absolute,overlay]{textpos}
+\setlength{\TPHorizModule}{1cm}
+\setlength{\TPVertModule}{1cm}
+\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}
+
+% Title and date
+
+\title{Forecasting interrupted time series}
+\date{19 February 2024}
+
+\def\Date{\number\day}
+\def\Month{\ifcase\month\or
+ January\or February\or March\or April\or May\or June\or
+ July\or August\or September\or October\or November\or December\fi}
+\def\Year{\number\year}
+
+% Working paper number and JEL codes
+
+\makeatletter
+\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
+\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
+\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
+\def\nojel{\def\showjel{}}
+\makeatother
+
+\wp{no/yr}
+\jel{C10,C14,C22}
+
+% Title page
+
+\makeatletter
+\def\cover{{\sffamily\setcounter{page}{0}
+        \thispagestyle{empty}
+        \placefig{2}{1.5}{width=5cm}{monash2}
+        \placefig{16.9}{1.5}{width=2.1cm}{MBSportrait}
+        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
+        \begin{textblock}{7}(12.7,27.9)\hfill
+        \includegraphics[height=0.7cm]{AACSB}~~~
+        \includegraphics[height=0.7cm]{EQUIS}~~~
+        \includegraphics[height=0.7cm]{AMBA}
+        \end{textblock}
+        \vspace*{2.5cm}
+        \begin{center}\Large
+        Department of Econometrics and Business Statistics\\[.5cm]
+        \footnotesize http://monash.edu/business/ebs/research/publications
+        \end{center}\vspace{2cm}
+        \begin{center}
+        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
+                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
+                \LARGE
+                \expandafter{\@author}
+                \end{onehalfspace}
+        }}
+        \end{center}
+        \vfill
+                \begin{center}\Large
+                \Month~\Year\\[1cm]
+                Working Paper \@wp
+        \end{center}\vspace*{2cm}}}
+        \def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
+        \def\pageone{{\sffamily\setstretch{1}%
+        \thispagestyle{empty}%
+        \vbox to \textheight{%
+        \raggedright\baselineskip=1.2cm
+     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
+        \vspace{2cm}\par
+        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
+        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
+        \hspace{1cm}\showjel\vss}}}
+\def\blindtitle{{\sffamily
+     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
+     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
+        }}
+\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}
+
+\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
+\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
+\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
+\let\maketitle\titlepage
+\makeatother
+
+% Authors
+
+\nocover
+  \author{Rob J Hyndman, Bahman Rostami-Tabar}
+  \addresses{%
+    %
+      \textbf{Rob J Hyndman}\\%
+      %
+        Department of Econometrics \& Business Statistics\\%
+        %
+        %
+        %
+      %
+        %
+        Monash University\\%
+        %
+        %
+      %
+        %
+        %
+        Clayton VIC \\%
+        %
+      %
+        %
+        %
+        %
+        Australia\\%
+      %
+        %
+        %
+        %
+        %
+      %
+      {Email: Rob.Hyndman@monash.edu}\\%
+      \textit{Corresponding author}\\[0.5cm]%
+   %
+      \textbf{Bahman Rostami-Tabar}\\%
+      %
+        %
+        Cardiff Business School\\%
+        %
+        %
+      %
+        %
+        %
+        %
+        %
+      %
+        %
+        %
+        %
+        United Kingdom\\%
+      %
+      {Email: rostami-tabarb@cardiff.ac.uk}\\%
+      \\[0.5cm]%
+   %
+   }%
+   \lfoot{\sf Hyndman, Rostami-Tabar: 19 February 2024}
+
+% Keywords
+
+\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}
+
+% Abstract
+\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
+\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline\setstretch{1.5}}
+  {\end{minipage}}
 \begin{document}
 \maketitle
+
 \begin{abstract}
-A brief summary of our ideas
+Forecasting interrupted time series data is a major challenge for
+forecasting teams, especially in light of events such as the COVID-19
+pandemic. This paper investigates several strategies for dealing with
+interruptions in time series forecasting, including highly adaptable
+models, intervention models, marking interrupted periods as missing,
+forecasting what may have been, downweighting the interruption period,
+and ensemble models. Each approach offers specific advantages and
+disadvantages, such as adaptability, memory retention, data integrity,
+flexibility, and accuracy. We evaluate the effectiveness of these
+strategies using two actual datasets that were interrupted by COVID-19,
+and we provide recommendations for how to handle these interruptions.
+This work contributes to the literature on time series forecasting,
+offering insights for academics and practitioners dealing with
+interrupted data in numerous domains.
 \end{abstract}
+
 \begin{keywords}
-blah, blah
+  Forecasting, interrupted time series, disruptive events, COVID-19; 
+  Forecasting, interrupted time series, disruptive events, COVID-19
 \end{keywords}
 
+\setstretch{1.5}
 \section{Introduction}\label{introduction}
 
-Time series are sometimes interrupted by unusual events; for example, a
-natural disaster may occur, or a there may be a temporary policy change.
-Many forecasters have faced this issue recently with the COVID-19
-pandemic, where historical patterns have been severely disrupted due to
-lockdowns and other restrictions. In this paper, we consider the problem
-of forecasting during such an event, and after it has occurred.
-
-This is a different problem from change point detection. In the
-situations we consider, we know that a change has occurred, and we want
-to forecast the future after the change. Change point detection is about
-identifying \emph{when} a change has occurred.
-
-Time series models usually assume that the data evolve in the future in
-a similar way to how they have evolved in the past. But a big event
-(such as COVID-19) can result in a future that is different from the
-past, at least in the short term.
-
-The disruption may result in relatively simple changes in the series;
+Time series forecasting models use historical data to estimate future
+values \autocite{fildes2008forecasting} based on consistent patterns
+observed in the past. However, time series are sometimes disrupted by
+unusual events that jeopardise the regularity of the time series
+pattern. This can be defined as an \emph{interrupted time series}. Time
+series data may be disrupted by a variety of reasons, ranging from
+temporary policy changes to natural disasters. Outbreaks, epidemics, and
+pandemics, for example, may have a considerable impact on population
+behavior, causing discrepancies and interruptions of data in a variety
+of industries such as healthcare, pharmaceuticals, transportation,
+retail, tourism, and traffic management. Similarly, supply chain
+disruptions induced by factors such as strikes or shortages cause
+irregularities in sales, inventory levels, production, and delivery,
+often resulting in data gaps. Transportation-related data may be
+interrupted due to equipment failures, such as malfunctions or
+breakdowns in vehicles, aircraft, or rail systems, resulting in missing
+or inaccurate data points. Furthermore, cybersecurity breaches data
+quality by possibly skewing readings in datasets relating to financial
+transactions, network traffic, and user activity, complicating time
+series analysis.
+
+The disruptions may result in relatively simple changes in the series;
 for example a level shift at the start of the disruption, and another at
@@
 the seasonal patterns, and changes to the level, which evolve over time.
-In this paper, we consider a range of models that can be used to handle
-such changes, and compare their performance on some real data sets.
-
-We introduce several approaches to handling interruptions when
+
+The presence of such disrupted events poses a significant challenge to
+time series forecasting approaches, restricting their ability to
+reliably capture systematic information and forecast future values. The
+implications of these temporal interruptions are far-reaching: time
+series forecasting techniques, which are a key component of the
+forecaster's toolbox, may no longer be viable options since time series
+models usually assume that the data will evolve in the future in a
+similar way to how it has evolved in the past. But a big event (such as
+COVID-19) can result in a future that is different from the past, at
+least in the short term. Therefore, practitioners face a critical
+question: how to effectively forecast time series data that are impacted
+by such disruptive events? Many forecasters have faced this issue
+recently with the COVID-19 pandemic, where historical patterns have been
+severely disrupted due to policy interventions, lockdowns, and other
+restrictions. This study is motivated by numerous conversations with
+academics and practitioners, which demonstrate widespread concern about
+dealing with data affected by events like the COVID-19 pandemic. In this
+paper, we consider a range of models that can be used to forecast time
+series influenced by disrupted events, and compare their performance on
+some real data sets.
+
+It is important to note that in this study, we consider the problem of
+forecasting interrupted time series both during and after the event.
+This is a separate problem from change point detection
+\autocite{truong2020selective} and anomaly detection
+\autocite{talagala2020anomaly}. In the situations we investigate, we
+know that a change has occurred which has caused some unusual
+observations, and we want to forecast what will happen next. Change
+point detection refers to determining \emph{when} a change occurred
+\autocite{blazquez2021review}, and anomaly detection aims to identify
+unusual observations with the intention of either minimising their
+value, or paying particular attention to them, and doing a root-cause
+investigation. Our study focus also differs from \emph{intervention
+modeling}, sometimes referred to as \emph{interrupted time series
+analysis and modelling} \autocite{mcdowall2019interrupted}, which is
+used when we have data about an outcome over time and the aim is to
+assess the impact of an intervention, policy, or program implemented at
+a specific time point. We are not trying to measure the impact of the
+interruption; rather we are trying to produce sensible forecasts during
+and after the interruption.
+
+To our knowledge, this is the first study to describe and compare
+general strategies for forecasting interrupted time series, such as
+COVID-19 affected data. The study proposes and evaluates multiple
+forecasting strategies designed to handle interruptions in time series
+forecasting caused by disruptive events. These approaches include a
+range of models capable of capturing different types of changes, from
+simple level shifts to more complex alterations in seasonal patterns and
+trends. By comparing the performance of these models on two real data
+sets affected by the COVID-19 pandemic, the paper offers valuable
+insights into the effectiveness and practical implications of each
+approach. By discussing the advantages and disadvantages of each
+strategy, and providing recommendations based on the findings, the paper
+equips practitioners with actionable insights for addressing the
+challenges associated with forecasting during and after disruptive
+events. Further, by adhering to reproducibility principles and providing
+both data and R code for the forecasting models, the research
+contributes to transparency and promotes the generalizability of the
+suggested strategies to a variety of domains. This openness allows for
+the replication of results, ensuring the reliability of findings, and
+enhancing the accessibility and utility of study outputs for the larger
+research community.
+
+We introduce several strategies to handling interruptions when
 forecasting in Section~\ref{sec-methods}. We then apply these approaches
@@
 
+\section{Related works}\label{related-works}
+
+The presence of unusual observations in time series analysis is a
+significant concern that has received attention in numerous studies. In
+this section, we provide a brief overview of some of these relevant
+publications.
+
+One modelling approach that has gained popularity is the intervention
+model, also known as interrupted time series (ITS) modelling, introduced
+by \textcite{box1975intervention}. This approach focuses on
+understanding how and whether outcomes change following the
+implementation of an intervention, policy, or programme at a specific
+time point. ITS models provide a robust tool for analysing policy and
+programme evaluation across diverse domains
+\autocite{bernal2017interrupted,mcdowall2019interrupted}. These models
+can forecast future values by developing models that incorporate a range
+of components, such as pre-intervention levels, trends, seasonality,
+covariates related to the intervention time, and other external factors
+that might affect the outcome variable. A related approach is the
+multiple regimes model, which is particularly useful in forecasting due
+to its ability to capture structural changes and transitions observed in
+interrupted time series data \autocite{Koopg2007}.
+
+There are also statistical models that are used to detect changes in the
+time series data and that can model unusual observation and have proven
+effective in similar situations. A range of methods such as simple
+location shift and scale shift models \autocite{aue2013structural},
+Bayesian \autocite{barry1993bayesian}, and supervised and supervised
+machine learning \autocite{aminikhanghahi2017survey}. Facebook's Prophet
+\autocite{taylor2018forecasting} is an example of a time series
+forecasting method that can incorporate changepoint detection
+(automatically or manually), and adjust model parameters to capture the
+newest trends in a time series. Anomaly detection is another related
+area in time series analysis. In the context of anomaly detection, time
+series anomalies are defined as points or subsequences with unexpected
+or abnormal values, and the detection goal is to identify such anomalies
+\autocite{blazquez2021review}.
+
+In the context of forecasting disruptions such as the COVID-19 pandemic,
+Bayesian approaches have been proposed for detecting changing points in
+outbreaks and forecasting case numbers under counterfactual scenarios
+\autocite{dehning2020inferring}. Counterfactual analysis examines what
+would have happened to the forecast variable of interest if the
+disruptive event had not occurred.
+\textcite{athanasopoulos2023probabilistic} proposed a variant of this
+approach for forecasting tourism recovery from COVID-19, combining
+forecast reconciliation and forecast combinations applied to historical
+data, to generate COVID-free counterfactual forecasts of what might have
+been if the pandemic never occurred. Then scenario-based judgemental
+probabilistic forecasts were compared with the counterfactual forecasts,
+to better understand the future recovery of the tourism sector from the
+pandemic.
+
+Inventory stockouts are a common cause of data interruption in the
+supply chain. To deal with these interruptions, a popular technique is
+to provide an estimate for the observation corresponding to the
+stock-out period as if there were no stock-outs, and then generate
+projections appropriately. \textcite{Bell2000} presented an adjustment
+strategy for stock-out periods that involves smoothing demand volatility
+and correcting for stockouts using predicted variance conditioned on the
+observed stockout. \textcite{trapero2023demand} also used the Tobit
+Kalman filter (TKF) for models presented in a state space framework for
+forecasting purposes. This method can efficiently deal with trends,
+seasonality, and exogenous influences in censored data, as long as the
+forecasting model functions within a state space framework.
+
+Certain models offer flexibility in adjusting to changes in time series
+characteristics over time. Exponential Smoothing State Space models
+(ETS) represent an adaptive method that has demonstrated competitiveness
+in time series forecasting \autocite{gardner2006exponential}. These
+approaches allow for the adaptation of model parameters over time to
+accommodate shifts in time series characteristics
+\autocite{hyndman2002state}. Similarly, time-varying parameter models
+exhibit high adaptability \autocite{harvey2006forecasting}, enabling
+forecasting models to better accommodate interruptions in data over
+time. Time-varying parameters are frequently utilized in dynamic
+forecasting models, such as state-space models or Bayesian structural
+time series models \autocite{talih2005structural}. Some models employ
+approaches to reduce the influence of affected observations during model
+training by assigning different weights. The utilization of weights
+extends beyond merely including or excluding observations and can also
+balance the degree of influence each observation has on the forecasting
+model \autocite{khoshgoftaar2007empirical}.
+
+Disruptive events may potentially alter time series by causing gaps, or
+sometimes observations can be converted into missing data. While many
+time series forecasting models struggle with missing data, several
+approaches can internally handle missing data \autocite{twala2008good}.
+\textcite{7157837} suggested a forecasting approach based on the Least
+Squares Support Vector Machine (LSSVM) that was particularly designed to
+handle time series forecasting with missing data. A different study by
+\textcite{tang2020joint} focused on local and global temporal dynamics
+for multivariate time series forecasting with missing data. Their
+proposed framework uses a memory network to capture global temporal
+patterns with local data as key components.
+
+Although several approaches exist, the literature lacks recommendations
+and comparisons of alternative strategies for forecasting time series
+data influenced by interruptions such as the COVID-19 pandemic. This
+paper aims to close this gap by proposing several methodologies for
+forecasting time series influenced by such disruptions.
+
 \section{Handling interruptions when forecasting}\label{sec-methods}
 
-We consider several possible ways to handle interruptions when
-forecasting.
+In this section, we describe several possible strategies to handle
+interruptions when forecasting time series data.
 
@@
 
-Highly adaptive models can adjust to the interruption as it is
-happening, and will therefore be able to approximate the data generating
-process relatively well. For example, an ETS model with large smoothing
+Highly adaptive models can adjust to the interruption as it is happening
+and will therefore be able to approximate the data generating process
+relatively well. For example, an ETS model with large smoothing
 parameters will be able to adjust to the interruption relatively
@@
 
-However, the prediction intervals will be large, because the model will
+However, the prediction intervals will be large because the model will
 have heavily discounted past data. In fact, the model will largely
@@
 model the interruption explicitly. For example, if the intervention
-involves a simple level shift, with a reverse level shift at the end of
+involves a simple level shift with a reverse level shift at the end of
 the intervention, we can use a dummy variable to indicate the
-interruption period, and allow the model to adjust to the interruption.
+interruption period and allow the model to adjust to the interruption.
 More complicated interventions can be handled by using more covariates.
@@
 seasonal patterns and other dynamics will be retained. This allows the
-change period to be effectively modelled provided the intervention
+change period to be effectively modelled, provided the intervention
 variables are chosen well. However, the model will assume that the
 post-interruption period will be similar to the pre-interruption period,
-and so the prediction intervals may be be too narrow, especially if
-there is a lasting effect beyond the end of the interruption.
+and so the prediction intervals may be too narrow, especially if there
+is a lasting effect beyond the end of the interruption.
 
@@
 intervals will become large during the disruption, and after the
-disruption they will remain large until the model has enough data to
+disruption, they will remain large until the model has enough data to
 estimate the forecast distribution more accurately.
@@
 A fourth solution is to estimate what might have been during the period
-of disruption, and then use the adjusted data to fit a model. The
+of disruption and then use the adjusted data to fit a model. The
 estimates could be made using any convenient method. One approach would
@@
 pre-interruption data. This then becomes almost equal to the previous
-solution except that the prediction intervals will be narrower because
+solution, except that the prediction intervals will be narrower because
 the estimation uncertainty has not been taken into account. On the other
@@
 forecasts obtained in this way during the disruption will not be true
-forecasts, because they will have used data from the future in computing
+forecasts because they will have used data from the future in computing
 the adjusted data during the disruption. But post-disruption forecasts
-will be true forecasts, and should be almost the same as those obtained
+will be true forecasts and should be almost the same as those obtained
 using the previous solution.
@@
 
-If the weights during the disruption are set to zero, and the weights at
+If the weights during the disruption are set to zero and the weights at
 other times are set to 1, this becomes equivalent to the missing value
@@
 weighting approach allows the model to not be completely blind to the
-disruption's effects, and adapt to changes or shifts in the time series
+disruption's effects and adapt to changes or shifts in the time series
 caused by the event.
@@
 While conceptually simple to implement, in practice, this approach may
-require more work than the other methods discussed here as most
+require more work than the other methods discussed here, as most
 forecasting software does not allow for the explicit use of weights.
@@
 
-As with may forecasting problems, taken an ensemble approach often leads
-to more accurate forecasts \autocite{combinations}. Here, we could
+As with many forecasting problems, taking an ensemble approach often
+leads to more accurate forecasts \autocite{combinations}. Here, we could
 combine some or all four of the approaches discussed above to obtain an
 ensemble forecast. In fact, if we were unable to implement the weighted
-approach, by using an ensemble of the other approaches we are
-effectively downweighting the observations during the disruption period
+approach, by using an ensemble of the other approaches, we are
+effectively downweighting the observations during the disruption period,
 as they are only explicitly used in the first two approaches (using a
-highly adaptive model, or using an intervention model).
-
-However, a disadvantage of this approach is that it is averaging
-forecasts that aim to achieve different objectives. For example, the
-highly adaptive model and the intervention model are trying to forecast
-what happened during the disruption, while the missing value and
-estimation approaches are trying to forecast what might have happened.
-So the ensemble approach will lie somewhere between these two
+highly adaptive model or using an intervention model).
+
+However, a disadvantage of this approach is that it is based on
+averaging forecasts that aim to achieve different objectives. For
+example, the highly adaptive model and the intervention model are trying
+to forecast what happened during the disruption, while the missing value
+and estimation approaches are trying to forecast what might have
+happened. So the ensemble approach will lie somewhere between these two
 objectives.
@@
 
-Let's consider some examples to data observed during the last few years,
-where the effect of the COVID-19 pandemic has been particularly evident.
+In this section, we implement and evaluate the strategies described in
+Section~\ref{sec-methods} using two daily and monthly datasets collected
+over the last several years, where the impact of the COVID-19 pandemic
+has been particularly evident.
 
@@
 are shown in Figure~\ref{fig-tourism-plot1}. As the borders closed in
-March 2020, the number of visitors to Australia dropped to near zero,
-and remained there until towards the end of 2021. The borders officially
+March 2020, the number of visitors to Australia dropped to near zero and
+remained there until towards the end of 2021. The borders officially
 reopened on 21 February 2022, although it seems visitors began to arrive
@@
 these data, making 12 month forecasts at the end of each year from 2019
-to 2022. We fit ETS, ARIMA and dynamic ARIMA models to the data
+to 2022. We fit ETS, ARIMA, and dynamic ARIMA models to the data
 \autocite{fpp3}, first applying a log transformation to ensure the
-resulting forecasts are positive. This is possible, because the
+resulting forecasts are positive. This is possible because the
 observations never reach exactly zero, with the smallest number of
@@
 these forecasts. The forecasts for 2020 were made using data before
-COVID-19 had any affect, and so they show similar patterns to the past.
+COVID-19 had any effect, and so they show similar patterns to the past.
 The forecasts for 2021 were made after 9 months of very low levels of
-arrivals, and both models show forecasts consistent with the recent
-history. The use of logarithms is particularly important here as the
-variance is much smaller during 2020 than previously, but after taking
-logarithms, the variance is more stable over time. The forecasts made at
-the start of 2022 have struggled to detect the small increase in traffic
-at the end of 2021, and both models have forecast relatively flat
-trajectories as a result, although the prediction intervals are wide
-indicating model uncertainty. Finally, the forecasts made at the end of
-2022 have captured the increasing trend, but ETS is much closer to the
-reality, adapting more quickly to the changing patterns. Again, the wide
+arrivals, and both models show forecasts consistent with recent history.
+The use of logarithms is particularly important here as the variance is
+much smaller during 2020 than previously, but after taking logarithms,
+the variance is more stable over time. The forecasts made at the start
+of 2022 have struggled to detect the small increase in traffic at the
+end of 2021, and both models have forecast relatively flat trajectories
+as a result, although the prediction intervals are wide, indicating
+model uncertainty. Finally, the forecasts made at the end of 2022 have
+captured the increasing trend, but ETS is much closer to reality,
+adapting more quickly to the changing patterns. Again, the wide
 prediction intervals indicate a high level of uncertainty. It is
@@
 the value of the smoothing parameters. For example, a high value of
-\(\beta\) (the smoothing parameter for the slope), will result in
-changes of trend being incorporated into the forecasts more quickly, at
-the risk of over-reacting to noise in the data, and increasing the size
-of the prediction intervals even more.
+\(\beta\) (the smoothing parameter for the slope) will result in changes
+in trend being incorporated into the forecasts more quickly, at the risk
+of overreacting to noise in the data, and increasing the size of the
+prediction intervals even more.
 
@@
 level shift variable has worked well, giving relatively good forecasts
-for 2021. However, the forecasts for 2022 are particularly poor, because
+for 2021. However, the forecasts for 2022 are particularly poor because
 the ramp slope has been greatly overestimated, as it was based on only
@@
 The third solution (Figure~\ref{fig-tsol3-plot}) involved setting the
-observations during the disruption period to missing, and then fitting
-an ARIMA model to the series. Consequently, the first three years of
+observations during the disruption period to missing and then fitting an
+ARIMA model to the series. Consequently, the first three years of
 forecasts show what might have been without the COVID-19 pandemic, based
 on the history to the end of 2019. The final forecasts for 2023 use the
-data to the end of 2019, and the two observations in November and
+data to the end of 2019 and the two observations in November and
 December 2022. These are much better, but the prediction intervals are
@@
 This time, no prediction intervals have been computed. The first set of
-forecasts have not taken any account of the disruption, and so are
+forecasts has not taken any account of the disruption, and so they are
 almost identical to those obtained in the previous plots. The last set
-of forecasts are reasonably good, showing what is obtained by combining
+of forecasts is reasonably good, showing what is obtained by combining
 the forecasting methods used to handle the disruption period. However,
@@
 the Pedestrian Counting System maintained by the City of Melbourne
-\autocite{pedestrians}, and are based on automated hourly counts from
+\autocite{pedestrians} and are based on automated hourly counts from
 sensors at 66 locations over the period from 2019-01-01 to 2021-12-31.
@@
 days in length. The first official lockdown period was from 31 March
-2020 to 12 May 2020, and the last lockdown was from 05 Aug 2021 to 21
+2020, to 12 May 2020, and the last lockdown was from 05 Aug 2021 to 21
 Oct 2021. However, the first period was preceded by over a week when
@@
 comprising the whole of 2019, and subsequent training sets growing by 1
-week at a time, to the end of 2021. The test sets are always 1 week
-long. Thus, we have evaluated the forecasts during 2020 and 2021,
-covering all lockdown periods, and the start of the recovery period. No
+week at a time, to the end of 2021. The test sets are always one week
+long. Thus, we have evaluated the forecasts for 2020 and 2021, covering
+all lockdown periods and the start of the recovery period. No
 transformations of the data have been used. The results are shown in
@@
 
-In Figure~\ref{fig-walkers-plot2}, we have used an intervention model,
+In Figure~\ref{fig-walkers-plot2}, we have used an intervention model
 with a dummy variable indicating the lockdown periods. This has worked a
@@
 period. Many people were in self-imposed lockdown before the start of
-the first lockdown period, and were reluctant to return to the city
-after it ended, resulting in poor forecasts on either side of the first
+the first lockdown period and were reluctant to return to the city after
+it ended, resulting in poor forecasts on either side of the first
 lockdown.
@@
 Figure~\ref{fig-walkers-plot3} shows the results obtained by setting the
-observations during the lockdown periods to missing, and then fitting an
+observations during the lockdown periods to missing and then fitting an
 ARIMA model to the remaining data. The forecasts during lockdowns show
@@
 a result, the ``forecasts'' appear to interpolate across the lockdown
-periods, reflecting neither the true lockdown pattern, or the pattern
+periods, reflecting neither the true lockdown pattern, nor the pattern
 that might have occurred without the lockdowns.
@@
 adaptive model is also useful when the timing of the interruption is
-unknown, or when the post-interruption period is expected to be very
+unknown or when the post-interruption period is expected to be very
 different from the pre-interruption period. The missing value approach
 is particularly useful when only post-interruption forecasts are
-required, and forecasts during the interruption period are not needed.
+required and forecasts during the interruption period are not needed.
 
 We have not previously used the weighted approach in practice, and
-software to implement it is not currently available. However, it seems
-to be a good compromise between the highly adaptive model and the
-intervention model, and we expect it to be useful in many situations,
-once software is available to implement it easily.
+open-source software to implement it is not currently available.
+However, it seems to be a good compromise between the highly adaptive
+model and the intervention model, and we expect it to be useful in many
+situations once software is available to implement it easily.
 
 Finally, the ensemble approach is useful when there is uncertainty about
-which approach to use, and will often result in greater accuracy due the
-power of averaging.
+which approach to use, and will often result in greater accuracy due to
+the power of averaging.
 
@@
 
+\section{Conclusion}\label{conclusion}
+
+It is not uncommon for modellers to experience interruptions in time
+series data. Both those in charge of developing forecasting models, and
+managers relying on forecasts to inform decisions and policies, have
+recently faced the issue of effectively forecasting time series data
+influenced by the COVID-19 pandemic. While this can stand as a critical
+challenge for many forecasting teams, it is important to recognise that
+similar challenges could arise in circumstances other than COVID-19.
+Time series can be disrupted by various factors, such as natural
+disasters, policy changes, price fluctuations, definition changes,
+sensor failures in the IoT (Internet of Things), maintenance, and
+breakdowns in factories, just to mention a few.
+
+The aim of this study is not to compile a comprehensive list of
+forecasting models applicable to interrupted time series data. Rather,
+we intend to propose overarching strategies to address such situations,
+with each strategy allowing for a variety of forecasting methodologies.
+As we provide publicly accessible R code, readers are encouraged to
+adapt the study by employing other forecasting methods tailored to their
+specific contexts.
+
+The proposed strategies to forecast interrupted time series offer
+various advantages and disadvantages. Highly adaptive models offer swift
+adjustments to interruptions with ease of implementation and speedy
+computation. However, this simplicity comes at the expense of wider
+prediction intervals and the loss of historical pattern memory, limiting
+their effectiveness in capturing pre-interruption dynamics. Intervention
+models explicitly account for disruptions, retaining past patterns, yet
+may yield overly narrow prediction intervals by assuming
+post-interruption similarity without proper calibration. Setting
+interrupted periods to missing values preserves data integrity but may
+result in inaccurate forecasts during interruptions. Estimating what
+might have been provides flexibility but may underestimate uncertainty.
+Downweighting the interruption period balances pre- and post-disruption
+data but may necessitate additional implementation complexity. Ensemble
+models combine various approaches for enhanced accuracy but may
+introduce challenges in interpretation.
+
+When determining a suitable approach to forecasting interrupted time
+series, several factors should be considered. To begin, the complexity
+of the interruption itself is critical to approach selection.
+Intervention models are a suitable approach for interruptions with known
+features since they can represent the disruption explicitly. In
+contrast, highly adaptable models give a simpler option for fast
+adaptations to interruptions without explicit modelling, which is
+especially useful when the time or effect of the interruption is
+unknown. Balancing simplicity and flexibility is critical to ensuring
+that the selected technique is consistent with the features of the
+interruption and the required forecasting accuracy. Further, the width
+of prediction intervals should be carefully assessed to determine
+forecast uncertainty, particularly during and after interruptions.
+Finally, for greater robustness and accuracy, investigating ensemble
+models that integrate various forecasting methods might be beneficial.
+
+Future research might focus on enhancing probabilistic forecasting
+methods that are designed specifically for interrupted time series.
+Large-scale empirical research evaluating the effectiveness of different
+strategies in real-world settings might also provide useful insights.
+Comparative studies that assess the accuracy, robustness, and computing
+efficiency of different techniques across several datasets and
+disruption situations can help discover the most effective approaches
+for a range of practical applications.
+
+\section*{Reproducibility}\label{reproducibility}
+\addcontentsline{toc}{section}{Reproducibility}
+
+To enhance reproducibility and facilitate the adoption of proposed
+strategies, we provide the data and R code, as well as the entire paper
+written in R using \href{https://quarto.org/}{Quarto} and the targets
+package for R \autocite{targetsr2021}. All materials required to
+reproduce this paper will be accessible via a public GitHub repository
+once the paper is accepted for publication.
+
 \printbibliography[title=References]

### 2024-02-19 fits.qmd
Combined lit review with methods and discussion with conclusions

@@
 
-# Introduction
+# Introduction {#sec-introduction}
 
@@
 
+Although several approaches to the problem exist, the literature lacks recommendations and comparisons of alternative strategies for forecasting time series data influenced by interruptions such as the COVID-19 pandemic.
+This paper aims to close this gap by proposing several methodologies for forecasting time series influenced by such disruptions.
+
 It is important to note that in this study, we consider the problem of forecasting interrupted time series both during and after the event.
-This is a separate problem from change point detection [@truong2020selective] and anomaly detection [@talagala2020anomaly].
+This is a separate problem from change point detection [@aue2013structural;@truong2020selective] and anomaly detection [@talagala2020anomaly].
 In the situations we investigate, we know that a change has occurred which has caused some unusual observations, and we want to forecast what will happen next.
@@
 We introduce several strategies to handling interruptions when forecasting in @sec-methods.
-We then apply these approaches to two real data sets in @sec-sec-examples.
+We then apply these approaches to two real data sets in @sec-examples.
 Finally, in @sec-conclusions, we discuss some of the advantages and disadvantages of each approach, and provide some recommendations for practitioners.
-
-# Related works
-
-The presence of unusual observations in time series analysis is a significant concern that has received attention in numerous studies.
-In this section, we provide a brief overview of some of these relevant publications.
-
-One modelling approach that has gained popularity is the intervention model, also known as interrupted time series (ITS) modelling, introduced by @box1975intervention.
-This approach focuses on understanding how and whether outcomes change following the implementation of an intervention, policy, or programme at a specific time point.
-ITS models provide a robust tool for analysing policy and programme evaluation across diverse domains [@bernal2017interrupted;@mcdowall2019interrupted].
-These models can forecast future values by developing models that incorporate a range of components, such as pre-intervention levels, trends, seasonality, covariates related to the intervention time, and other external factors that might affect the outcome variable.
-A related approach is the multiple regimes model, which is particularly useful in forecasting due to its ability to capture structural changes and transitions observed in interrupted time series data [@Koopg2007].
-
-There are also statistical models that are used to detect changes in the time series data and that can model unusual observation and have proven effective in similar situations.
-A range of methods such as simple location shift and scale shift models [@aue2013structural], Bayesian [@barry1993bayesian], and supervised and supervised machine learning [@aminikhanghahi2017survey].
-Facebook’s Prophet [@taylor2018forecasting] is an example of a time series forecasting method that can incorporate changepoint detection (automatically or manually), and adjust model parameters to capture the newest trends in a time series.
-Anomaly detection is another related area in time series analysis.
-In the context of anomaly detection, time series anomalies are defined as points or subsequences with unexpected or abnormal values, and the detection goal is to identify such anomalies [@blazquez2021review].
-
-In the context of forecasting disruptions such as the COVID-19 pandemic, Bayesian approaches have been proposed for detecting changing points in outbreaks and forecasting case numbers under counterfactual scenarios [@dehning2020inferring].
-Counterfactual analysis examines what would have happened to the forecast variable of interest if the disruptive event had not occurred.
-@athanasopoulos2023probabilistic proposed a variant of this approach for forecasting tourism recovery from COVID-19, combining forecast reconciliation and forecast combinations applied to historical data, to generate COVID-free counterfactual forecasts of what might have been if the pandemic never occurred.
-Then scenario-based judgemental probabilistic forecasts were compared with the counterfactual forecasts, to better understand the future recovery of the tourism sector from the pandemic.
-
-Inventory stockouts are a common cause of data interruption in the supply chain.
-To deal with these interruptions, a popular technique is to provide an estimate for the observation corresponding to the stock-out period as if there were no stock-outs, and then generate projections appropriately.
-@Bell2000 presented an adjustment strategy for stock-out periods that involves smoothing demand volatility and correcting for stockouts using predicted variance conditioned on the observed stockout.
-@trapero2023demand also used the Tobit Kalman filter (TKF) for models presented in a state space framework for forecasting purposes.
-This method can efficiently deal with trends, seasonality, and exogenous influences in censored data, as long as the forecasting model functions within a state space framework.
-
-Certain models offer flexibility in adjusting to changes in time series characteristics over time.
-Exponential Smoothing State Space models (ETS) represent an adaptive method that has demonstrated competitiveness in time series forecasting [@gardner2006exponential].
-These approaches allow for the adaptation of model parameters over time to accommodate shifts in time series characteristics [@hyndman2002state].
-Similarly, time-varying parameter models exhibit high adaptability [@harvey2006forecasting], enabling forecasting models to better accommodate interruptions in data over time.
-Time-varying parameters are frequently utilized in dynamic forecasting models, such as state-space models or Bayesian structural time series models [@talih2005structural].
-Some models employ approaches to reduce the influence of affected observations during model training by assigning different weights.
-The utilization of weights extends beyond merely including or excluding observations and can also balance the degree of influence each observation has on the forecasting model [@khoshgoftaar2007empirical].
-
-Disruptive events may potentially alter time series by causing gaps, or sometimes observations can be converted into missing data.
-While many time series forecasting models struggle with missing data, several approaches can internally handle missing data [@twala2008good].
-@7157837 suggested a forecasting approach based on the Least Squares Support Vector Machine (LSSVM) that was particularly designed to handle time series forecasting with missing data.
-A different study by @tang2020joint focused on local and global temporal dynamics for multivariate time series forecasting with missing data.
-Their proposed framework uses a memory network to capture global temporal patterns with local data as key components.
-
-Although several approaches exist, the literature lacks recommendations and comparisons of alternative strategies for forecasting time series data influenced by interruptions such as the COVID-19 pandemic.
-This paper aims to close this gap by proposing several methodologies for forecasting time series influenced by such disruptions.
 
@@
 
-Highly adaptive models can adjust to the interruption as it is happening and will therefore be able to approximate the data generating process relatively well.
+Highly adaptive models offer flexibility in adjusting to changes in time series characteristics over time.
+These models can adjust to the interruption as it is happening and will therefore be able to approximate the data generating process relatively well.
+
+Exponential Smoothing State Space models (ETS) represent an adaptive method that has demonstrated competitiveness in time series forecasting [@gardner2006exponential].
+These models allow for the adaptation of model parameters over time to accommodate shifts in time series characteristics [@hyndman2002state].
 For example, an ETS model with large smoothing parameters will be able to adjust to the interruption relatively quickly.
@@
 
-However, the prediction intervals will be large because the model will have heavily discounted past data.
+Similarly, time-varying parameter models exhibit high adaptability [@harvey2006forecasting], enabling forecasting models to better accommodate interruptions in data over time.
+Time-varying parameters are frequently utilized in dynamic forecasting models, such as state-space models or Bayesian structural time series models [@talih2005structural].
+Some models employ approaches to reduce the influence of affected observations during model training by assigning different weights.
+The utilization of weights extends beyond merely including or excluding observations and can also balance the degree of influence each observation has on the forecasting model [@khoshgoftaar2007empirical].
+
+The prediction intervals from highly adaptive models tend to be wide because the model will have heavily discounted past data.
 In fact, the model will largely forget the past data other than the most recent observations, so there is no memory of the seasonal patterns and other dynamics that were present before the interruption.
@@
 
-A dynamic regression model with intervention covariates can be used to model the interruption explicitly.
+One modelling approach that has gained popularity in handling disruptions is the intervention model, also known as interrupted time series (ITS) modelling, introduced by @box1975intervention.
+This approach focuses on understanding how and whether outcomes change following the implementation of an intervention, policy, or programme at a specific time point.
+ITS models provide a robust tool for analysing policy and programme evaluation across diverse domains [@bernal2017interrupted;@mcdowall2019interrupted].
+These models can forecast future values by developing models that incorporate a range of components, such as pre-intervention levels, trends, seasonality, covariates related to the intervention time, and other external factors that might affect the outcome variable.
+A related approach is the multiple regimes model, which is particularly useful in forecasting due to its ability to capture structural changes and transitions observed in interrupted time series data [@Koopg2007].
+
+A special case of the intervention models proposed by @box1975intervention is a dynamic regression (or regression with ARIMA errors) model.
+This model includes covariates describing the interruption explicitly, with an ARIMA process describing the remaaining time series dynamics.
 For example, if the intervention involves a simple level shift with a reverse level shift at the end of the intervention, we can use a dummy variable to indicate the interruption period and allow the model to adjust to the interruption.
@@
 
-Many time series models will handle periods of missing values.
-So the problematic observations that occur during the period of disruption can be set to missing, and the model should continue to produce forecasts as if the interruption had not occurred.
+Disruptive events may potentially alter time series by causing gaps.
+Alternatively, problematic observations that occur during the period of disruption can be set to missing, and a model fitted to the remaining observations should continue to produce forecasts as if the interruption had not occurred.
 Of course, the forecasts will not be accurate for the period of disruption, but they can be interpreted as "what might have been".
@@
 
-Because no information is retained during the disruption, the prediction intervals will become large during the disruption, and after the disruption, they will remain large until the model has enough data to estimate the forecast distribution more accurately.
+While many time series forecasting models struggle with missing data, several approaches can internally handle missing data [@twala2008good].
+@7157837 suggested a forecasting approach based on a Least Squares Support Vector Machine (LSSVM) that was particularly designed to handle time series forecasting with missing data.
+A different study by @tang2020joint focused on local and global temporal dynamics for multivariate time series forecasting with missing data.
+Their proposed framework uses a memory network to capture global temporal patterns with local data as key components.
+
+In this approach, no information is retained during the disruption, so the prediction intervals will become large during the disruption, and after the disruption, they will remain large until the model has enough data to estimate the forecast distribution more accurately.
 
@@
 But post-disruption forecasts will be true forecasts and should be almost the same as those obtained using the previous solution.
+
+These approaches are sometimes called "counterfactual" methods, as they examine what would have happened to the forecast variable of interest if the disruptive event had not occurred.
+@athanasopoulos2023probabilistic proposed a variant of this approach for forecasting tourism recovery from COVID-19, combining forecast reconciliation and forecast combinations applied to historical data, to generate COVID-free counterfactual forecasts of what might have been if the pandemic never occurred.
+Then scenario-based judgemental probabilistic forecasts were compared with the counterfactual forecasts, to better understand the future recovery of the tourism sector from the pandemic.
+
+Inventory stockouts are a common cause of data interruption in the supply chain.
+To deal with these interruptions, a counterfactual approach is popular, whereby the demand corresponding to the stock-out period is estimated as if there were no stock-outs, and then forecasts are obtained.
+@Bell2000 presented an adjustment strategy for stock-out periods that involves smoothing demand volatility and correcting for stockouts using predicted variance conditioned on the observed stockout.
+@trapero2023demand also used the Tobit Kalman filter (TKF) for models presented in a state space framework for forecasting purposes.
+This method can efficiently deal with trends, seasonality, and exogenous influences in censored data, as long as the forecasting model functions within a state space framework.
+
+In the context of forecasting disruptions such as the COVID-19 pandemic, Bayesian approaches have also been proposed for detecting changing points in outbreaks and forecasting case numbers under counterfactual scenarios [@dehning2020inferring].
 
@@
 
-# Discussion
+# Discussion and conclusions {#sec-conclusions}
 
@@
 Prediction intervals for ensemble forecasts can be obtained by averaging the end points of the component prediction intervals [@lichtendahl2013better], for those models where the prediction intervals seem reasonable.
-
-# Conclusion
 

### 2024-02-19 fits.tex
Combined lit review with methods and discussion with conclusions

@@
 \setstretch{1.5}
-\section{Introduction}\label{introduction}
+\section{Introduction}\label{sec-introduction}
 
@@
 
+Although several approaches to the problem exist, the literature lacks
+recommendations and comparisons of alternative strategies for
+forecasting time series data influenced by interruptions such as the
+COVID-19 pandemic. This paper aims to close this gap by proposing
+several methodologies for forecasting time series influenced by such
+disruptions.
+
 It is important to note that in this study, we consider the problem of
@@
 This is a separate problem from change point detection
-\autocite{truong2020selective} and anomaly detection
+\autocite{aue2013structural,truong2020selective} and anomaly detection
 \autocite{talagala2020anomaly}. In the situations we investigate, we
@@
 forecasting in Section~\ref{sec-methods}. We then apply these approaches
-to two real data sets in \textbf{?@sec-sec-examples}. Finally, in
-\textbf{?@sec-conclusions}, we discuss some of the advantages and
+to two real data sets in Section~\ref{sec-examples}. Finally, in
+Section~\ref{sec-conclusions}, we discuss some of the advantages and
 disadvantages of each approach, and provide some recommendations for
@@
 
-\section{Related works}\label{related-works}
-
-The presence of unusual observations in time series analysis is a
-significant concern that has received attention in numerous studies. In
-this section, we provide a brief overview of some of these relevant
-publications.
-
-One modelling approach that has gained popularity is the intervention
-model, also known as interrupted time series (ITS) modelling, introduced
-by \textcite{box1975intervention}. This approach focuses on
-understanding how and whether outcomes change following the
-implementation of an intervention, policy, or programme at a specific
-time point. ITS models provide a robust tool for analysing policy and
-programme evaluation across diverse domains
+\section{Handling interruptions when forecasting}\label{sec-methods}
+
+In this section, we describe several possible strategies to handle
+interruptions when forecasting time series data.
+
+\subsection{Use a highly adaptive
+model}\label{use-a-highly-adaptive-model}
+
+Highly adaptive models offer flexibility in adjusting to changes in time
+series characteristics over time. These models can adjust to the
+interruption as it is happening and will therefore be able to
+approximate the data generating process relatively well.
+
+Exponential Smoothing State Space models (ETS) represent an adaptive
+method that has demonstrated competitiveness in time series forecasting
+\autocite{gardner2006exponential}. These models allow for the adaptation
+of model parameters over time to accommodate shifts in time series
+characteristics \autocite{hyndman2002state}. For example, an ETS model
+with large smoothing parameters will be able to adjust to the
+interruption relatively quickly. This has the advantage of being a very
+simple solution that is easy to implement and fast to compute. There is
+no need to explicitly model the interruption, and so the model can be
+used for forecasting even if the timing or effect of the interruption is
+unknown.
+
+Similarly, time-varying parameter models exhibit high adaptability
+\autocite{harvey2006forecasting}, enabling forecasting models to better
+accommodate interruptions in data over time. Time-varying parameters are
+frequently utilized in dynamic forecasting models, such as state-space
+models or Bayesian structural time series models
+\autocite{talih2005structural}. Some models employ approaches to reduce
+the influence of affected observations during model training by
+assigning different weights. The utilization of weights extends beyond
+merely including or excluding observations and can also balance the
+degree of influence each observation has on the forecasting model
+\autocite{khoshgoftaar2007empirical}.
+
+The prediction intervals from highly adaptive models tend to be wide
+because the model will have heavily discounted past data. In fact, the
+model will largely forget the past data other than the most recent
+observations, so there is no memory of the seasonal patterns and other
+dynamics that were present before the interruption. Consequently, the
+approach works best if there is no assumption that the post-interruption
+period will be similar to the pre-interruption period.
+
+\subsection{Use an intervention model}\label{use-an-intervention-model}
+
+One modelling approach that has gained popularity in handling
+disruptions is the intervention model, also known as interrupted time
+series (ITS) modelling, introduced by \textcite{box1975intervention}.
+This approach focuses on understanding how and whether outcomes change
+following the implementation of an intervention, policy, or programme at
+a specific time point. ITS models provide a robust tool for analysing
+policy and programme evaluation across diverse domains
 \autocite{bernal2017interrupted,mcdowall2019interrupted}. These models
@@
 
-There are also statistical models that are used to detect changes in the
-time series data and that can model unusual observation and have proven
-effective in similar situations. A range of methods such as simple
-location shift and scale shift models \autocite{aue2013structural},
-Bayesian \autocite{barry1993bayesian}, and supervised and supervised
-machine learning \autocite{aminikhanghahi2017survey}. Facebook's Prophet
-\autocite{taylor2018forecasting} is an example of a time series
-forecasting method that can incorporate changepoint detection
-(automatically or manually), and adjust model parameters to capture the
-newest trends in a time series. Anomaly detection is another related
-area in time series analysis. In the context of anomaly detection, time
-series anomalies are defined as points or subsequences with unexpected
-or abnormal values, and the detection goal is to identify such anomalies
-\autocite{blazquez2021review}.
-
-In the context of forecasting disruptions such as the COVID-19 pandemic,
-Bayesian approaches have been proposed for detecting changing points in
-outbreaks and forecasting case numbers under counterfactual scenarios
-\autocite{dehning2020inferring}. Counterfactual analysis examines what
-would have happened to the forecast variable of interest if the
-disruptive event had not occurred.
-\textcite{athanasopoulos2023probabilistic} proposed a variant of this
-approach for forecasting tourism recovery from COVID-19, combining
-forecast reconciliation and forecast combinations applied to historical
-data, to generate COVID-free counterfactual forecasts of what might have
-been if the pandemic never occurred. Then scenario-based judgemental
-probabilistic forecasts were compared with the counterfactual forecasts,
-to better understand the future recovery of the tourism sector from the
-pandemic.
-
-Inventory stockouts are a common cause of data interruption in the
-supply chain. To deal with these interruptions, a popular technique is
-to provide an estimate for the observation corresponding to the
-stock-out period as if there were no stock-outs, and then generate
-projections appropriately. \textcite{Bell2000} presented an adjustment
-strategy for stock-out periods that involves smoothing demand volatility
-and correcting for stockouts using predicted variance conditioned on the
-observed stockout. \textcite{trapero2023demand} also used the Tobit
-Kalman filter (TKF) for models presented in a state space framework for
-forecasting purposes. This method can efficiently deal with trends,
-seasonality, and exogenous influences in censored data, as long as the
-forecasting model functions within a state space framework.
-
-Certain models offer flexibility in adjusting to changes in time series
-characteristics over time. Exponential Smoothing State Space models
-(ETS) represent an adaptive method that has demonstrated competitiveness
-in time series forecasting \autocite{gardner2006exponential}. These
-approaches allow for the adaptation of model parameters over time to
-accommodate shifts in time series characteristics
-\autocite{hyndman2002state}. Similarly, time-varying parameter models
-exhibit high adaptability \autocite{harvey2006forecasting}, enabling
-forecasting models to better accommodate interruptions in data over
-time. Time-varying parameters are frequently utilized in dynamic
-forecasting models, such as state-space models or Bayesian structural
-time series models \autocite{talih2005structural}. Some models employ
-approaches to reduce the influence of affected observations during model
-training by assigning different weights. The utilization of weights
-extends beyond merely including or excluding observations and can also
-balance the degree of influence each observation has on the forecasting
-model \autocite{khoshgoftaar2007empirical}.
-
-Disruptive events may potentially alter time series by causing gaps, or
-sometimes observations can be converted into missing data. While many
-time series forecasting models struggle with missing data, several
-approaches can internally handle missing data \autocite{twala2008good}.
-\textcite{7157837} suggested a forecasting approach based on the Least
-Squares Support Vector Machine (LSSVM) that was particularly designed to
-handle time series forecasting with missing data. A different study by
-\textcite{tang2020joint} focused on local and global temporal dynamics
-for multivariate time series forecasting with missing data. Their
-proposed framework uses a memory network to capture global temporal
-patterns with local data as key components.
-
-Although several approaches exist, the literature lacks recommendations
-and comparisons of alternative strategies for forecasting time series
-data influenced by interruptions such as the COVID-19 pandemic. This
-paper aims to close this gap by proposing several methodologies for
-forecasting time series influenced by such disruptions.
-
-\section{Handling interruptions when forecasting}\label{sec-methods}
-
-In this section, we describe several possible strategies to handle
-interruptions when forecasting time series data.
-
-\subsection{Use a highly adaptive
-model}\label{use-a-highly-adaptive-model}
-
-Highly adaptive models can adjust to the interruption as it is happening
-and will therefore be able to approximate the data generating process
-relatively well. For example, an ETS model with large smoothing
-parameters will be able to adjust to the interruption relatively
-quickly. This has the advantage of being a very simple solution that is
-easy to implement and fast to compute. There is no need to explicitly
-model the interruption, and so the model can be used for forecasting
-even if the timing or effect of the interruption is unknown.
-
-However, the prediction intervals will be large because the model will
-have heavily discounted past data. In fact, the model will largely
-forget the past data other than the most recent observations, so there
-is no memory of the seasonal patterns and other dynamics that were
-present before the interruption. Consequently, the approach works best
-if there is no assumption that the post-interruption period will be
-similar to the pre-interruption period.
-
-\subsection{Use an intervention model}\label{use-an-intervention-model}
-
-A dynamic regression model with intervention covariates can be used to
-model the interruption explicitly. For example, if the intervention
-involves a simple level shift with a reverse level shift at the end of
-the intervention, we can use a dummy variable to indicate the
-interruption period and allow the model to adjust to the interruption.
-More complicated interventions can be handled by using more covariates.
+A special case of the intervention models proposed by
+\textcite{box1975intervention} is a dynamic regression (or regression
+with ARIMA errors) model. This model includes covariates describing the
+interruption explicitly, with an ARIMA process describing the remaaining
+time series dynamics. For example, if the intervention involves a simple
+level shift with a reverse level shift at the end of the intervention,
+we can use a dummy variable to indicate the interruption period and
+allow the model to adjust to the interruption. More complicated
+interventions can be handled by using more covariates.
 
@@
 
-Many time series models will handle periods of missing values. So the
-problematic observations that occur during the period of disruption can
-be set to missing, and the model should continue to produce forecasts as
-if the interruption had not occurred. Of course, the forecasts will not
-be accurate for the period of disruption, but they can be interpreted as
-``what might have been''. This solution requires a judgement to be made
-about when the disruption has begun, and when normality resumes.
-
-Because no information is retained during the disruption, the prediction
-intervals will become large during the disruption, and after the
-disruption, they will remain large until the model has enough data to
-estimate the forecast distribution more accurately.
+Disruptive events may potentially alter time series by causing gaps.
+Alternatively, problematic observations that occur during the period of
+disruption can be set to missing, and a model fitted to the remaining
+observations should continue to produce forecasts as if the interruption
+had not occurred. Of course, the forecasts will not be accurate for the
+period of disruption, but they can be interpreted as ``what might have
+been''. This solution requires a judgement to be made about when the
+disruption has begun, and when normality resumes.
+
+While many time series forecasting models struggle with missing data,
+several approaches can internally handle missing data
+\autocite{twala2008good}. \textcite{7157837} suggested a forecasting
+approach based on a Least Squares Support Vector Machine (LSSVM) that
+was particularly designed to handle time series forecasting with missing
+data. A different study by \textcite{tang2020joint} focused on local and
+global temporal dynamics for multivariate time series forecasting with
+missing data. Their proposed framework uses a memory network to capture
+global temporal patterns with local data as key components.
+
+In this approach, no information is retained during the disruption, so
+the prediction intervals will become large during the disruption, and
+after the disruption, they will remain large until the model has enough
+data to estimate the forecast distribution more accurately.
 
@@
 using the previous solution.
+
+These approaches are sometimes called ``counterfactual'' methods, as
+they examine what would have happened to the forecast variable of
+interest if the disruptive event had not occurred.
+\textcite{athanasopoulos2023probabilistic} proposed a variant of this
+approach for forecasting tourism recovery from COVID-19, combining
+forecast reconciliation and forecast combinations applied to historical
+data, to generate COVID-free counterfactual forecasts of what might have
+been if the pandemic never occurred. Then scenario-based judgemental
+probabilistic forecasts were compared with the counterfactual forecasts,
+to better understand the future recovery of the tourism sector from the
+pandemic.
+
+Inventory stockouts are a common cause of data interruption in the
+supply chain. To deal with these interruptions, a counterfactual
+approach is popular, whereby the demand corresponding to the stock-out
+period is estimated as if there were no stock-outs, and then forecasts
+are obtained. \textcite{Bell2000} presented an adjustment strategy for
+stock-out periods that involves smoothing demand volatility and
+correcting for stockouts using predicted variance conditioned on the
+observed stockout. \textcite{trapero2023demand} also used the Tobit
+Kalman filter (TKF) for models presented in a state space framework for
+forecasting purposes. This method can efficiently deal with trends,
+seasonality, and exogenous influences in censored data, as long as the
+forecasting model functions within a state space framework.
+
+In the context of forecasting disruptions such as the COVID-19 pandemic,
+Bayesian approaches have also been proposed for detecting changing
+points in outbreaks and forecasting case numbers under counterfactual
+scenarios \autocite{dehning2020inferring}.
 
@@
 
-\section{Discussion}\label{discussion}
+\section{Discussion and conclusions}\label{sec-conclusions}
 
@@
 those models where the prediction intervals seem reasonable.
-
-\section{Conclusion}\label{conclusion}
 

### 2024-02-19 fits.qmd
Keywords

@@
 This work contributes to the literature on time series forecasting, offering insights for academics and practitioners dealing with interrupted data in numerous domains."
-keywords: "Forecasting, interrupted time series, disruptive events, COVID-19"
+keywords:
+  - counterfactual forecasting
+  - COVID-19
+  - disruptive events
+  - intervention analysis
+  - structural breaks
 bibliography: references.bib

### 2024-02-19 fits.qmd
Make blind

--- 2024-02-19 fits.qmd
@@
 jelcodes: C10,C14,C22
-blind: false
+blind: true
 cover: false

### 2024-02-19 fits.tex
Make blind

@@
   pdfauthor={Rob J Hyndman; Bahman Rostami-Tabar},
-  pdfkeywords={Forecasting, interrupted time series, disruptive events,
-COVID-19},
+  pdfkeywords={counterfactual forecasting, COVID-19, disruptive
+events, intervention analysis, structural breaks},
   colorlinks=true,
@@
 \title{Forecasting interrupted time series}
-\date{19 February 2024}
+\date{20 February 2024}
 
@@
 \nocover
-  \author{Rob J Hyndman, Bahman Rostami-Tabar}
-  \addresses{%
-    %
-      \textbf{Rob J Hyndman}\\%
-      %
-        Department of Econometrics \& Business Statistics\\%
-        %
-        %
-        %
-      %
-        %
-        Monash University\\%
-        %
-        %
-      %
-        %
-        %
-        Clayton VIC \\%
-        %
-      %
-        %
-        %
-        %
-        Australia\\%
-      %
-        %
-        %
-        %
-        %
-      %
-      {Email: Rob.Hyndman@monash.edu}\\%
-      \textit{Corresponding author}\\[0.5cm]%
-   %
-      \textbf{Bahman Rostami-Tabar}\\%
-      %
-        %
-        Cardiff Business School\\%
-        %
-        %
-      %
-        %
-        %
-        %
-        %
-      %
-        %
-        %
-        %
-        United Kingdom\\%
-      %
-      {Email: rostami-tabarb@cardiff.ac.uk}\\%
-      \\[0.5cm]%
-   %
-   }%
-   \lfoot{\sf Hyndman, Rostami-Tabar: 19 February 2024}
+  \blind
+  \lfoot{\sf20 February 2024}
 
@@
 \begin{keywords}
-  Forecasting, interrupted time series, disruptive events, COVID-19; 
-  Forecasting, interrupted time series, disruptive events, COVID-19
+  counterfactual forecasting; COVID-19; disruptive events; intervention
+analysis; 
+  structural breaks.
 \end{keywords}

### 2024-05-15 response_letter/response_letter_JORS.qmd
add response letter

---
title: "Response letter - Manuscript ID JSR-22-468"
bibliography: bibliography.bib
format:
  pdf:
    pdf-engine: pdflatex
    papersize: a4
    geometry: "margin = 2.5cm"
    include-in-header: header.tex
---

Dear Professor

Please find enclosed a revised version of our paper submitted for peer review for the *Journal of Operational Research Society*. Your comments have been very much appreciated and we have tried to suitably modify/amend our paper for those comments to be reflected in the revised version.

We quote below the detailed referees’ comments that are followed by our response. In all cases, we point out, where necessary, the corresponding amendments in the new version of the paper and highlighted them in teal (greenish-blue color).

# Editor comment

_Your manuscript entitled "Forecasting interrupted time series", which you submitted to Journal of the Operational Research Society, has been reviewed. The referee comments are included at the bottom of this letter. The referee(s) would like to see some revisions made to your manuscript before publication. Therefore, I invite you to respond to the referee(s)' comments and revise your manuscript. When you revise your manuscript please highlight the changes you make in the manuscript by using the track changes mode in MS Word or by using bold or coloured text._

Response:

# Reviewer 1

_**General comment:**The paper deals with the prediction of time series when a disruption happens in the data. Different prediction strategies are proposed, simulated and performances are compared. The methodology is well explained, all data sources and codes are available for replication and improvements. Results are clearly exposed, criticized and recommandations are provided. In my opinion, the paper is very good n its present form and I recommend publication._

_Minor comments_

_Comment1. one typo, line 49, page 4 : "remaining"_

Response:

_Comment2. ref Trapejo (2023) is incomplete_

Response:

# Reviewer 2

_Comment 1:The topic is very interesting and very timely, in the wake of the Covid -19 related disruptions in supply chains, changes in consumer preferences, and inflationary environment. The time series patterns and causal relationships that were seen before the pandemic did not apply and whether the patterns observed before the pandemic will continue to be relevant remains to be evaluated on a case by case basis._

Response:

_Comment 2: I like that the authors contrast the topic with others that it may be confused with. But need to check the claim that it is indeed the first to address this problem._

Response:

_Comment 3:The paper is well written, but in terms of contributions, it falls short of the expectations for an original research paper. Improvement avenues are proposing a viable new method that has advantages over these existing methods, or a large-scale empirical research with carefully curated set of actual data representing different interruption patterns from different domains._

Response:

# Reviewer 3

_General comment: I would appreciate the editor for providing me an opportunity to review the article “Forecasting interrupted time series”. In this paper, the authors investigate several strategies for dealing with interruptions in time series forecasting and make some comparisons among these methods. Though the paper is overall interesting, but the empirical design seems to be too simple. I hope these comments may be helpful for the authors._

Response:

_Comment 1. The authors regard the description and comparison of general strategies for forecasting interrupted time series as a major contribution, which in my opinion is quite limited._

Response:

_Comment 2. The patterns of interrupted events may be very different, which makes the suitable approach varies. In this case, it may be better to focus on a certain event (e.g., COVID-19 pandemic) and make comprehensive comparisons._

Response:

_Comment 3. The authors provide some visualizations on the forecasting. This should be appreciated since it is quite straightforward. However, the authors may not expect the readers to judge the optimal strategy from these figures. Instead, the authors should consider some out-of-sample validation and may introduce some metrics or measures to evaluate the performance different strategies._

Response:

_Comment 4. If the authors focus on the interrupted time series for a certain event, some exogeneous variables are encouraged for potential better performance._

Response:

# Reviewer 4

_General comment: The manuscript deals with forecasting interrupted time series for forecasting teams, especially in light of emergency events like COVID-19. The manuscript investigates several strategies for dealing with interruptions in time series forecasting, and their advantages and disadvantages. The authors evaluated two datasets to understand the effectiveness of these strategies. The manuscripts deal with important areas and insight into the strategies._

A few suggestions need to be included such as:
_Comment 1: Different strategies are discussed in section 2 in a few words without giving a proper introduction of the strategies. It makes the reading of the manuscript difficult. The authors are required to give practical examples of the strategies to make it understandable to read._

_Comment 2. No mathematical support is provided as to how the strategies work. And what are the practical implementation of the strategies? Authors are required to provide detailed insight into the strategies i.e. which datasets these strategies are applied?_

_Comment 3. The figures presented in the manuscripts require detailed explanation. Authors are required to explain the given figures in detail one by one._

_Comment 4. Two datasets are used and the strategies are applied to them. However, the descriptive statistics of these data sets missing.. The authors are required to explain the datasets._

_Comment 5. Literature must reflect the relevant studies i.e.
a) OntoCOVID: Ontology for Semantic Modeling of COVID19 Statistical Data
Authors
Shaukat Ali, Shah Khusro, Sajid Anwar, Abrar Ullah
Publication date
2022
Book
Proceedings of International Conference on Information Technology and Applications
Pages
183-194
Publisher
Springer, Singapore
b) COVID-19 patient count prediction using LSTM
M Iqbal, F Al-Obeidat, F Maqbool, S Razzaq, S Anwar, A Tubaishat, ...
IEEE Transactions on Computational Social Systems 8 (4), 974-981
c)Countering malicious URLs in internet of things using a knowledge-based approach and a simulated expert
S Anwar, F Al-Obeidat, A Tubaishat, S Din, A Ahmad, FA Khan, G Jeon, ...
IEEE Internet of Things Journal 7 (5), 4497-4504

# References {#references .unnumbered}

### 2024-05-15 response_letter/response_letter_JORS.qmd
response letter-first go

@@
 
-Response:
+**Response:** Thank you very much for giving us an opportunity to revise the manuscript.
+The valuable feedback provided by the referees.
+We have thoroughly reviewed their comments and made any required changes.
+We have highlighted all changes with coloured text in both the response letter and the manuscript.
 
@@
 
+**Response:** Thank you for your positive feedback on our paper.
+We are pleased to hear that you found the methodology well-explained and the results clearly presented.
+We appreciate your recommendation for publication.
+
 _Minor comments_
@@
 
-Response:
+**Response:** This has been corrected
 
@@
 
-Response:
+**Response:** This paper is currently available online. We have included the designation 'In Press' in the reference to reflect its status.
 
@@
 
-Response:
+**Response:** Thank you for your insightful feedback.
+We are delighted to hear that you appreciate the timeliness and importance of the paper.
+We agree that the topic is particularly relevant given the COVID-19-related disruptions in supply chains and various other areas.
+We also agree that evaluating whether pre-pandemic time series patterns and causal relationships remain applicable is a critical challenge that requires careful case-by-case analysis.
+That is why, in this paper, we propose various strategies rather than a single approach to address different situations and cases effectively.
+
+We have added the following paragraph in the revised manuscript to address this comment:
+
+\textcolor{teal}{The assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis.
+Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
 
@@
 
-Response:
+**Response:** Thank you for your comment.
+The idea for this paper emerged from our close contact with practitioners, who highlighted the challenges they face in forecasting time series data affected by COVID-19, particularly in handling observations impacted by the disruption.
+Following these discussions, we carried out a search to find out if any existing papers addressed this issue.
+As world-leading experts in the field of forecasting and well-connected to the research community, we could not find any published research in this area.
+To our knowledge, this is the first study to describe and compare general strategies for forecasting interrupted time series, such as COVID-19-affected data.
 
@@
 
-Response:
+Response: Thank you for your feedback.
+We appreciate your comments on the writing quality and your recommendations for improving the paper's contributions.
+We believe that the originality of our research comes from addressing the actual challenges that many forecasters and modellers experience when their time series data is disturbed by recommending multiple viable and reliable strategies suitable for different cases.
+However, we recognise that presenting a new technique may be an exciting direction, and we have included it as a potential topic for future research in this area.
+Also, we included two empirical datasets from two different application areas in the paper.
+Moreover, the whole paper is reproducible, and the datasets and R codes are accessible on GitHub, allowing strategies to be tailored to any dataset in any application.
+
+\textcolor{teal}{This study proposes multiple viable and reliable strategies for forecasting time series data interrupted by different type of events.
+Future research could focus on developing a new strategy specifically for such disrupted time series, using large-scale data from various applications that are interrupted by different events, and analysing which strategy may work best in which scenarios.}
 
@@
 
-Response:
+**Response:** Thank you for taking the time to review our article.
+We appreciate your feedback and your acknowledgment of the interesting aspects of our work.
+Your comments are valuable to us and will help improve the quality of our work.
+We understand your concern about the empirical design, and we will discuss this in our response to your specific comments below.
 
@@
 
-Response:
+**Response:** Thank you for your feedback.
+We have not claimed that our paper represents a major or groundbreaking discovery.
+However, we believe that our research tackles fundamental challenges that many forecasters and modellers face when working with time series data that has been disrupted by events such as COVID-19, strikes, etc.
+This challenge was highlighted throughout our conversations with academics and practitioners in the field, who emphasised the issues they face when forecasting such time series data.
+Therefore, we think addressing this practical challenge is where our work stands out.
+Additionally, our work may open avenues for further research and reflection in this area.
+Also, we made sure that the entire work was reproducible, with datasets and R codes available on GitHub, allowing other academics and practitioners to tailor the strategies to their own needs.
 
@@
 
-Response:
+**Response:** Thank you for your helpful comment.
+We agree that the patterns of interrupted events may vary, which would affect the suitability of proposed approaches.
+That is why we have proposed different strategies to consider that might be appropriate in different cases, each with its own pros and cons.
+In future work, a large scale empirical study covering several types of disruption could be conducted to provide deeper insights and any possible link between strategies and types of disruption events.
+We have included the following paragraph in the conclusion section of the revised manuscript:
+
+\textcolor{teal}{This study proposes multiple viable and reliable strategies for forecasting time series data interrupted by different type of events.
+Future research could focus on developing a new strategy specifically for such disrupted time series, using large-scale data from various applications that are interrupted by different events, and analysing which strategy may work best in which scenarios.}
 
@@
 However, the authors may not expect the readers to judge the optimal strategy from these figures.
-Instead, the authors should consider some out-of-sample validation and may introduce some metrics or measures to evaluate the performance different strategies._
+Instead, the authors should consider some out-of-sample validation and may introduce some metrics or measures to evaluate the performance different strategies._ Rob
 
@@
 
-Response:
+**Response:** Thank you for your suggestion.
+We completely agree that incorporating exogenous variables related to specific events could enhance the performance of forecasting models for interrupted time series.
+We have acknowledged this by adding the following paragraph to the revised manuscript:
+
+\textcolor{teal}{When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated, since they may improve forecasting model performance.}
 
@@
 A few suggestions need to be included such as:
+
 _Comment 1: Different strategies are discussed in section 2 in a few words without giving a proper introduction of the strategies.
 It makes the reading of the manuscript difficult.
-The authors are required to give practical examples of the strategies to make it understandable to read._
+The authors are required to give practical examples of the strategies to make it understandable to read._ Rob
 
@@
 Authors are required to provide detailed insight into the strategies i.e.
-which datasets these strategies are applied?_
+which datasets these strategies are applied?_ Rob
 
@@
 
-_Comment 4. Two datasets are used and the strategies are applied to them. However, the descriptive statistics of these data sets missing.. The authors are required to explain the datasets._
+_Comment 4. Two datasets are used and the strategies are applied to them. However, the descriptive statistics of these data sets missing.. The authors are required to explain the datasets._ Rob
 

### 2024-05-15 response_letter/response_letter_JORS.qmd
add response for comment 5 R4

@@
 
-_Comment 3. The figures presented in the manuscripts require detailed explanation. Authors are required to explain the given figures in detail one by one._
+_Comment 3. The figures presented in the manuscripts require detailed explanation. Authors are required to explain the given figures in detail one by one._ Bahman & Rob
 
@@
 
-_Comment 5. Literature must reflect the relevant studies i.e.
-a) OntoCOVID: Ontology for Semantic Modeling of COVID19 Statistical Data
-Authors
-Shaukat Ali, Shah Khusro, Sajid Anwar, Abrar Ullah
-Publication date
-2022
-Book
-Proceedings of International Conference on Information Technology and Applications
-Pages
-183-194
-Publisher
-Springer, Singapore
-b) COVID-19 patient count prediction using LSTM
-M Iqbal, F Al-Obeidat, F Maqbool, S Razzaq, S Anwar, A Tubaishat, ...
-IEEE Transactions on Computational Social Systems 8 (4), 974-981
-c)Countering malicious URLs in internet of things using a knowledge-based approach and a simulated expert
-S Anwar, F Al-Obeidat, A Tubaishat, S Din, A Ahmad, FA Khan, G Jeon, ...
-IEEE Internet of Things Journal 7 (5), 4497-4504
+_Comment 5. Literature must reflect the relevant studies i.e._
+
+_a) OntoCOVID: Ontology for Semantic Modeling of COVID19 Statistical Data Authors Shaukat Ali, Shah Khusro, Sajid Anwar, Abrar Ullah Publication date 2022 Book Proceedings of International Conference on Information Technology and Applications Pages 83-194 Publisher Springer, Singapore_
+
+_b) COVID-19 patient count prediction using LSTM M Iqbal, F Al-Obeidat, F Maqbool, S Razzaq, S Anwar, A Tubaishat, ... IEEE Transactions on Computational Social Systems 8 (4), 974-981_
+
+_c) Countering malicious URLs in internet of things using a knowledge-based approach and a simulated expert S Anwar, F Al-Obeidat, A Tubaishat, S Din, A Ahmad, FA Khan, G Jeon, ...
+IEEE Internet of Things Journal 7 (5), 4497-4504_
+
+**Response:** Thank you for your comment.
+We have made sure to include relevant literature in our paper.
+We would like to clarify that our study does not focus on forecasting COVID-19-related data and variables, such as COVID19 patient count, or statistical data, as suggested.
+Instead, our paper addresses interrupted time series data caused by events like pandemics, strikes, and other disruptions.
+We have included literature relevant to this topic.
+Your feedback is appreciated, and we hope this clarifies the scope and focus of our research.
 

### 2024-05-16 response_letter/response_letter_JORS.qmd
Updated responses

@@
 
-**Response:** Thank you very much for giving us an opportunity to revise the manuscript.
-The valuable feedback provided by the referees.
-We have thoroughly reviewed their comments and made any required changes.
+**Response:** Thank you for giving us an opportunity to revise the manuscript.
+We have thoroughly reviewed the feedback provided by the referees, and made some changes as a result.
 We have highlighted all changes with coloured text in both the response letter and the manuscript.
@@
 
-_**General comment:**The paper deals with the prediction of time series when a disruption happens in the data.
+_**General comment:** The paper deals with the prediction of time series when a disruption happens in the data.
 Different prediction strategies are proposed, simulated and performances are compared.
@@
 
-_Comment1. one typo, line 49, page 4 : "remaining"_
+_Comment 1. one typo, line 49, page 4 : "remaining"_
 
@@
 
-_Comment2. ref Trapejo (2023) is incomplete_
+_Comment 2. ref Trapejo (2023) is incomplete_
 
@@
 
-_Comment 1:The topic is very interesting and very timely, in the wake of the Covid -19 related disruptions in supply chains, changes in consumer preferences, and inflationary environment.
+_Comment 1: The topic is very interesting and very timely, in the wake of the Covid -19 related disruptions in supply chains, changes in consumer preferences, and inflationary environment.
 The time series patterns and causal relationships that were seen before the pandemic did not apply and whether the patterns observed before the pandemic will continue to be relevant remains to be evaluated on a case by case basis._
 
-**Response:** Thank you for your insightful feedback.
-We are delighted to hear that you appreciate the timeliness and importance of the paper.
+**Response:** We are delighted to hear that you appreciate the timeliness and importance of the paper.
 We agree that the topic is particularly relevant given the COVID-19-related disruptions in supply chains and various other areas.
@@
 The idea for this paper emerged from our close contact with practitioners, who highlighted the challenges they face in forecasting time series data affected by COVID-19, particularly in handling observations impacted by the disruption.
-Following these discussions, we carried out a search to find out if any existing papers addressed this issue.
-As world-leading experts in the field of forecasting and well-connected to the research community, we could not find any published research in this area.
+Following these discussions, we carried out a search to find out if any existing papers addressed this issue, but we could not find any published research in this area.
 To our knowledge, this is the first study to describe and compare general strategies for forecasting interrupted time series, such as COVID-19-affected data.
 
-_Comment 3:The paper is well written, but in terms of contributions, it falls short of the expectations for an original research paper.
+_Comment 3: The paper is well written, but in terms of contributions, it falls short of the expectations for an original research paper.
 Improvement avenues are proposing a viable new method that has advantages over these existing methods, or a large-scale empirical research with carefully curated set of actual data representing different interruption patterns from different domains._
@@
 We believe that the originality of our research comes from addressing the actual challenges that many forecasters and modellers experience when their time series data is disturbed by recommending multiple viable and reliable strategies suitable for different cases.
-However, we recognise that presenting a new technique may be an exciting direction, and we have included it as a potential topic for future research in this area.
+We recognise that presenting a new technique may be an exciting direction, but it is not the aim of this paper.
+So we have included it as a potential topic for future research in this area.
 Also, we included two empirical datasets from two different application areas in the paper.
@@
 
-\textcolor{teal}{This study proposes multiple viable and reliable strategies for forecasting time series data interrupted by different type of events.
-Future research could focus on developing a new strategy specifically for such disrupted time series, using large-scale data from various applications that are interrupted by different events, and analysing which strategy may work best in which scenarios.}
+\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
+Future research could focus on exploring new strategies for handling such disrupted time series, or on further exploring which strategy may work best in different scenarios.}
 
@@
 
-**Response:** Thank you for taking the time to review our article.
-We appreciate your feedback and your acknowledgment of the interesting aspects of our work.
-Your comments are valuable to us and will help improve the quality of our work.
-We understand your concern about the empirical design, and we will discuss this in our response to your specific comments below.
+**Response:** Thank you for taking the time to review our article. We appreciate your feedback and your acknowledgment of the interesting aspects of our work.
 
@@
 
-**Response:** Thank you for your feedback.
-We have not claimed that our paper represents a major or groundbreaking discovery.
-However, we believe that our research tackles fundamental challenges that many forecasters and modellers face when working with time series data that has been disrupted by events such as COVID-19, strikes, etc.
+**Response:** We believe that our research tackles fundamental challenges that many forecasters and modellers face when working with time series data that has been disrupted by events such as COVID-19, strikes, etc.
 This challenge was highlighted throughout our conversations with academics and practitioners in the field, who emphasised the issues they face when forecasting such time series data.
 Therefore, we think addressing this practical challenge is where our work stands out.
+Further, we have not found any other papers addressing this issue.
+Consequently, we feel that we are addressing an important and widespread problem.
+
 Additionally, our work may open avenues for further research and reflection in this area.
@@
 
-**Response:** Thank you for your helpful comment.
-We agree that the patterns of interrupted events may vary, which would affect the suitability of proposed approaches.
+**Response:** We agree that the patterns of interrupted events may vary, which would affect the suitability of proposed approaches.
 That is why we have proposed different strategies to consider that might be appropriate in different cases, each with its own pros and cons.
+Focussing on one event would greatly limit the scope of the paper, making it far less useful.
+
 In future work, a large scale empirical study covering several types of disruption could be conducted to provide deeper insights and any possible link between strategies and types of disruption events.
@@
 However, the authors may not expect the readers to judge the optimal strategy from these figures.
-Instead, the authors should consider some out-of-sample validation and may introduce some metrics or measures to evaluate the performance different strategies._ Rob
+Instead, the authors should consider some out-of-sample validation and may introduce some metrics or measures to evaluate the performance different strategies._
 
-Response:
+Response: The visualizations are intended to help the reader understand the effect and intent of different strategies.
+In determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
+We think the visualizations help the reader do that.
+Also, it does not always make sense to do a quantitative comparison of forecast accuracy.
+We have attempted to address this issue by adding the following paragraph in the discussion section.
+
+\textcolor{teal}{In determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
+This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy.
+The strategies we discuss have different ways of handling the disrupted period, and many of them do not seek to provide accurate forecasts during the disruption.
+For example, the "set to missing" strategy ignores the disrupted period, while the "estimate what might have been" strategy replaces the observations during the disruption with interpolations.
+In both cases, it does not make sense to evaluate the performance of the method during the disruption.
+Instead, the analyst needs to decide what approach makes the most sense for the application at hand.
+}
 
@@
 **Response:** Thank you for your suggestion.
-We completely agree that incorporating exogenous variables related to specific events could enhance the performance of forecasting models for interrupted time series.
+We agree that incorporating exogenous variables related to specific events could enhance the performance of forecasting models for interrupted time series.
 We have acknowledged this by adding the following paragraph to the revised manuscript:
@@
 It makes the reading of the manuscript difficult.
-The authors are required to give practical examples of the strategies to make it understandable to read._ Rob
+The authors are required to give practical examples of the strategies to make it understandable to read._
+
+We are not clear what is being requested here.
+We describe the strategies in Section 2, and we provide fully reproducible code showing how they are implemented.
+We also give practical examples of the strategies in Section 3.
+Please explain what more is required.
 
@@
 Authors are required to provide detailed insight into the strategies i.e.
-which datasets these strategies are applied?_ Rob
+which datasets these strategies are applied?_
+
+While we could add equations in Section 2 for each strategy, we do not think it would add to the readability or accessibility of the paper.
+We are not undertaking a mathematical analysis of the strategies, so we have opted for a verbal description accompanied by code that implements each strategy.
+However, if the editor would prefer us to add equations providing a more mathematical description of each strategy, we can do so.
+
+The practical implementation of the strategies is provided in Section 3 and in the supplementary code.
+
+The data sets are described in Section 3.
+
+Insights are provided in Sections 3 and 4.
 
@@
 
-_Comment 4. Two datasets are used and the strategies are applied to them. However, the descriptive statistics of these data sets missing.. The authors are required to explain the datasets._ Rob
+_Comment 4. Two datasets are used and the strategies are applied to them. However, the descriptive statistics of these data sets missing.. The authors are required to explain the datasets._
+
+We do not believe the addition of descriptive statistics will add anything useful to the manuscript.
+We have provided the raw data, and we have provided time plots to help visualize the data.
+Each data set is described in Section 3.
 
@@
 **Response:** Thank you for your comment.
-We have made sure to include relevant literature in our paper.
 We would like to clarify that our study does not focus on forecasting COVID-19-related data and variables, such as COVID19 patient count, or statistical data, as suggested.
+So none of these references appear to be relevant to our paper.
 Instead, our paper addresses interrupted time series data caused by events like pandemics, strikes, and other disruptions.
 We have included literature relevant to this topic.
-Your feedback is appreciated, and we hope this clarifies the scope and focus of our research.
+We hope this clarifies the scope and focus of our research.
 

### 2024-05-16 fits.qmd
Responding to Stefan's comments

@@
 We fit ETS, ARIMA, and dynamic ARIMA models to the data [@fpp3], first applying a log transformation to ensure the resulting forecasts are positive.
-This is possible because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> filter(Visitors == min(Visitors)) |> pull(Month) |> format("%B %Y")`.
+(A bias correction is also applied to ensure the back-transformed forecasts are the means of the forecast distributions.) A log transformation is possible because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> filter(Visitors == min(Visitors)) |> pull(Month) |> format("%B %Y")`.
 For all forecasts, we also show 90% prediction intervals.
@@
 The first and fourth solutions carry double weight, because both the ETS and ARIMA forecasts were included in the ensemble.
-This time, no prediction intervals have been computed.
+The prediction intervals are formed from the mixture distributions obtained by combining the four forecast distributions for each time horizon.
 The first set of forecasts has not taken any account of the disruption, and so they are almost identical to those obtained in the previous plots.
@@
 #| fig-pos: "!t"
-#| fig-cap: "**Ensemble approach**, combining the previous four approaches shown in Figures [-@fig-tsol1-plot]--[-@fig-tsol4-plot]."
+#| fig-cap: "**Ensemble approach**, combining the previous four approaches shown in Figures [-@fig-tsol1-plot]--[-@fig-tsol4-plot].
+The point forecasts are the average of the four approaches, while the prediction intervals are formed by combining the four forecast distributions at each horizon to form a mixture distribution."
 tar_read(tensemble_plot)
@@
 #| fig-height: 3
-#| fig-cap: "**Highly adaptive model**: the forecasts do not take into account the lockdown periods, but the ARIMA model is largely able to adjust to the changing level of the series."
+#| fig-cap: "**Highly adaptive model**: the forecasts do not take into account the lockdown periods, but the ARIMA model is largely able to adjust to the changing level of the series.
+Forecasts are produced on a rolling basis each week, with a 7-day forecast horizon in each case."
 tar_read(walkers_plot1)

### 2024-05-16 R/tourism.R

@@ -124,7 +124,7 @@ tourism_sol4 <- function(austa) {
     mutate(month = month(Month)) |>
     left_join(ave_3y, by = "month") |>
     mutate(
-      Visitors_adj = if_else(Month < yearmonth("2022 Nov") &
+      Visitors = if_else(Month < yearmonth("2022 Nov") &
         Month > yearmonth("2020 Feb"),
       ave, Visitors
       ),
@@ -132,8 +132,8 @@ tourism_sol4 <- function(austa) {
   austa |>
     stretch_tsibble(.step = 12, .init = 240) |>
     model(
-      ets = ETS(log(Visitors_adj)),
-      arima = ARIMA(log(Visitors_adj))
+      ets = ETS(log(Visitors)),
+      arima = ARIMA(log(Visitors))
     ) |>
     forecast(h = 12) |>
     mutate(.id = paste("Forecasts of", .id + 2019))
@@ -146,13 +146,13 @@ tourism_ensemble <- function(object) {
       x |> as_tibble() |> mutate(.model = NULL)
     }) |> 
     bind_rows() |> 
-    group_by(Month) |> 
-    summarise(
-      .id = unique(.id),
-      .mean = mean(.mean)
-    ) |> 
-    mutate(Visitors = distributional::dist_degenerate(.mean)) |> 
-    as_fable(index = Month, key = .id, response = "Visitors", distribution = Visitors)
+    mutate(Visitors = generate(Visitors, 2000)) |> 
+    group_by(.id, Month) |> 
+    summarise(Visitors = distributional::dist_sample(list(c(unlist(Visitors))))) |> 
+    ungroup() |> 
+    as_fable(index = Month, key = .id,
+        response = "Visitors", distribution = Visitors) |> 
+    suppressWarnings()
 }
 
 

### 2024-05-16 _targets.R

@@ -57,7 +57,7 @@ list(
   ),
   tar_target(
     tensemble_plot,
-    autoplot(tensemble, level = NULL) + tourism_history + tourism_labels
+    autoplot(tensemble, level = 90) + tourism_history + tourism_labels
   ),
   
   # Pedestrians

### 2024-05-16 response_letter/response_letter_JORS.qmd
finalising response letter, and revise the manuscript.

@@
 ---
-title: "Response letter - Manuscript ID JSR-22-468"
+title: "Response letter - Manuscript ID 249381018"
 bibliography: bibliography.bib
@@
 We have highlighted all changes with coloured text in both the response letter and the manuscript.
+
+We also wanted to provide further clarification to the editor regarding the comment provided by Reviewer 4.
+
+- Reviewer 4 (comment 2) suggested including mathematical equations for each strategy.
+While we could add these equations in Section 2, we believe it would not improve the readability or accessibility of the paper.
+However, if the editor prefers a more mathematical description of each strategy, we are willing to include the equations.
+Since we are not conducting a mathematical analysis of the strategies, we have opted for a verbal description supported by code implementations.
+
+- Reviewer 4 (comment 4) asked to add descriptive statistics.
+We do not believe the addition of descriptive statistics will add anything useful to the manuscript.
+We argue that the usefulness of descriptive statistics for time series analysis is limited.
+We have provided the raw data, and we have provided time plots to help visualize the data.
+Each data set is described in Section 3.
+
+- Reviewer 4 (comment 5) recommended including three papers from the literature.
+However, these papers are not relevant to our current study.
+The suggested papers focus on forecasting COVID-19 variables, which is not the focus of our paper.
 
@@
 
-\textcolor{teal}{The assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis.
+\textcolor{teal}{We note that the assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis.
 Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
@@
 \textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
-Future research could focus on exploring new strategies for handling such disrupted time series, or on further exploring which strategy may work best in different scenarios.}
+Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data.
+Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}
 
@@
 That is why we have proposed different strategies to consider that might be appropriate in different cases, each with its own pros and cons.
-Focussing on one event would greatly limit the scope of the paper, making it far less useful.
+Focusing on one event would greatly limit the scope of the paper, making it far less useful.
 
@@
 
-\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
-Future research could focus on exploring new strategies for handling such disrupted time series, or on further exploring which strategy may work best in different scenarios.}
+\textcolor{teal}Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}
 
@@
 
-\textcolor{teal}{In determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
+\textcolor{teal}{Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
 This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy.
@@
 In both cases, it does not make sense to evaluate the performance of the method during the disruption.
-Instead, the analyst needs to decide what approach makes the most sense for the application at hand.
-}
+Instead, the analyst needs to decide what approach makes the most sense for the application at hand.}
 
@@
 
-\textcolor{teal}{When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated, since they may improve forecasting model performance.}
+\textcolor{teal}{We should also note that when forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated, since they may improve forecasting model performance.}
 
@@
 
+**Response:** Thank you.
 We are not clear what is being requested here.
@@
 We also give practical examples of the strategies in Section 3.
-Please explain what more is required.
+Please explain what more is required to be included, and we are willing to do that.
 
@@
 
+**Response:** Thank you for the comment.
+We have carefully thought through this request.
 While we could add equations in Section 2 for each strategy, we do not think it would add to the readability or accessibility of the paper.
@@
 However, if the editor would prefer us to add equations providing a more mathematical description of each strategy, we can do so.
+To respond to other remarks:
 
-The practical implementation of the strategies is provided in Section 3 and in the supplementary code.
+- The practical implementation of the strategies is provided in Section 3 and in the supplementary code.
 
-The data sets are described in Section 3.
+- The data sets are described in Section 3.
 
-Insights are provided in Sections 3 and 4.
+- Insights are provided in Sections 3 and 4.
 
-_Comment 3. The figures presented in the manuscripts require detailed explanation. Authors are required to explain the given figures in detail one by one._ Bahman & Rob
+_Comment 3. The figures presented in the manuscripts require detailed explanation. Authors are required to explain the given figures in detail one by one._
+
+**Response:** Thank you for your feedback.
+We have provide a detailed explanation of each figure in the manuscript, as well as in the captions of each figure, discussing them one by one to ensure clarity.
 
@@
 
+**Response:** Thank you for this comment.
 We do not believe the addition of descriptive statistics will add anything useful to the manuscript.
-We have provided the raw data, and we have provided time plots to help visualize the data.
-Each data set is described in Section 3.
+We argue that the usefulness of descriptive statistics for time series analysis is limited.
+First, they cannot highlight patterns in the data, identify unusual observations, describe changes over time, or help to understand relationships between variables.
+Second, they could be misleading, since similar statistics can stem from very different datasets (please see Anscombe’s Quartet).
+
+We have provided the raw data, and we have provided time plots to help visualize the data. Each data set is described in Section 3.
 

### 2024-05-16 fits.qmd
finalising response letter, and revise the manuscript.

@@
 Outbreaks, epidemics, and pandemics, for example, may have a considerable impact on population behavior, causing discrepancies and interruptions of data in a variety of industries such as healthcare, pharmaceuticals, transportation, retail, tourism, and traffic management.
-Similarly, supply chain disruptions induced by factors such as strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps.
-Transportation-related data may be interrupted due to equipment failures, such as malfunctions or breakdowns in vehicles, aircraft, or rail systems, resulting in missing or inaccurate data points.
+Similarly, transportation-related data may be interrupted due to equipment failures, such as malfunctions or breakdowns in vehicles, aircraft, or rail systems, resulting in missing or inaccurate data points.
 Furthermore, cybersecurity breaches data quality by possibly skewing readings in datasets relating to financial transactions, network traffic, and user activity, complicating time series analysis.
+Supply chain disruptions induced by factors such as strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps.
+\textcolor{teal}{The war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of COVID-19 toilet paper shortage.
+A power outage in Eastern Germany halted production at a Tesla facility and cut off electricity to a large retailer distribution centre, disrupting supplies to around 500 stores in and around Berlin for several weeks.
+Recent strikes in various forms of transport, including air and rail in Germany, France, and the United Kingdom, have had an impact on passenger flow and retail sales at airports and train stations.
+These issues highlight the significance of robust forecasting systems in addressing such disruptions effectively.}
 
@@
 In this section, we describe several possible strategies to handle interruptions when forecasting time series data.
+\textcolor{teal}{We note that the assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis.
+Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
 
@@
 
+\textcolor{teal}{The disadvantage of smoothing and other highly adaptive techniques is that they need time to adapt after a disturbance and can eventually end up chasing noise.
+The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting, ignoring the presence of noise in the data.}
+
 ## Use an intervention model
@@
 This allows the change period to be effectively modelled, provided the intervention variables are chosen well.
-However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.
+\textcolor{teal}{Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, might become subjective.} However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.
 
@@
 
+\textcolor{teal}{We should also note that when forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated, since they may improve forecasting model performance.}
+
 We have not discussed probabilistic forecasting in any detail in this paper, other than producing some prediction intervals for some of the examples.
@@
 
-Future research might focus on enhancing probabilistic forecasting methods that are designed specifically for interrupted time series.
-Large-scale empirical research evaluating the effectiveness of different strategies in real-world settings might also provide useful insights.
-Comparative studies that assess the accuracy, robustness, and computing efficiency of different techniques across several datasets and disruption situations can help discover the most effective approaches for a range of practical applications.
+\textcolor{teal}{Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
+This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy.
+The strategies we discuss have different ways of handling the disrupted period, and many of them do not seek to provide accurate forecasts during the disruption.
+For example, the "set to missing" strategy ignores the disrupted period, while the "estimate what might have been" strategy replaces the observations during the disruption with interpolations.
+In both cases, it does not make sense to evaluate the performance of the method during the disruption.
+Instead, the analyst needs to decide what approach makes the most sense for the application at hand.}
+
+\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
+Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data.
+Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.} Comparative studies that assess the accuracy, robustness, and computing efficiency of different techniques across several datasets and disruption situations can help discover the most effective approaches for a range of practical applications.
 
@@
 
+# Acknowledgment {.unnumbered}
+
+We would like to thank Dr.
+Stephan Kolassa for sharing his thoughts and providing valuable feedback on the first version of our paper.
+His voluntary feedback and recommendations have helped us to refine and improve our work.
+
 # References

### 2024-05-17 _targets.R
Updated header.tex and add to targets trigger

@@ -30,7 +30,7 @@ list(
   tar_target(tsol3, tourism_sol3(austa)),
   tar_target(tsol4, tourism_sol4(austa)),
   tar_target(tensemble, tourism_ensemble(list(tsol1,tsol2,tsol3,tsol4))),
-  
+
   # Forecast plots
   tar_target(
     tourism_history,
@@ -59,7 +59,7 @@ list(
     tensemble_plot,
     autoplot(tensemble, level = 90) + tourism_history + tourism_labels
   ),
-  
+
   # Pedestrians
   tar_target(raw_walkers, fetch_walkers(start = "2019-01-01", end = "2021-12-31")),
   tar_target(lockdowns, lockdown_table()),
@@ -80,7 +80,7 @@ list(
   tar_target(walkers_plot3, pedestrian_fc_plot(walkers_sol3, walkers)),
   tar_target(walkers_plot4, pedestrian_fc_plot(walkers_sol4, walkers)),
   tar_target(walkers_ensemble_plot, pedestrian_fc_plot(wensemble, walkers)),
-  
+
   # Attended incidents example
   tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format = "file"),
   tar_target(verified_csv, here::here("data/verified_incident.csv"), format = "file"),
@@ -121,6 +121,6 @@ list(
   tar_target(fc, forecast(fit, h = 365)),
   tar_quarto(paper,
     "fits.qmd",
-    extra_files = "references.bib"
+    extra_files = c("references.bib", "header.tex")
   )
 )

### 2024-05-17 response_letter/response_letter_JORS.qmd
Final changes to revised paper and response

@@
 
-Dear Professor
+Dear Professor Zhang
 
@@
 
-\textcolor{teal}{We note that the assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis.
+\textcolor{teal}{The handling of time series interruptions is a complex and multifaceted issue, and the choice of the most appropriate strategy depends on the specific characteristics of the data and the nature of the interruption.
 Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
@@
 
-\textcolor{teal}{We should also note that when forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated, since they may improve forecasting model performance.}
+\textcolor{teal}{When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated where possible, regardless of which approach is employed, since they may improve forecasting model performance.}
 
@@
 **Response:** Thank you for your feedback.
-We have provide a detailed explanation of each figure in the manuscript, as well as in the captions of each figure, discussing them one by one to ensure clarity.
+We have provide a detailed explanation of each figure in the manuscript, discussing them one by one to ensure clarity.
+We have now added some of the same information in the captions of each figure to aid readability.
 

### 2024-05-17 fits.qmd
Final changes to revised paper and response

@@
 Supply chain disruptions induced by factors such as strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps.
-\textcolor{teal}{The war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of COVID-19 toilet paper shortage.
+\textcolor{teal}{For example, the war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of the COVID-19 toilet paper shortage.
 A power outage in Eastern Germany halted production at a Tesla facility and cut off electricity to a large retailer distribution centre, disrupting supplies to around 500 stores in and around Berlin for several weeks.
@@
 
-It is important to note that in this study, we consider the problem of forecasting interrupted time series both during and after the event.
+We consider the problem of forecasting interrupted time series both during and after the event.
 This is a separate problem from change point detection [@aue2013structural;@truong2020selective] and anomaly detection [@talagala2020anomaly].
@@
 In this section, we describe several possible strategies to handle interruptions when forecasting time series data.
-\textcolor{teal}{We note that the assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis.
+\textcolor{teal}{The handling of time series interruptions is a complex and multifaceted issue, and the choice of the most appropriate strategy depends on the specific characteristics of the data and the nature of the interruption.
 Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
@@
 \textcolor{teal}{The disadvantage of smoothing and other highly adaptive techniques is that they need time to adapt after a disturbance and can eventually end up chasing noise.
-The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting, ignoring the presence of noise in the data.}
+The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting.}
 
@@
 This allows the change period to be effectively modelled, provided the intervention variables are chosen well.
-\textcolor{teal}{Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, might become subjective.} However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.
+\textcolor{teal}{Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, may become subjective.} However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.
 
@@
 Then the model can be re-estimated using the adjusted data.
-Note that forecasts obtained in this way during the disruption will not be true forecasts because they will have used data from the future in computing the adjusted data during the disruption.
+Forecasts obtained in this way during the disruption will not be true forecasts because they will have used data from the future in computing the adjusted data during the disruption.
 But post-disruption forecasts will be true forecasts and should be almost the same as those obtained using the previous solution.
@@
 #| label: fig-tourism-plot1
-#| fig-cap: "Short-term visitor arrivals to Australia (monthly): Jan 2000 -- May 2023."
+#| fig-cap: "Short-term visitor arrivals to Australia (monthly): Jan 2000 -- May 2023.
+Source: @tourismdata.
+The Australian international borders closed in March 2020 and officially reopened in February 2022."
 #| fig-height: 3
@@
 #| fig-pos: "!b"
-#| fig-cap: "**Highly adaptive models**. Forecasts from ETS and ARIMA models. Neither works particularly well for disruptions of this magnitude."
+#| fig-cap: "**Highly adaptive models**. Each panel shows one year of monthly forecasts from ETS and ARIMA models. Neither works particularly well for disruptions of this magnitude."
 tar_read(tsol1_plot)
@@
 #| fig-cap: "**Intervention model**.
-We use a level shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022.
+Each panel shows one year of monthly forecasts from an intervention model containing a level shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022.
 After October 2022, all intervention variables are set to zero."
@@
 #| fig-cap: "**Set to missing**.
+Each panel shows one year of monthly forecasts from an ARIMA model.
 The period from March 2020 to October 2022 is set to missing.
@@
 #| fig-pos: "!b"
-#| fig-cap: "**Estimate what might have been**. We replace the observations from March 2020 to October 2022 with the average of the same month in the three years prior to March 2020."
+#| fig-cap: "**Estimate what might have been**.
+Each panel shows one year of monthly forecasts from ETS and ARIMA models.
+We have replaced the observations from March 2020 to October 2022 with the average of the same month in the three years prior to March 2020."
 tar_read(tsol4_plot)
@@
 The resulting forecasts are similar to those from solution 3, but with narrower prediction intervals because the model is (falsely) assuming that the "observations" during the disruption are real.
-
-Finally, in @fig-tensemble-plot, we show the ensemble forecasts obtained by averaging the forecasts from the four solutions discussed above.
-The first and fourth solutions carry double weight, because both the ETS and ARIMA forecasts were included in the ensemble.
-The prediction intervals are formed from the mixture distributions obtained by combining the four forecast distributions for each time horizon.
-The first set of forecasts has not taken any account of the disruption, and so they are almost identical to those obtained in the previous plots.
-The last set of forecasts is reasonably good, showing what is obtained by combining the forecasting methods used to handle the disruption period.
-However, the forecasts for 2021 and 2022 are particularly poor, because they are averaging forecasts that aim to achieve different objectives, and are affected by the poor forecasts obtained using the dynamic regression model.
 
@@
 
-\FloatBarrier
+Finally, in @fig-tensemble-plot, we show the ensemble forecasts obtained by averaging the forecasts from the four solutions discussed above.
+The first and fourth solutions carry double weight, because both the ETS and ARIMA forecasts were included in the ensemble.
+The prediction intervals are formed from the mixture distributions obtained by combining the four forecast distributions for each time horizon.
+The first set of forecasts has not taken any account of the disruption, and so they are almost identical to those obtained in the previous plots.
+The last set of forecasts is reasonably good, showing what is obtained by combining the forecasting methods used to handle the disruption period.
+However, the forecasts for 2021 and 2022 are particularly poor, because they are averaging forecasts that aim to achieve different objectives, and are affected by the poor forecasts obtained using the dynamic regression model.
 
@@
 #| fig-height: 3.7
-#| fig-cap: "Pedestrian traffic in Melbourne, Australia, from 2018 to 2023. The gray shaded regions indicate periods of lockdown due to COVID-19."
+#| fig-cap: "Average pedestrian traffic in Melbourne, Australia, from 2018 to 2023, based on automated hourly counts from sensors around the city.
+The counts are averaged over `r ngoodsensors` sensors, each of which had no more than 30 days of missing data.
+The gray shaded regions indicate periods of lockdown due to COVID-19.
+Source: @pedestrians"
 tar_read(walkers_plot)
@@
 
+We use ARIMA models in @fig-walkers-plot1, which have mostly been able to adapt to the changing level of the series, apart from during the first few weeks of the first lockdown, and the first day or two of subsequent lockdowns.
+
 ```{r}
@@
 #| fig-height: 3
-#| fig-cap: "**Highly adaptive model**: the forecasts do not take into account the lockdown periods, but the ARIMA model is largely able to adjust to the changing level of the series.
-Forecasts are produced on a rolling basis each week, with a 7-day forecast horizon in each case."
+#| fig-cap: "**Highly adaptive model**: Rolling 7-day forecasts of daily pedestrian data, made each week, using ARIMA models.
+The forecasts do not take into account the lockdown periods, but the ARIMA models are largely able to adjust to the changing level of the series."
 tar_read(walkers_plot1)
 ```
-
-We use ARIMA models in @fig-walkers-plot1, which have mostly been able to adapt to the changing level of the series, apart from during the first few weeks of the first lockdown, and the first day or two of subsequent lockdowns.
-
-In @fig-walkers-plot2, we have used an intervention model with a dummy variable indicating the lockdown periods.
-This has worked a little better than the ARIMA model used in @fig-walkers-plot1, apart from around the first lockdown period.
-Many people were in self-imposed lockdown before the start of the first lockdown period and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
 
@@
 #| fig-height: 3
-#| fig-cap: "**Intervention model**: the forecasts are based on an ARIMA model with a dummy variable indicating when the lockdown periods occurred.
+#| fig-cap: "**Intervention model**: Rolling 7-day forecasts of daily pedestrian data, made each week, using ARIMA models with a dummy variable indicating when the lockdown periods occurred.
 Many people were in self-imposed lockdown before the start of the first lockdown period, and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
@@
 #| fig-height: 3
-#| fig-cap: "**Set to missing**: the lockdown periods are set to missing, and the ARIMA model uses the remaining data to produce forecasts.
+#| fig-cap: "**Set to missing**: Rolling 7-day forecasts of daily pedestrian data, made each week, using ARIMA models, where the lockdown periods are set to missing.
+The ARIMA models use the remaining data to produce forecasts.
 Thus, the forecasts are based on only previous non-lockdown data.
@@
 #| fig-height: 3
-#| fig-cap: "**Estimate what might have been**: the lockdown periods are replaced with estimated counts based on an ARIMA model applied to the remaining data.
-Then a new ARIMA model is fitted to the whole data set, and used to produce forecasts.
+#| fig-cap: "**Estimate what might have been**: Rolling 7-day forecasts of daily pedestrian data, made each week.
+The lockdown periods are replaced with estimated counts based on an ARIMA model applied to the remaining data.
+Then a new ARIMA model is fitted to each data set, and used to produce forecasts.
 As a result, the forecasts during the lockdown periods are not true forecasts, as they use non-lockdown data from the future."
@@
 #| fig-height: 3
-#| fig-cap: "**Ensemble forecasts** based on the first two solutions shown in Figures [-@fig-walkers-plot1] and [-@fig-walkers-plot2].
+#| fig-cap: "**Ensemble forecasts** combining the first two solutions shown in Figures [-@fig-walkers-plot1] and [-@fig-walkers-plot2].
 Other than having some difficulty around the period of the first lockdown, the results are relatively good."
@@
 ```
+
+In @fig-walkers-plot2, we have used an intervention model with a dummy variable indicating the lockdown periods.
+This has worked a little better than the ARIMA model used in @fig-walkers-plot1, apart from around the first lockdown period.
+Many people were in self-imposed lockdown before the start of the first lockdown period and were reluctant to return to the city after it ended, resulting in poor forecasts on either side of the first lockdown.
 
@@
 
-We would like to thank Dr.
-Stephan Kolassa for sharing his thoughts and providing valuable feedback on the first version of our paper.
-His voluntary feedback and recommendations have helped us to refine and improve our work.
+We thank Dr Stephan Kolassa and several referees for their thoughts and feedback on the first version of our paper. These comments have helped us to refine and improve our work.
 

### 2024-05-17 fits.tex
Final changes to revised paper and response

@@
 \usepackage[section]{placeins}
-\usepackage{orcidlink}
+\widowpenalty=10000
+\clubpenalty=10000
 \definecolor{mypink}{RGB}{219, 48, 122}
@@
 \title{Forecasting interrupted time series}
-\date{20 February 2024}
+\date{17 May 2024}
 
@@
   \blind
-  \lfoot{\sf20 February 2024}
+  \lfoot{\sf17 May 2024}
 
@@
 of industries such as healthcare, pharmaceuticals, transportation,
-retail, tourism, and traffic management. Similarly, supply chain
-disruptions induced by factors such as strikes or shortages cause
+retail, tourism, and traffic management. Similarly,
+transportation-related data may be interrupted due to equipment
+failures, such as malfunctions or breakdowns in vehicles, aircraft, or
+rail systems, resulting in missing or inaccurate data points.
+Furthermore, cybersecurity breaches data quality by possibly skewing
+readings in datasets relating to financial transactions, network
+traffic, and user activity, complicating time series analysis. Supply
+chain disruptions induced by factors such as strikes or shortages cause
 irregularities in sales, inventory levels, production, and delivery,
-often resulting in data gaps. Transportation-related data may be
-interrupted due to equipment failures, such as malfunctions or
-breakdowns in vehicles, aircraft, or rail systems, resulting in missing
-or inaccurate data points. Furthermore, cybersecurity breaches data
-quality by possibly skewing readings in datasets relating to financial
-transactions, network traffic, and user activity, complicating time
-series analysis.
+often resulting in data gaps.
+\textcolor{teal}{For example, the war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of the COVID-19 toilet paper shortage.
+A power outage in Eastern Germany halted production at a Tesla facility and cut off electricity to a large retailer distribution centre, disrupting supplies to around 500 stores in and around Berlin for several weeks.
+Recent strikes in various forms of transport, including air and rail in Germany, France, and the United Kingdom, have had an impact on passenger flow and retail sales at airports and train stations.
+These issues highlight the significance of robust forecasting systems in addressing such disruptions effectively.}
 
@@
 
-It is important to note that in this study, we consider the problem of
-forecasting interrupted time series both during and after the event.
-This is a separate problem from change point detection
-\autocite{aue2013structural,truong2020selective} and anomaly detection
-\autocite{talagala2020anomaly}. In the situations we investigate, we
-know that a change has occurred which has caused some unusual
-observations, and we want to forecast what will happen next. Change
-point detection refers to determining \emph{when} a change occurred
-\autocite{blazquez2021review}, and anomaly detection aims to identify
-unusual observations with the intention of either minimising their
-value, or paying particular attention to them, and doing a root-cause
-investigation. Our study focus also differs from \emph{intervention
-modeling}, sometimes referred to as \emph{interrupted time series
-analysis and modelling} \autocite{mcdowall2019interrupted}, which is
-used when we have data about an outcome over time and the aim is to
-assess the impact of an intervention, policy, or program implemented at
-a specific time point. We are not trying to measure the impact of the
-interruption; rather we are trying to produce sensible forecasts during
-and after the interruption.
+We consider the problem of forecasting interrupted time series both
+during and after the event. This is a separate problem from change point
+detection \autocite{aue2013structural,truong2020selective} and anomaly
+detection \autocite{talagala2020anomaly}. In the situations we
+investigate, we know that a change has occurred which has caused some
+unusual observations, and we want to forecast what will happen next.
+Change point detection refers to determining \emph{when} a change
+occurred \autocite{blazquez2021review}, and anomaly detection aims to
+identify unusual observations with the intention of either minimising
+their value, or paying particular attention to them, and doing a
+root-cause investigation. Our study focus also differs from
+\emph{intervention modeling}, sometimes referred to as \emph{interrupted
+time series analysis and modelling} \autocite{mcdowall2019interrupted},
+which is used when we have data about an outcome over time and the aim
+is to assess the impact of an intervention, policy, or program
+implemented at a specific time point. We are not trying to measure the
+impact of the interruption; rather we are trying to produce sensible
+forecasts during and after the interruption.
 
@@
 interruptions when forecasting time series data.
+\textcolor{teal}{The handling of time series interruptions is a complex and multifaceted issue, and the choice of the most appropriate strategy depends on the specific characteristics of the data and the nature of the interruption.
+Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
 
@@
 period will be similar to the pre-interruption period.
+
+\textcolor{teal}{The disadvantage of smoothing and other highly adaptive techniques is that they need time to adapt after a disturbance and can eventually end up chasing noise.
+The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting.}
 
@@
 change period to be effectively modelled, provided the intervention
-variables are chosen well. However, the model will assume that the
-post-interruption period will be similar to the pre-interruption period,
-and so the prediction intervals may be too narrow, especially if there
-is a lasting effect beyond the end of the interruption.
+variables are chosen well.
+\textcolor{teal}{Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, may become subjective.}
+However, the model will assume that the post-interruption period will be
+similar to the pre-interruption period, and so the prediction intervals
+may be too narrow, especially if there is a lasting effect beyond the
+end of the interruption.
 
@@
 and use it to estimate what might have been during the disruption. Then
-the model can be re-estimated using the adjusted data. Note that
-forecasts obtained in this way during the disruption will not be true
-forecasts because they will have used data from the future in computing
-the adjusted data during the disruption. But post-disruption forecasts
-will be true forecasts and should be almost the same as those obtained
-using the previous solution.
+the model can be re-estimated using the adjusted data. Forecasts
+obtained in this way during the disruption will not be true forecasts
+because they will have used data from the future in computing the
+adjusted data during the disruption. But post-disruption forecasts will
+be true forecasts and should be almost the same as those obtained using
+the previous solution.
 
@@
 \caption{\label{fig-tourism-plot1}Short-term visitor arrivals to
-Australia (monthly): Jan 2000 -- May 2023.}
+Australia (monthly): Jan 2000 -- May 2023. Source:
+\textcite{tourismdata}. The Australian international borders closed in
+March 2020 and officially reopened in February 2022.}
 
@@
 \autocite{fpp3}, first applying a log transformation to ensure the
-resulting forecasts are positive. This is possible because the
+resulting forecasts are positive. (A bias correction is also applied to
+ensure the back-transformed forecasts are the means of the forecast
+distributions.) A log transformation is possible because the
 observations never reach exactly zero, with the smallest number of
@@
 
-\caption{\label{fig-tsol1-plot}\textbf{Highly adaptive models}.
-Forecasts from ETS and ARIMA models. Neither works particularly well for
-disruptions of this magnitude.}
+\caption{\label{fig-tsol1-plot}\textbf{Highly adaptive models}. Each
+panel shows one year of monthly forecasts from ETS and ARIMA models.
+Neither works particularly well for disruptions of this magnitude.}
 
@@
 
-\caption{\label{fig-tsol2-plot}\textbf{Intervention model}. We use a
-level shift from March 2020 to October 2022, and a ramp from October
-2021 to October 2022. After October 2022, all intervention variables are
-set to zero.}
+\caption{\label{fig-tsol2-plot}\textbf{Intervention model}. Each panel
+shows one year of monthly forecasts from an intervention model
+containing a level shift from March 2020 to October 2022, and a ramp
+from October 2021 to October 2022. After October 2022, all intervention
+variables are set to zero.}
 
@@
 
-\caption{\label{fig-tsol3-plot}\textbf{Set to missing}. The period from
-March 2020 to October 2022 is set to missing. So the first three years
-of forecasts show what might have been. The fourth set of forecasts uses
+\caption{\label{fig-tsol3-plot}\textbf{Set to missing}. Each panel shows
+one year of monthly forecasts from an ARIMA model. The period from March
+2020 to October 2022 is set to missing. So the first three years of
+forecasts show what might have been. The fourth set of forecasts uses
 observations from November and December 2022, and so the forecasts have
@@
 \caption{\label{fig-tsol4-plot}\textbf{Estimate what might have been}.
-We replace the observations from March 2020 to October 2022 with the
-average of the same month in the three years prior to March 2020.}
+Each panel shows one year of monthly forecasts from ETS and ARIMA
+models. We have replaced the observations from March 2020 to October
+2022 with the average of the same month in the three years prior to
+March 2020.}
 
@@
 
+\begin{figure}[!t]
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-tensemble-plot-1.pdf}
+
+}
+
+\caption{\label{fig-tensemble-plot}\textbf{Ensemble approach}, combining
+the previous four approaches shown in Figures
+\ref{fig-tsol1-plot}--\ref{fig-tsol4-plot}. The point forecasts are the
+average of the four approaches, while the prediction intervals are
+formed by combining the four forecast distributions at each horizon to
+form a mixture distribution.}
+
+\end{figure}%
+
 Finally, in Figure~\ref{fig-tensemble-plot}, we show the ensemble
@@
 because both the ETS and ARIMA forecasts were included in the ensemble.
-This time, no prediction intervals have been computed. The first set of
-forecasts has not taken any account of the disruption, and so they are
-almost identical to those obtained in the previous plots. The last set
-of forecasts is reasonably good, showing what is obtained by combining
-the forecasting methods used to handle the disruption period. However,
-the forecasts for 2021 and 2022 are particularly poor, because they are
-averaging forecasts that aim to achieve different objectives, and are
-affected by the poor forecasts obtained using the dynamic regression
-model.
-
-\begin{figure}[!t]
-
-\centering{
-
-\includegraphics{fits_files/figure-pdf/fig-tensemble-plot-1.pdf}
-
-}
-
-\caption{\label{fig-tensemble-plot}\textbf{Ensemble approach}, combining
-the previous four approaches shown in Figures
-\ref{fig-tsol1-plot}--\ref{fig-tsol4-plot}.}
-
-\end{figure}%
-
-\FloatBarrier
+The prediction intervals are formed from the mixture distributions
+obtained by combining the four forecast distributions for each time
+horizon. The first set of forecasts has not taken any account of the
+disruption, and so they are almost identical to those obtained in the
+previous plots. The last set of forecasts is reasonably good, showing
+what is obtained by combining the forecasting methods used to handle the
+disruption period. However, the forecasts for 2021 and 2022 are
+particularly poor, because they are averaging forecasts that aim to
+achieve different objectives, and are affected by the poor forecasts
+obtained using the dynamic regression model.
 
@@
 
-\caption{\label{fig-walkers}Pedestrian traffic in Melbourne, Australia,
-from 2018 to 2023. The gray shaded regions indicate periods of lockdown
-due to COVID-19.}
+\caption{\label{fig-walkers}Average pedestrian traffic in Melbourne,
+Australia, from 2018 to 2023, based on automated hourly counts from
+sensors around the city. The counts are averaged over
+\texttt{r\ ngoodsensors} sensors, each of which had no more than 30 days
+of missing data. The gray shaded regions indicate periods of lockdown
+due to COVID-19. Source: \textcite{pedestrians}}
 
@@
 
-\begin{figure}
-
-\centering{
-
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot1-1.pdf}
-
-}
-
-\caption{\label{fig-walkers-plot1}\textbf{Highly adaptive model}: the
-forecasts do not take into account the lockdown periods, but the ARIMA
-model is largely able to adjust to the changing level of the series.}
-
-\end{figure}%
-
 We use ARIMA models in Figure~\ref{fig-walkers-plot1}, which have mostly
@@
 two of subsequent lockdowns.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot1-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot1}\textbf{Highly adaptive model}:
+Rolling 7-day forecasts of daily pedestrian data, made each week, using
+ARIMA models. The forecasts do not take into account the lockdown
+periods, but the ARIMA models are largely able to adjust to the changing
+level of the series.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot2-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot2}\textbf{Intervention model}: Rolling
+7-day forecasts of daily pedestrian data, made each week, using ARIMA
+models with a dummy variable indicating when the lockdown periods
+occurred. Many people were in self-imposed lockdown before the start of
+the first lockdown period, and were reluctant to return to the city
+after it ended, resulting in poor forecasts on either side of the first
+lockdown. The model was able to adjust to the changing level of the
+series during subsequent lockdowns.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot3-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot3}\textbf{Set to missing}: Rolling 7-day
+forecasts of daily pedestrian data, made each week, using ARIMA models,
+where the lockdown periods are set to missing. The ARIMA models use the
+remaining data to produce forecasts. Thus, the forecasts are based on
+only previous non-lockdown data. Naturally, the forecasts during the
+lockdown periods do not reflect reality, but they can be interpreted as
+what might have been.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-plot4-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-plot4}\textbf{Estimate what might have
+been}: Rolling 7-day forecasts of daily pedestrian data, made each week.
+The lockdown periods are replaced with estimated counts based on an
+ARIMA model applied to the remaining data. Then a new ARIMA model is
+fitted to each data set, and used to produce forecasts. As a result, the
+forecasts during the lockdown periods are not true forecasts, as they
+use non-lockdown data from the future.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{fits_files/figure-pdf/fig-walkers-ensemble-plot-1.pdf}
+
+}
+
+\caption{\label{fig-walkers-ensemble-plot}\textbf{Ensemble forecasts}
+combining the first two solutions shown in Figures
+\ref{fig-walkers-plot1} and \ref{fig-walkers-plot2}. Other than having
+some difficulty around the period of the first lockdown, the results are
+relatively good.}
+
+\end{figure}%
 
@@
 lockdown.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot2-1.pdf}
-
-}
-
-\caption{\label{fig-walkers-plot2}\textbf{Intervention model}: the
-forecasts are based on an ARIMA model with a dummy variable indicating
-when the lockdown periods occurred. Many people were in self-imposed
-lockdown before the start of the first lockdown period, and were
-reluctant to return to the city after it ended, resulting in poor
-forecasts on either side of the first lockdown. The model was able to
-adjust to the changing level of the series during subsequent lockdowns.}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot3-1.pdf}
-
-}
-
-\caption{\label{fig-walkers-plot3}\textbf{Set to missing}: the lockdown
-periods are set to missing, and the ARIMA model uses the remaining data
-to produce forecasts. Thus, the forecasts are based on only previous
-non-lockdown data. Naturally, the forecasts during the lockdown periods
-do not reflect reality, but they can be interpreted as what might have
-been.}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot4-1.pdf}
-
-}
-
-\caption{\label{fig-walkers-plot4}\textbf{Estimate what might have
-been}: the lockdown periods are replaced with estimated counts based on
-an ARIMA model applied to the remaining data. Then a new ARIMA model is
-fitted to the whole data set, and used to produce forecasts. As a
-result, the forecasts during the lockdown periods are not true
-forecasts, as they use non-lockdown data from the future.}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{fits_files/figure-pdf/fig-walkers-ensemble-plot-1.pdf}
-
-}
-
-\caption{\label{fig-walkers-ensemble-plot}\textbf{Ensemble forecasts}
-based on the first two solutions shown in Figures
-\ref{fig-walkers-plot1} and \ref{fig-walkers-plot2}. Other than having
-some difficulty around the period of the first lockdown, the results are
-relatively good.}
-
-\end{figure}%
 
@@
 the power of averaging.
+
+\textcolor{teal}{When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated where possible, regardless of which approach is employed, since they may improve forecasting model performance.}
 
@@
 
-Future research might focus on enhancing probabilistic forecasting
-methods that are designed specifically for interrupted time series.
-Large-scale empirical research evaluating the effectiveness of different
-strategies in real-world settings might also provide useful insights.
+\textcolor{teal}{Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
+This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy.
+The strategies we discuss have different ways of handling the disrupted period, and many of them do not seek to provide accurate forecasts during the disruption.
+For example, the "set to missing" strategy ignores the disrupted period, while the "estimate what might have been" strategy replaces the observations during the disruption with interpolations.
+In both cases, it does not make sense to evaluate the performance of the method during the disruption.
+Instead, the analyst needs to decide what approach makes the most sense for the application at hand.}
+
+\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
+Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data.
+Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}
 Comparative studies that assess the accuracy, robustness, and computing
@@
 
+\section*{Acknowledgment}\label{acknowledgment}
+\addcontentsline{toc}{section}{Acknowledgment}
+
+We thank Dr Stephan Kolassa and several referees for their thoughts and
+feedback on the first version of our paper. These comments have helped
+us to refine and improve our work.
+
 \printbibliography[title=References]

### 2024-05-17 response_letter/response_letter_JORS.qmd
Further changes to response letter

@@
 title: "Response letter - Manuscript ID 249381018"
-bibliography: bibliography.bib
 format:
@@
     papersize: a4
-    geometry: "margin = 2.5cm"
-    include-in-header: header.tex
+    geometry: "margin = 2.2cm"
+    header-includes:
+      - \usepackage{bm,mathpazo}
+      - \usepackage[lf]{FiraSans}
+      - \usepackage[scale=0.9]{sourcecodepro}
+
 ---
@@
 
+\newpage
+
 # Reviewer 1
@@
 
-_Minor comments_
+_**Minor comments**_
 
@@
 **Response:** This paper is currently available online. We have included the designation 'In Press' in the reference to reflect its status.
+
+\newpage
 
@@
 
-Response: Thank you for your feedback.
+**Response**: Thank you for your feedback.
 We appreciate your comments on the writing quality and your recommendations for improving the paper's contributions.
@@
 
+We have added the following paragraph in the revised manuscript.
+
 \textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
@@
 Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}
+
+\newpage
 
@@
 
-\textcolor{teal}Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}
+\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
+Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data.
+Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}
 
@@
 
-Response: The visualizations are intended to help the reader understand the effect and intent of different strategies.
+**Response**: The visualizations are intended to help the reader understand the effect and intent of different strategies.
 In determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
@@
 \textcolor{teal}{When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated where possible, regardless of which approach is employed, since they may improve forecasting model performance.}
+
+\newpage
 
@@
 
-_a) OntoCOVID: Ontology for Semantic Modeling of COVID19 Statistical Data Authors Shaukat Ali, Shah Khusro, Sajid Anwar, Abrar Ullah Publication date 2022 Book Proceedings of International Conference on Information Technology and Applications Pages 83-194 Publisher Springer, Singapore_
-
-_b) COVID-19 patient count prediction using LSTM M Iqbal, F Al-Obeidat, F Maqbool, S Razzaq, S Anwar, A Tubaishat, ... IEEE Transactions on Computational Social Systems 8 (4), 974-981_
-
-_c) Countering malicious URLs in internet of things using a knowledge-based approach and a simulated expert S Anwar, F Al-Obeidat, A Tubaishat, S Din, A Ahmad, FA Khan, G Jeon, ...
+  a) _OntoCOVID: Ontology for Semantic Modeling of COVID19 Statistical Data Authors Shaukat Ali, Shah Khusro, Sajid Anwar, Abrar Ullah Publication date 2022 Book Proceedings of International Conference on Information Technology and Applications Pages 83-194 Publisher Springer, Singapore_
+  b) _COVID-19 patient count prediction using LSTM M Iqbal, F Al-Obeidat, F Maqbool, S Razzaq, S Anwar, A Tubaishat, ... IEEE Transactions on Computational Social Systems 8 (4), 974-981_
+  c) _Countering malicious URLs in internet of things using a knowledge-based approach and a simulated expert S Anwar, F Al-Obeidat, A Tubaishat, S Din, A Ahmad, FA Khan, G Jeon, ...
 IEEE Internet of Things Journal 7 (5), 4497-4504_
@@
 We hope this clarifies the scope and focus of our research.
-
-# References {#references .unnumbered}

### 2024-05-17 fits.qmd
Added authors to first page

@@
   email: Rob.Hyndman@monash.edu
-  corresponding: true
 - name: Bahman Rostami-Tabar
@@
   email: rostami-tabarb@cardiff.ac.uk
+  corresponding: true
 abstract: "Forecasting interrupted time series data is a major challenge for forecasting teams, especially in light of events such as the COVID-19 pandemic.
@@
 wpnumber: no/yr
-jelcodes: C10,C14,C22
-blind: true
+jelcodes: ""
+blind: false
 cover: false

### 2024-08-19 fits.qmd
add Github link

@@
 Supply chain disruptions induced by factors such as strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps.
-\textcolor{teal}{For example, the war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of the COVID-19 toilet paper shortage.
+For example, the war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of the COVID-19 toilet paper shortage.
 A power outage in Eastern Germany halted production at a Tesla facility and cut off electricity to a large retailer distribution centre, disrupting supplies to around 500 stores in and around Berlin for several weeks.
 Recent strikes in various forms of transport, including air and rail in Germany, France, and the United Kingdom, have had an impact on passenger flow and retail sales at airports and train stations.
-These issues highlight the significance of robust forecasting systems in addressing such disruptions effectively.}
+These issues highlight the significance of robust forecasting systems in addressing such disruptions effectively.
 
@@
 In this section, we describe several possible strategies to handle interruptions when forecasting time series data.
-\textcolor{teal}{The handling of time series interruptions is a complex and multifaceted issue, and the choice of the most appropriate strategy depends on the specific characteristics of the data and the nature of the interruption.
-Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
+The handling of time series interruptions is a complex and multifaceted issue, and the choice of the most appropriate strategy depends on the specific characteristics of the data and the nature of the interruption.
+Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.
 
@@
 
-\textcolor{teal}{The disadvantage of smoothing and other highly adaptive techniques is that they need time to adapt after a disturbance and can eventually end up chasing noise.
-The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting.}
+The disadvantage of smoothing and other highly adaptive techniques is that they need time to adapt after a disturbance and can eventually end up chasing noise.
+The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting.
 
@@
 A special case of the intervention models proposed by @box1975intervention is a dynamic regression (or regression with ARIMA errors) model.
-This model includes covariates describing the interruption explicitly, with an ARIMA process describing the remaaining time series dynamics.
+This model includes covariates describing the interruption explicitly, with an ARIMA process describing the remaining time series dynamics.
 For example, if the intervention involves a simple level shift with a reverse level shift at the end of the intervention, we can use a dummy variable to indicate the interruption period and allow the model to adjust to the interruption.
@@
 This allows the change period to be effectively modelled, provided the intervention variables are chosen well.
-\textcolor{teal}{Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, may become subjective.} However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.
+Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, may become subjective.
+However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.
 
@@
 
-\textcolor{teal}{When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated where possible, regardless of which approach is employed, since they may improve forecasting model performance.}
+When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated where possible, regardless of which approach is employed, since they may improve forecasting model performance.
 
@@
 
-\textcolor{teal}{Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
+Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
 This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy.
@@
 In both cases, it does not make sense to evaluate the performance of the method during the disruption.
-Instead, the analyst needs to decide what approach makes the most sense for the application at hand.}
+Instead, the analyst needs to decide what approach makes the most sense for the application at hand.
 
-\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
+This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
 Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data.
-Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.} Comparative studies that assess the accuracy, robustness, and computing efficiency of different techniques across several datasets and disruption situations can help discover the most effective approaches for a range of practical applications.
+Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.
+Comparative studies that assess the accuracy, robustness, and computing efficiency of different techniques across several datasets and disruption situations can help discover the most effective approaches for a range of practical applications.
 
@@
 To enhance reproducibility and facilitate the adoption of proposed strategies, we provide the data and R code, as well as the entire paper written in R using [Quarto](https://quarto.org/) and the targets package for R [@targetsr2021].
-All materials required to reproduce this paper will be accessible via a public GitHub repository once the paper is accepted for publication.
+All materials required to reproduce this paper is accessible via the [GitHub repository](https://github.com/bahmanrostamitabar/forecasting_interrupted_time_series) .
 
@@
 
-We thank Dr Stephan Kolassa and several referees for their thoughts and feedback on the first version of our paper. These comments have helped us to refine and improve our work.
+We thank Dr. Stephan Kolassa and several referees for their thoughts and feedback on the first version of our paper. These comments have helped us to refine and improve our work.
 

### 2024-08-19 fits.tex
add Github link

@@
 \title{Forecasting interrupted time series}
-\date{17 May 2024}
+\date{19 August 2024}
 
@@
 \wp{no/yr}
-\jel{C10,C14,C22}
+\nojel
 
@@
 \nocover
-  \blind
-  \lfoot{\sf17 May 2024}
+  \author{Rob J Hyndman, Bahman Rostami-Tabar}
+  \addresses{%
+    %
+      \textbf{Rob J Hyndman}\\%
+      %
+        Department of Econometrics \& Business Statistics\\%
+        %
+        %
+        %
+      %
+        %
+        Monash University\\%
+        %
+        %
+      %
+        %
+        %
+        Clayton VIC \\%
+        %
+      %
+        %
+        %
+        %
+        Australia\\%
+      %
+        %
+        %
+        %
+        %
+      %
+      {Email: Rob.Hyndman@monash.edu}\\%
+      \\[0.5cm]%
+   %
+      \textbf{Bahman Rostami-Tabar}\\%
+      %
+        %
+        Cardiff Business School\\%
+        %
+        %
+      %
+        %
+        %
+        %
+        %
+      %
+        %
+        %
+        %
+        United Kingdom\\%
+      %
+      {Email: rostami-tabarb@cardiff.ac.uk}\\%
+      \textit{Corresponding author}\\[0.5cm]%
+   %
+   }%
+   \lfoot{\sf Hyndman, Rostami-Tabar: 19 August 2024}
 
@@
 irregularities in sales, inventory levels, production, and delivery,
-often resulting in data gaps.
-\textcolor{teal}{For example, the war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of the COVID-19 toilet paper shortage.
-A power outage in Eastern Germany halted production at a Tesla facility and cut off electricity to a large retailer distribution centre, disrupting supplies to around 500 stores in and around Berlin for several weeks.
-Recent strikes in various forms of transport, including air and rail in Germany, France, and the United Kingdom, have had an impact on passenger flow and retail sales at airports and train stations.
-These issues highlight the significance of robust forecasting systems in addressing such disruptions effectively.}
+often resulting in data gaps. For example, the war in Ukraine caused
+disruptions in sunflower oil and wheat supply chains, which were most
+likely driven by social media and people's recent memories of the
+COVID-19 toilet paper shortage. A power outage in Eastern Germany halted
+production at a Tesla facility and cut off electricity to a large
+retailer distribution centre, disrupting supplies to around 500 stores
+in and around Berlin for several weeks. Recent strikes in various forms
+of transport, including air and rail in Germany, France, and the United
+Kingdom, have had an impact on passenger flow and retail sales at
+airports and train stations. These issues highlight the significance of
+robust forecasting systems in addressing such disruptions effectively.
 
@@
 In this section, we describe several possible strategies to handle
-interruptions when forecasting time series data.
-\textcolor{teal}{The handling of time series interruptions is a complex and multifaceted issue, and the choice of the most appropriate strategy depends on the specific characteristics of the data and the nature of the interruption.
-Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}
+interruptions when forecasting time series data. The handling of time
+series interruptions is a complex and multifaceted issue, and the choice
+of the most appropriate strategy depends on the specific characteristics
+of the data and the nature of the interruption. Therefore, in this
+paper, we propose multiple strategies rather than a single approach to
+effectively address different situations and cases.
 
@@
 
-\textcolor{teal}{The disadvantage of smoothing and other highly adaptive techniques is that they need time to adapt after a disturbance and can eventually end up chasing noise.
-The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting.}
+The disadvantage of smoothing and other highly adaptive techniques is
+that they need time to adapt after a disturbance and can eventually end
+up chasing noise. The latter is an intriguing issue since individuals
+often prefer adaptability while failing to recognise that it does not
+necessarily improve forecasting.
 
@@
 with ARIMA errors) model. This model includes covariates describing the
-interruption explicitly, with an ARIMA process describing the remaaining
+interruption explicitly, with an ARIMA process describing the remaining
 time series dynamics. For example, if the intervention involves a simple
@@
 change period to be effectively modelled, provided the intervention
-variables are chosen well.
-\textcolor{teal}{Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, may become subjective.}
-However, the model will assume that the post-interruption period will be
-similar to the pre-interruption period, and so the prediction intervals
-may be too narrow, especially if there is a lasting effect beyond the
-end of the interruption.
+variables are chosen well. Intervention models are useful because they
+enable the forecasting of similar future disruptions, but distinguishing
+between a disruption and a general driving factor, such as a promotion
+or a seasonal storm, may become subjective. However, the model will
+assume that the post-interruption period will be similar to the
+pre-interruption period, and so the prediction intervals may be too
+narrow, especially if there is a lasting effect beyond the end of the
+interruption.
 
@@
 
-\textcolor{teal}{When forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated where possible, regardless of which approach is employed, since they may improve forecasting model performance.}
+When forecasting interrupted time series data, exogenous variables
+associated with specific events should be identified and incorporated
+where possible, regardless of which approach is employed, since they may
+improve forecasting model performance.
 
@@
 
-\textcolor{teal}{Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used.
-This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy.
-The strategies we discuss have different ways of handling the disrupted period, and many of them do not seek to provide accurate forecasts during the disruption.
-For example, the "set to missing" strategy ignores the disrupted period, while the "estimate what might have been" strategy replaces the observations during the disruption with interpolations.
-In both cases, it does not make sense to evaluate the performance of the method during the disruption.
-Instead, the analyst needs to decide what approach makes the most sense for the application at hand.}
-
-\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events.
-Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data.
-Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}
-Comparative studies that assess the accuracy, robustness, and computing
-efficiency of different techniques across several datasets and
-disruption situations can help discover the most effective approaches
-for a range of practical applications.
+Moreover, in determining which strategy to apply in each case, it is
+important to think through what is being assumed, and what data are
+being used. This is a qualitative and subjective process, and cannot
+simply be tackled with a quantitative comparison of forecast accuracy.
+The strategies we discuss have different ways of handling the disrupted
+period, and many of them do not seek to provide accurate forecasts
+during the disruption. For example, the ``set to missing'' strategy
+ignores the disrupted period, while the ``estimate what might have
+been'' strategy replaces the observations during the disruption with
+interpolations. In both cases, it does not make sense to evaluate the
+performance of the method during the disruption. Instead, the analyst
+needs to decide what approach makes the most sense for the application
+at hand.
+
+This study proposes multiple viable strategies for forecasting time
+series data interrupted by different type of events. Future research
+could focus on exploring new strategies for handling such disrupted time
+series, or enhancing probabilistic forecasting methods that are designed
+specifically for such data. Large-scale empirical research on further
+exploring which strategy may work best in different scenarios also
+provide useful insights. Comparative studies that assess the accuracy,
+robustness, and computing efficiency of different techniques across
+several datasets and disruption situations can help discover the most
+effective approaches for a range of practical applications.
 
@@
 package for R \autocite{targetsr2021}. All materials required to
-reproduce this paper will be accessible via a public GitHub repository
-once the paper is accepted for publication.
+reproduce this paper is accessible via the
+\href{https://github.com/bahmanrostamitabar/forecasting_interrupted_time_series}{GitHub
+repository} .
 
@@
 
-We thank Dr Stephan Kolassa and several referees for their thoughts and
+We thank Dr.~Stephan Kolassa and several referees for their thoughts and
 feedback on the first version of our paper. These comments have helped

### 2024-11-18 R/read.R
Removed attended incidents example

@@ -1,12 +1,0 @@
-create_tsibble <- function(attendant, verified) {
-  # Create tsibble
-  left_join(
-    attendant |>
-      pivot_longer(-date, names_to = "Area", values_to = "Attended"),
-    verified |>
-      pivot_longer(-date, names_to = "Area", values_to = "Verified"),
-    by = c("date", "Area")
-  ) |>
-    rename(Date = date) |>
-    as_tsibble(index = Date, key = Area)
-}

### 2024-11-18 _targets.R

@@ -3,14 +3,10 @@ library(targets)
 library(tarchetypes)
 
 # Set target options:
-tar_option_set(
-  packages = c("fpp3")
-  # Set other options as needed.
-)
+tar_option_set(packages = c("fpp3"))
 
-# Run the R scripts in the R/ folder with your custom functions:
+# Run the R scripts in the R/ folder
 tar_source()
-# source("other_functions.R") # Source other scripts as needed.
 
 # List of targets
 list(
@@ -81,44 +77,6 @@ list(
   tar_target(walkers_plot4, pedestrian_fc_plot(walkers_sol4, walkers)),
   tar_target(walkers_ensemble_plot, pedestrian_fc_plot(wensemble, walkers)),
 
-  # Attended incidents example
-  tar_target(attendant_csv, here::here("data/attendent_incident.csv"), format = "file"),
-  tar_target(verified_csv, here::here("data/verified_incident.csv"), format = "file"),
-  tar_target(attendant, readr::read_csv(attendant_csv)),
-  tar_target(verified, readr::read_csv(verified_csv)),
-  tar_target(df, create_tsibble(attendant, verified)),
-  tar_target(attended_plot, autoplot(df, Attended)),
-  tar_target(
-    stl, # STL Decomposition
-    df |>
-      filter(Area == "BC") |>
-      model(
-        stl = STL(Attended ~ trend(window = 81) + season(period = 7, window = Inf)),
-      )
-  ),
-  tar_target(
-    stl_plot,
-    stl |>
-      components() |>
-      autoplot()
-  ),
-  tar_target(
-    stl_model,
-    decomposition_model(
-      STL(Attended ~ trend(window = 81) + season(period = 7, window = Inf)),
-      ETS(season_adjust ~ season("N"))
-    )
-  ),
-  tar_target(
-    fit,
-    df |>
-      model(
-        ets = ETS(Attended),
-        arima = ARIMA(Attended),
-        stl_ets = stl_model
-      )
-  ),
-  tar_target(fc, forecast(fit, h = 365)),
   tar_quarto(paper,
     "fits.qmd",
     extra_files = c("references.bib", "header.tex")

### 2024-11-18 _targets_packages.R
Removed files not needed

@@ -1,13 +1,0 @@
-# Generated by targets::tar_renv(): do not edit by hand
-library(bs4Dash)
-library(clustermq)
-library(fpp3)
-library(future)
-library(gt)
-library(markdown)
-library(pingr)
-library(rstudioapi)
-library(shiny)
-library(shinybusy)
-library(shinyWidgets)
-library(visNetwork)

### 2024-11-19 fits.tex
Fixed renv update and recompiled

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
+
 \usepackage[style=authoryear-comp,]{biblatex}
@@
 }{}
-\makeatother% Placement of logos
+\makeatother
+% Placement of logos
 
@@
 \title{Forecasting interrupted time series}
-\date{19 August 2024}
+\date{19 November 2024}
 
@@
    }%
-   \lfoot{\sf Hyndman, Rostami-Tabar: 19 August 2024}
+   \lfoot{\sf Hyndman, Rostami-Tabar: 19 November 2024}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-tourism-plot1-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-tourism-plot1-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-tsol1-plot-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-tsol1-plot-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-tsol2-plot-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-tsol2-plot-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-tsol3-plot-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-tsol3-plot-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-tsol4-plot-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-tsol4-plot-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-tensemble-plot-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-tensemble-plot-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-walkers-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-walkers-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot1-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-walkers-plot1-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot2-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-walkers-plot2-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot3-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-walkers-plot3-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-walkers-plot4-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-walkers-plot4-1.pdf}}
 
@@
 
-\includegraphics{fits_files/figure-pdf/fig-walkers-ensemble-plot-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{fits_files/figure-pdf/fig-walkers-ensemble-plot-1.pdf}}
 

### 2024-11-19 Old_Rcode/pedestrians.R
Added "original" R code (pre targets)

@@ -1,0 +1,217 @@
+## PEDESTRIANS EXAMPLE ----------------------------------------------------------
+
+## Read data --------------------------------------------------------------------
+
+# Download hourly data from start to end
+if (fs::file_exists(here::here("data/raw_walkers.rds"))) {
+  raw_walkers <- readRDS(here::here("data/raw_walkers.rds"))
+} else {
+  raw_walkers <- rwalkr::melb_walk(from = as.Date("2019-01-01"), to = as.Date("2023-12-31")) |>
+    arrange(Sensor, Date, Time) |>
+    group_by(Sensor, Date) |>
+    mutate(obs = row_number() - 1) |>
+    filter(obs > 0) |>
+    select(-obs) |>
+    mutate(Count = readr::parse_integer(Count))
+  saveRDS(raw_walkers, here::here("data/raw_walkers.rds"))
+}
+
+## Find sensors that are used throughout the whole period ------------------------
+
+sensors_to_use <- raw_walkers |>
+  group_by(Sensor, Date_Time) |>
+  summarise(Count = sum(Count), .groups = "drop") |>
+  as_tsibble(index = Date_Time, key = Sensor) |>
+  fill_gaps(.full = TRUE) |>
+  as_tibble() |>
+  group_by(Sensor) |>
+  summarise(nmiss = sum(is.na(Count))) |>
+  filter(nmiss < 30 * 24) |>
+  pull(Sensor)
+
+# Compute daily averages across all sensors -------------------------------------
+
+walkers <- raw_walkers |>
+  filter(Sensor %in% sensors_to_use) |>
+  group_by(Date, Sensor) |>
+  summarise(Count = sum(Count), .groups = "drop") |>
+  group_by(Date) |>
+  summarise(Count = mean(Count, na.rm = TRUE) / 1e3, .groups = "drop") |>
+  tsibble::as_tsibble(index = Date)
+
+## Lockdown periods ------------------------------------------------------------
+
+# First lockdown officially started 31 March, but self-lockdown happened from
+# start of previous week. Other dates below are official periods as per
+# https://en.wikipedia.org/wiki/COVID-19_pandemic_in_Victoria#Lockdown_statistics
+# See also https://www.platinumaccounting.com.au/blog/melbourne-lockdown-dates
+lockdowns <- tribble(
[…]

### 2024-11-19 Old_Rcode/tourism.R

@@ -1,0 +1,189 @@
+## TOURISM EXAMPLE -------------------------------------------------------------
+
+library(fpp3)
+
+# Read data --------------------------------------------------------------------
+
+austa <- readxl::read_excel(
+    path = here::here("data/340101.xlsx"),
+    sheet = "Data1", skip = 9
+  ) |>
+  transmute(
+    Month = as.Date(`Series ID`),
+    Visitors = A85375847A / 1e3
+  ) |>
+  filter(Month >= "2000-01-01")
+
+# Tourism time plot ------------------------------------------------------------
+
+austa |>
+  ggplot(aes(x = Month, y = Visitors)) +
+  geom_line() +
+  labs(
+    y = "Thousands of visitors",
+    title = "Total short-term visitors to Australia"
+  )
+
+# Convert to tsibble -----------------------------------------------------------
+
+austa <- austa |>
+  mutate(Month = yearmonth(Month)) |>
+  as_tsibble(index = Month)
+
+# Solution 1 -------------------------------------------------------------------
+
+tsol1 <- austa |>
+  stretch_tsibble(.step = 12, .init = 240) |>
+  model(
+    ets = ETS(log(Visitors)),
+    arima = ARIMA(log(Visitors))
+  ) |>
+  forecast(h = 12) |>
+  mutate(.id = paste("Forecasts of", .id + 2019))
+
+# Solution 2 -------------------------------------------------------------------
+
+austa_stretch <- austa |>
+  mutate(
+    covid = as.numeric(Month >= yearmonth("2020 Mar") &
+      Month <= yearmonth("2022 Oct")),
+    recovery = Month >= yearmonth("2021 Oct") &
+      Month <= yearmonth("2022 Oct"),
+    recovery = recovery * (Month - yearmonth("2021 Sep"))
+  ) |>
+  stretch_tsibble(.step = 12, .init = 240)
+
+# Fit models using none, one or two covariates
+fit1 <- austa_stretch |>
+  model(arima1 = ARIMA(log(Visitors)))
+fit2 <- austa_stretch |>
+  model(arima2 = ARIMA(log(Visitors) ~ covid)) |>
+  suppressWarnings()
+fit3 <- austa_stretch |>
+  model(arima3 = ARIMA(log(Visitors) ~ covid + recovery)) |>
+  suppressWarnings()
+
[…]

### 2024-11-19 _targets.R
Added random seed

@@ -4,6 +4,7 @@ library(tarchetypes)
 
 # Set target options:
 tar_option_set(packages = c("fpp3"))
+tar_option_set(seed = 3)
 
 # Run the R scripts in the R/ folder
 tar_source()

### 2024-11-19 fits.qmd
Added jors template

@@
     - department: Department of Econometrics & Business Statistics
-    - name: Monash University
-    - city: Clayton VIC
-    - country: Australia
-    - postal-code: 3800
+      name: Monash University
+      city: Clayton VIC
+      country: Australia
+      postal-code: 3800
   email: Rob.Hyndman@monash.edu
+  orcid: 0000-0002-2140-5352
 - name: Bahman Rostami-Tabar
@@
     - name: Cardiff Business School
-    - postal-code: Wales CF10 3EU
-    - country: United Kingdom
+      postal-code: Wales CF10 3EU
+      country: United Kingdom
   email: rostami-tabarb@cardiff.ac.uk
+  orcid: 0000-0002-3730-0045
   corresponding: true
@@
 bibliography: references.bib
-wpnumber: no/yr
+format:
+  wp-pdf+wp:
+    include-in-header: header.tex
+    keep-tex: true
+    linestretch: 1.5
+    wpnumber: no/24
     jelcodes: ""
     blind: false
-cover: false
-linestretch: 1.5
-format:
-  wp-pdf:
+    cover: true
+  tandf-pdf:
     include-in-header: header.tex
@@
 library(targets)
-library(fpp3)
 
@@
 )
-#theme_set(theme_get() + theme(text = element_text(family = "Fira Sans")))
 ggplot2::theme_set(ggthemes::theme_few())
-# Load main data objects from targets folder
-tar_config_set(store = here::here("_targets"))
-tar_load(c(austa))
+# Load austa data 
+tar_load(austa)
 ```
@@
 We fit ETS, ARIMA, and dynamic ARIMA models to the data [@fpp3], first applying a log transformation to ensure the resulting forecasts are positive.
-(A bias correction is also applied to ensure the back-transformed forecasts are the means of the forecast distributions.) A log transformation is possible because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> filter(Visitors == min(Visitors)) |> pull(Month) |> format("%B %Y")`.
+(A bias correction is also applied to ensure the back-transformed forecasts are the means of the forecast distributions.) A log transformation is possible because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> dplyr::filter(Visitors == min(Visitors)) |> dplyr::pull(Month) |> format("%B %Y")`.
 For all forecasts, we also show 90% prediction intervals.
@@
 lockdowns <- lockdowns |>
-  mutate(Duration = End - Start + 1)
+  dplyr::mutate(Duration = End - Start + 1)
 ```

### 2024-11-19 Old_Rcode/tourism.R

@@ -20,6 +20,7 @@ austa |>
   ggplot(aes(x = Month, y = Visitors)) +
   geom_line() +
   labs(
+    x = "Month",
     y = "Thousands of visitors",
     title = "Total short-term visitors to Australia"
   )
